---
title: "Thematic accounts - encroachment extent accounts in recreation areas"
subtitle: 
# delete the line with the second indicatorID (line 5) 
# if it's not relevant for you.
# In the parentheses you can add a word to distinguish multiple indicators 
# from each other, for example the ecosystem type that they cover
indicatorID: 
  - ENCACC_001
  - ENCACC_002 # delete whole line if not needed
# If you are presenting multiple indicators, and have multiple 
# indicator IDs, then add them as a list here, and make sure you add
# the same number of list entries, 
# and in the same order that corresponds to the indicator IDs,  
# to the meta data fields where the indicators differ. 
# For example, if the indicators cover different ecosystem types, 
# add the ecosystems as a list under 'Ecosystems', but if they are in the
# same realm, same country etc., then you only need to write the realm and
# country once. 
indicatorName: 
  - Extent accounts of all encroachment in municipalities # add more by adding another line starting '-'
  - Extent accounts of all sværlig vikitig natur lost due to encroachments in municipalities
  
folderName: code # name of the folder under 'indicators/'. 
# If it's a single indicator being reported, the folder name should be the same as the indicatorID. 
# If there are multiple ID's then the folder name could look something like XX_YYY_ZZZ-ÅÅÅ.
country: Norway
continent: Europe
# ECT: write ECT here # e.g. A2 - Chemical State characteristics
Realm: 
  - write realm here # e.g. Freshwater (F)
Biome: 
  - write biome here # e.g. F1 Rivers and streams biome
Ecosystem: 
  - write ecosystem here # e.g. F1.3 Freeze-thaw streams
yearAdded: 2025 # the year when the indicator was first submitted to ecRxiv
yearLastUpdate: 2025
status: incomplete # options are complete, incomplete, or deprecated
Version: "000.001" # Beta-versions are numbers incrementally with values below 0. Operational versions can be given the version numbers above 0. Use this terms similar as for software versioning,
VersionComment: This is the first version of the thematic encroachment extent accounts in municipalities being compiled # Describe briefly how this version differs from earlier versions 
url: https://github.com/NINAnor/recereation_encroachment_accounts
# Leave the url as is, but replace XX_YYYY_ZZ with the name of the root folder for the current indicator(s). It should be the same as the indicatorID, but if there are more indicator IDs presented in the same folder, the folder name might be different, e.g. XX_YYYY_AAA-BBB 
Normalised: No # Yes or no depending on whether the metric has been normliased or not.
SpatialAggregationPathway: Describe the aggregation using the standrdaise terminology # e.g. Scale - Transform - Aggregate (length-weighted arithmetic mean) - Truncate
format: 
  html:
    embed-resources: true
    code-fold: true
    toc: true
    toc-title: Contents
    toc-depth: 3
    smooth-scroll: true
execute: 
  cache: true
author:
  - name: Sylvie Clappe             # Enter name
    email: sylvie.clappe@nina.no  # Enter email
    affiliations:
      - id: myID
        name: Norwegian Institute for Nature Research  # Enter affiliations
  - name: David Barton              #  Enter subsequent authors like this, or remove if not relevant
    affiliations:
      - ref: myID               # To reuse affiliations reference the id like this
  - name: Reto Spielhofer             #  Enter subsequent authors like this, or remove if not relevant
    affiliations:
  - ref: myID 
AuthorList: Clappe, S., Barton, D.N., Spielhofer, R.  
date: January 22, 2026 # Enter date
data_availability: none
code_reproducibility: none
open_science_badge: none
callout-icon: false
lightbox: true
css: ../../../style.css
code-links:
      - text: Add a review
        icon: github
        href: https://github.com/NINAnor/ecRxiv
hide: TRUE # Change to FALSE
---

<!--# This is a template for how to document the indicator analyses. Make sure also to not change the order, or modify, the headers, unless you really need to. This is because it easier to read if all the indicators are presented using the same layout. If there is one header where you don't have anything to write, just leave the header as is, and don't write anything below it. If you are providing code, be careful to annotate and comment on every step in the analysis. Before starting it is recommended to fill in as much as you can in the metadata file. This file will populate the initial table in your output.-->

<!--# Load all you dependencies here -->

```{r setup}
#| include: false
#| cache: false
library(knitr)
library(tidyverse)
library(kableExtra)
library(here)
library(anybadger)
library(yaml)
library(tibble)
library(conflicted)
library(httr2)
library(giscoR)
library(zen4R)
library(jsonlite)
library(sf)
library(purrr)
library(numform)
library(DT)
library(tmap)

# weird workarounds for cache issues
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::lag)
conflicts_prefer(dplyr::select)
pull<-dplyr::pull

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Remove dplyr and messages
options(dplyr.summarise.inform = FALSE)
```

```{r source}
#| echo: false
#| cache: false
source(here::here("_common.R"))
```

```{r working_directory}
#| include: false
#| cache: false

# Workaround to force the working directory
# Needs to be set after {r source} otherwise
# {_common.R is not found}
opts_knit$set(root.dir = here:::here())
```


<!--# The following parts are autogenerated. Do not edit. -->

::: {layout-ncol="5"}

```{r statusBadge}
#| echo: false
status_badge(as.character(st))
```

```{r versionBadge}
#| echo: false
version_badge(my_version_number = as.character(version[[1]]),
  folder_name = folder_name)
```

```{r dataBadge}
#| echo: false
data_badge(as.character(badges[[1]]))
```

```{r codeBadge}
#| echo: false
code_badge(as.character(badges[[2]]))
```

```{r openScienceBadge}
#| echo: false
open_science_badge(as.character(badges[[3]]))
```
::: 
 

> **Recommended citation**: `r paste(auth, " ", year, ". ", name, " (ID(s): ", folder_name, ") ", "v. ", version, ". ", url, sep="")`



::: {.callout-tip collapse="true"}
## Metadata
```{r tbl-meta2}
#| tbl-cap: 'Indicator metadata'
#| echo: false
#| warning: false
meta |>
  dplyr::filter(Variable %in% c(
    "Indicator ID",
    "Indicator Name",
    "Continent",
    "Country",
    "Ecosystem Condition Typology Class", 
    "Realm",
    "Biome",
    "Ecosystem",
    "Year added", 
    "Last update",
    "Version",
    "Version comment",
    "Normalised",
    "Spatial aggregation pathway")) |>   
  kable()

```
:::

::: {.callout-tip collapse="true"}

## Log

<!--# Update this logg with short messages for each update -->
- 22 January 2026 - Analyses completed
- XXX - First draft version
:::


<hr />

<!--# Document you work below.  -->

## 1. Les meg 
Norway is losing nature at high speed. According to the NRK article “Norgei Rødt, Hvitt og Grått”, Norway lost 208 km2 of semi-natural lands to artificial land expansion; a process known as land take[^1,2] . Almost a year later, in October 2025, NRK released another article looking at the rate of land take in Europe, «Europe from Green to Grey – Norway has the highest loss of nature per capita in Europe». In this article, Norway had the alarming rate of more than  38 m2 of nature loss per inhabitant in the period 2018-2024 (NRK, 2025). 

Although we know what is happening at the scale of the nation, we don't know what is happening within outdoor recreation areas. How much natur is being lost in these area due to land take? Also, what type of nature was lost? To answer these question, we compile arealregnskap over 2018-2024 for:

1. All nedbygging whitin the municipality / fylke

2. Nedbygging whitin all of outdoor recreation areas per municipality / fylke

3. Nedbygging whitin important and very important outdoor recreation areas per municipality / fylke

4. Versatte naturtyper tapt within 1, 2 and 3 above.


## 2. Hvilke datasett brukte vi?
### 2.1 Kart over kommunegrensene i Norge 



### 2.2 Kart over området og verdiene av kommunale friluftsområder



### 2.3 Kart over potensielle «nedbygging» i Norge i perioden 2018–2024




### 2.3 Kart over habitatene som anses som «særlig viktig natur»


## 5. Uncertainties
Resultatene må være pålitelige for å ta informerte beslutninger. Rapporten bruker et tra-fikklyssystem for å vise viser relativ (u)sikkerhet i resultatene fra din kommune og / eller fylke: 

FIGURE

## 6. Fylke tematisk arealregnskap for nedbygging
### 6.1 Totalt areal av nedbygging
```{r ENCACC001}
#| eval: false
#| warning: false
#| message: false
#| output: true
#| echo: false

# Load the accounting table: encacc_001_acc

# Display the table dynamically
datatable(encacc_002_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list( dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                          columnDefs = list(list(targets = c(which(colnames(encacc_002_acc) == "Usikkerhet for areal i regnskapsperiode")),
                                                 visible = FALSE))),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Tabel 3 - Tematiske arealregnskap: areal av særlig viktig natur tapt på grunn av nedbygging i kommuner i perioden 2018–2024."))) %>%
    formatCurrency(
    columns = c("Areal særlig viktige naturtyper gått tapt (dekar)"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",") %>%
  ## Example: color Årlig areal i dekar by var1
  formatStyle(
    "Areal særlig viktige naturtyper gått tapt (dekar)",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red", "NA"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A", "")
    ),
    valueColumns = "Usikkerhet for areal i regnskapsperiode"           # column used for the intervals
  )

```

### 6.2 Totalt areal av nedbygging i friluftlivsområder

### 6.3 Totalt areal av verdsatte «verdsatte naturtyper» av «stor verdi« og «svært stor verdi» tapt til nedbygging
```{r ENCACC002}
#| eval: false
#| warning: false
#| message: false
#| output: true
#| echo: false

# Load the accounting table: encacc_001_fy_acc

# Display the table dynamically
datatable(encacc_002_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list( dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                          columnDefs = list(list(targets = c(which(colnames(encacc_002_acc) == "Usikkerhet for areal i regnskapsperiode")),
                                                 visible = FALSE))),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Tabel 3 - Tematiske arealregnskap: areal av særlig viktig natur tapt på grunn av nedbygging i kommuner i perioden 2018–2024."))) %>%
    formatCurrency(
    columns = c("Areal særlig viktige naturtyper gått tapt (dekar)"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",") %>%
  ## Example: color Årlig areal i dekar by var1
  formatStyle(
    "Areal særlig viktige naturtyper gått tapt (dekar)",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red", "NA"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A", "")
    ),
    valueColumns = "Usikkerhet for areal i regnskapsperiode"           # column used for the intervals
  )

```

### 6.3 Totalt areal av verdsatte «verdsatte naturtyper» av «stor verdi« og «svært stor verdi» tapt til nedbygging i friluftlivsområder

## 7. Kommune tematisk arealregnskap for nedbygging
### 6.1 Totalt areal av nedbygging

### 6.2 Totalt areal av nedbygging i friluftlivsområder

### 6.3 Totalt areal av verdsatte «verdsatte naturtyper» av «stor verdi« og «svært stor verdi» tapt til nedbygging

### 6.3 Totalt areal av verdsatte «verdsatte naturtyper» av «stor verdi« og «svært stor verdi» tapt til nedbygging i friluftlivsområder

## 8. References
(Waiting, 2024; Venter et al., 2024)

::: {.callout-tip collapse="true"}
## Session Info

This information is mainly aimed at developers that need to recreate the R environment that produced the html.
Important note: this information is about the R session that was running when the html was rendered, and if some of the code chunks were not evaluated when the html was rendered, they may have been executed in a different environment.  

```{r sessionInfo}
#| echo: false
sessionInfo()
```

:::
