---
title: "Thematic accounts - land take extent accounts in Norway"
subtitle: |
  [ENCACC_001] (Extent accounts of all land take in municipalities) \
  [ENCACC_005] (Extent accounts of specaially important nature lost due to land take in municipalities)
# delete the line with the second indicatorID (line 5) 
# if it's not relevant for you.
# In the parentheses you can add a word to distinguish multiple indicators 
# from each other, for example the ecosystem type that they cover
indicatorID: 
  - ENCACC_001
  - ENCACC_002 # delete whole line if not needed
# If you are presenting multiple indicators, and have multiple 
# indicator IDs, then add them as a list here, and make sure you add
# the same number of list entries, 
# and in the same order that corresponds to the indicator IDs,  
# to the meta data fields where the indicators differ. 
# For example, if the indicators cover different ecosystem types, 
# add the ecosystems as a list under 'Ecosystems', but if they are in the
# same realm, same country etc., then you only need to write the realm and
# country once. 
indicatorName: 
  - Extent accounts of all land take in municipalities # add more by adding another line starting '-'
  - Extent accounts of verdsatte natur av "stor" og "svært stor" verdi lost due to land take in municipalities
  
folderName: code # name of the folder under 'indicators/'. 
# If it's a single indicator being reported, the folder name should be the same as the indicatorID. 
# If there are multiple ID's then the folder name could look something like XX_YYY_ZZZ-ÅÅÅ.
country: Norway
continent: Europe
# ECT: write ECT here # e.g. A2 - Chemical State characteristics
Realm: 
  - all terrestrial # e.g. Freshwater (F)
Biome: 
  - all terrestrial # e.g. F1 Rivers and streams biome
Ecosystem: 
  - all terrestrial # e.g. F1.3 Freeze-thaw streams
yearAdded: 2026 # the year when the indicator was first submitted to ecRxiv
yearLastUpdate: 2026
status: complete # options are complete, incomplete, or deprecated
Version: "000.001" # Beta-versions are numbers incrementally with values below 0. Operational versions can be given the version numbers above 0. Use this terms similar as for software versioning,
VersionComment: This is the first version of the thematic land take extent accounts in Norway's municipalities and counties being compiled # Describe briefly how this version differs from earlier versions 
url: https://github.com/NINAnor/recereation_encroachment_accounts
# Leave the url as is, but replace XX_YYYY_ZZ with the name of the root folder for the current indicator(s). It should be the same as the indicatorID, but if there are more indicator IDs presented in the same folder, the folder name might be different, e.g. XX_YYYY_AAA-BBB 
Normalised: No normalisation, except when ENCACC_001-002 are presented as percentages. In that case we normalised by the total area of the municiplaities / counties (ENCACC_001); or the total area of land take whithin municipalities / counties (ENCACC_002). # Yes or no depending on whether the metric has been normliased or not.
SpatialAggregationPathway: Area of land take calculated at polygon level, then aggregated at municipal level, then aggregated at county and national levels. Standard or cumulative sums were used. # e.g. Scale - Transform - Aggregate (length-weighted arithmetic mean) - Truncate
format: 
  html:
    embed-resources: true
    code-fold: true
    toc: true
    toc-title: Contents
    toc-depth: 3
    smooth-scroll: true
execute: 
  cache: true
author:
  - name: Sylvie Clappe             # Enter name
    email: sylvie.clappe@nina.no  # Enter email
    affiliations:
      - id: myID
        name: Norwegian Institute for Nature Research  # Enter affiliations
  - name: David Barton              #  Enter subsequent authors like this, or remove if not relevant
    affiliations:
      - ref: myID               # To reuse affiliations reference the id like this
  - name: Reto Spielhofer             #  Enter subsequent authors like this, or remove if not relevant
    affiliations:
      - ref: myID 
  - name: Zander Venter             #  Enter subsequent authors like this, or remove if not relevant
    affiliations:
      - ref: myID 
AuthorList: Clappe, S., Barton, D.N., Spielhofer, R., Venter, Z.  
date: March 04, 2026 # Enter date
data_availability: none
code_reproducibility: none
open_science_badge: none
callout-icon: false
lightbox: true
css: ../../../style.css
code-links:
      - text: Add a review
        icon: github
        href: https://github.com/NINAnor/ecRxiv
hide: TRUE # Change to FALSE
---

<!--# This is a template for how to document the indicator analyses. Make sure also to not change the order, or modify, the headers, unless you really need to. This is because it easier to read if all the indicators are presented using the same layout. If there is one header where you don't have anything to write, just leave the header as is, and don't write anything below it. If you are providing code, be careful to annotate and comment on every step in the analysis. Before starting it is recommended to fill in as much as you can in the metadata file. This file will populate the initial table in your output.-->

<!--# Load all you dependencies here -->

```{r setup}
#| include: false
#| cache: false
library(knitr)
library(tidyverse)
library(kableExtra)
library(here)
library(anybadger)
library(yaml)
library(tibble)
library(conflicted)
library(httr2)
library(giscoR)
library(zen4R)
library(jsonlite)
library(sf)
library(purrr)
library(numform)
library(DT)
library(tmap)

# weird workarounds for cache issues
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::lag)
conflicts_prefer(dplyr::select)
pull<-dplyr::pull

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Remove dplyr and messages
options(dplyr.summarise.inform = FALSE)
```

```{r source}
#| echo: false
#| cache: false
source(here::here("_common.R"))
```

```{r working_directory}
#| include: false
#| cache: false

# Workaround to force the working directory
# Needs to be set after {r source} otherwise
# {_common.R is not found}
opts_knit$set(root.dir = here:::here())
```


<!--# The following parts are autogenerated. Do not edit. -->

::: {layout-ncol="5"}

```{r statusBadge}
#| echo: false
status_badge(as.character(st))
```

```{r versionBadge}
#| echo: false
version_badge(my_version_number = as.character(version[[1]]),
  folder_name = folder_name)
```

```{r dataBadge}
#| echo: false
data_badge(as.character(badges[[1]]))
```

```{r codeBadge}
#| echo: false
code_badge(as.character(badges[[2]]))
```

```{r openScienceBadge}
#| echo: false
open_science_badge(as.character(badges[[3]]))
```
::: 
 

> **Recommended citation**: `r paste(auth, " ", year, ". ", name, " (ID(s): ", folder_name, ") ", "v. ", version, ". ", url, sep="")`



::: {.callout-tip collapse="true"}
## Metadata
```{r tbl-meta2}
#| tbl-cap: 'Indicator metadata'
#| echo: false
#| warning: false
meta |>
  dplyr::filter(Variable %in% c(
    "Indicator ID",
    "Indicator Name",
    "Continent",
    "Country",
    "Ecosystem Condition Typology Class", 
    "Realm",
    "Biome",
    "Ecosystem",
    "Year added", 
    "Last update",
    "Version",
    "Version comment",
    "Normalised",
    "Spatial aggregation pathway")) |>   
  kable()

```
:::

::: {.callout-tip collapse="true"}

## Log

<!--# Update this logg with short messages for each update -->
- 25 February 2026 - Analyses completed
- 04 March 2026 - First version
:::


<hr />

<!--# Document you work below.  -->

## 1. Summary
<!--# 
With a maximum of 300 words, describe the indicator in general terms as you would to a non-expert. Think of this as a kind of common language summary. Follow the four numbers sub-headers below. under point 2, it is possible under  include a bullet point list of the specific steps in the workflow. 
 -->

1. **Background and rationale**:

Norway is losing nature at high speed. According to the NRK article “Norge i Rødt, Hvitt og Grått”, Norway lost 208 km2 of semi-natural lands to artificial land expansion; a process known as land take (Venter et al., 2024a,b,c). Almost a year later, in October 2025, NRK released another article looking at the rate of land take in Europe, «Europe from Green to Grey – Norway has the highest loss of nature per capita in Europe». In this article, Norway had the alarming rate of more than  38 m2 of nature loss per inhabitant in the period 2018-2024 (NRK, 2025a). 

In this document, we explore how much of this nature loss happened in both counties and municipalities and what type of nature was lost. To be relevant in a planning context, we also focused on the nature types referred to Environmental Administration's Objection Practice under the collective term «nature of particular importance».

2. **Methods**:

We created extent accounts for land take within the county and municipalities as a whole. Three main datasets were used to create our accounts: (i) a map of municipal administrative boundaries from Eurostat; (ii) maps of built-up expansion over Norway 2017-2014 from Venter 2025;  and (iii) a map of valued nature types from the Norwegian Environmental Agency. The analyses mostly relied on various spatial intersection of the data. 

We also developed a simple traffic-light system to inform users of the uncertainty of our results. This system is mainly based on the user accuracy of the prediction of land take, and the temporal discrepancy between datasets.

3. **Main headlines**:

- Despite a total of about 362 000 decares (362 km2) of nature lost in Norway due to land take between 2018 and 2024, the area of nature disappearing each year is decreasing since 2021.

- Innlandet saw the largest loss with more than 7 600 dekar of nature disappearing annualy over 2018-2024. While Oslo lost the smallest area with about 1 100 decares over seven years.

- Over 5 200 decares of specially important nature disappeared due to land take be-tween 2018 and 2024

- Over 2 300 decares of coastal heathlands disappeared all over Norway due to land take between 2018 and 2024.

- With almost 4,5%, Roagaland has the highest percentage of land take built on spe-cially important nature. While less than 1% of the nature lost to nedbygging in Østfold, Troms, Agder, Oslo was specially important nature.



## 2. About the underlying data
We used three main datasets:

- **DT01**: [Local Administrative Units](^https://ec.europa.eu/eurostat/web/gisco/geodata/statistical-units/local-administrative-units$): Map of the administrative boundaries of municipalities in Norway. It is pro-duced by Eurostat. We chose not to use the Kartverket map of the munic-ipalities as their boundaries extend to the marine habitats. Eurostat’s boundaries are restricted to the ter-restrial habitats. As DT02 is limited to terrestrial habitats, we chose to restrict the extent of outdoor recrea-tion areas to terrestrial habitats only.


- **DT02**: [Map of built-up expansion ("nedbygging") over Norway 2017-2024 with year of loss, Venter (2025)](^https://zenodo.org/records/15130992$): Map of potential “nedbygging” (i.e., land take) event in Norway over 2018-2024. This data have been using Sentinal-2 satellite imagery and  a convolutional neural network from Google called Dynamic World.

- **DT03**: [Naturtyper - verdsatte](^https://kartkatalog.geonorge.no/metadata/naturtyper-verdsatte/64cbb884-a19d-4356-a114-380cfe4a7314$): Map of the habitats considered as specially important nature  in accordance with the valuation criteria in guideline M-1941 Impact assessments for climate and the environment.

<!--# Give a general introduction the datasets you have used, and provide more specific details in the sub-chapters.  -->

### 2.1 Spatial and temporal resolution and extent

- **DT01**: vector data last updated in 2025 and covering 2024 administrative boundaries. DT01 covers all of Norway's municipalities.

- **DT02**: vector data last updated in 2025 and covering the land take that appeared between 2018 and 2024. DT02 has a national coverage.

- **DT03**: vector data last updated in 2025 and whose mapping years of habitats range between 1899 and 2025. DT03 does not have a national coverage (see bias below).

<!--# Describe the temporal and spatial resolution and extent of the data used -->

### 2.2 Sampling design

- **DT01**: Not applicable

- **DT02**: detection of encroachment is based on Sentinel-2 data and a convulotional neural network from Google called Dynamic World. These data are associated with a 85% user accuracy, this means that the detected encroachment are incorrect 18% of the time. This acuracy however varies depending on the year of detection of the land take: 80.6% for 2018; 81.8 for 2019; 82.2 for 2020; 84% for 2021; 86% for 2022; 95% for 2023. 2024 is not reported, we will assume it is equal to the overall user accuracy of 82%.

- **DT03**: No noational coverage or revisit or sites. This means two things: (i) that if a habitat is not mapped at a particular location, we do not know if it's because it is indeed not there or that it simply hasn't been mapped yet; (ii) once a habitat has been mapped, there is no temporal indication for ensuring it is still there over the years.

<!-- How was the data sampled? Describe potential sampling biases -->

### 2.3 Original units

- **DT01**: Municipalities

- **DT02**: Land take polygons

- **DT03**: Habitats polygons
<!--# What are the original units for the most relevant  variables in the data-->


### 2.3 Instructions for citing, using and accessing data

- **DT01**: [Acces here](^https://ec.europa.eu/eurostat/web/gisco/geodata/statistical-units/local-administrative-units$). Can be cited as: Eurostat (2025). Local Administrative Units (LAU) 2024. 

- **DT02**: [Access here](^https://zenodo.org/records/15130992$). Can be cited as: Miljødirektoratet (2025). Verdsatte naturtyper 2025. 

- **DT03**: [Access here](^https://kartkatalog.geonorge.no/metadata/naturtyper-verdsatte/64cbb884-a19d-4356-a114-380cfe4a7314$). Can be cited as: Venter, Z. (2025). Map of built-up expansion ("nedbygging") over Norway 2017-2024 with year of loss [Data set]. Zenodo. https://doi.org/10.5281/zenodo.15130992

<!--# Is the data openly available? If not, how can one access it? What are the key references to the datasets?   -->

### 2.4 Additional comments about the dataset

- **DT01**: was used to delineate the administrative boundaries of the municipalities and counties.

- **DT02**: was used to identify the area of land take.



<!--# Optional. Text here -->

## 3. Indicator properties

### 3.1 Mis-alignment with the System of Environmental Economic Accounting - Ecosystem Accounting
For ENCACC_001, we have two deviations from the SEEA-EA traditional extent accounts:

1.	A land take represent a land change over time. The SEEA-EA uses an “ecosystem-based” approach referring to ecosystem types and assets, whereby we record the extent in the opening and closing year of the accounting period, as well as a net change. land take being already a change, the accounts that are created will be for a land change which is not an “ecosystem” asset  per se. Here, we however decided to consider the land change as an “ecosystem” and report an opening (2018) and closing (2024) extent for the land change. This led to a subsequent deviation regarding the calculation of a total change in land take extent during the accounting period (see point 2 below).

2. SEEA-EA extent accounts report the total area of an ecosystem during the accounting period and details the reductions, additions and net changes in this extent. As per point 1 above, land take intrinsically represent change in land cover only as the data only record the area of new land take in “natural areas” each year. Even though we consider it as a “pseudo-ecosystem” in our extent accounts, it meant that we could not calculate a change over the opening and closing years in the traditional way the SEEA-EA presents by subtracting the yearly opening extent to the yearly closing extent. To counteract this, we decided to also compile the cumulative area so we could compile a change in area between the opening and closing year. However, as previously noted, the land take data only record the apparition of new land take each year. This means that we only report the additions in area over the years, not the reductions (e.g., restoration works). In other words, we report when an encroachment appear but not when it disappears. There is thus a bias in the way the information is presented and the total change we present over the accounting period (2018-2024) is not a “net change”. Overall this means that we do not compile full SEEA-EA extent accounts, but only partial ones.

For ENCACC_002, we only presented the total area lost over the accounting period per habitat type. This was to simplify the interpretation by the municpalities as areas are not lost constantly over time.

<!--# Describe known collinearities with other metrics (indicators or variables) that could become problematic if they were also included in the same Ecosystem Condition Assessment as the indicator described here. -->


### 3.2 Spatial aggregation

For both ENCACC_001 and ENCACC_002, areas were first calculated at the municipal level, then aggregated at county and national level. 
When we present the ENCACC_001 map of the land take as a share of the total municipal / county area, we normalise the total land take area at the municipal / county level by the total administrative area of the municipality / county. 

When we present the ENCACC_002 map of percentage of land take that was specially important nature, we normalise the total area of specially important nature lost to land take by the total area of land take for the municipality / county. 

<!-- Describe in word the same aggregation pathway presented in the top yaml (the metadata). If you have scaled against the reference value, it is especially important to know if you did any avergaing or summations etc. beforehand. An example text for a rather simple case could be: "The variable is scaled at the initial polygon level and the point estimate is a area-weighted arithmetic mean of all polygins. There is no transformation, and truncation is not needed as the indicator becomes naturally bound between 0 and 1".  -->


## 5. Uncertainties
### 5.1 General uncertainties
There are several sources of uncertainty that underlay the results that are presented in this report. Among them, we identified four major sources:

1. Prediction uncertainty linked to DT03, whereby 18% of the land take areas are false positive. These means that 18% of the land take  areas are not “real” land take. 

2. Discrepancy in temporal coverages between the datasets used (DT01, DT02, DT03). As already described previously, all the datasets used in our analysis do not have the same temporal coverage. This creates uncertainty when performing the spatial intersection as data are not representative from the same year. We took this source of uncertainty into account as described in section 5.2 Reported uncertainties. Note that we omitted DT01 as boundaries within municipalities are unlikely to change over time.

3.Spatial quality of the boundaries of DT03. It was not possible to assess the quality of the boundaries of the mapped specially important nature, neither the rate of error in the identification of the nature types. We are however conscious that they would impact the quality of the results if a habitat is mapped smaller or bigger than it is in reality, or if a habitat is not recorded as occurring due to misclassification in the habitat type. Due to the lack of potential proxies for these uncertainties, it was not possible to take them into account in our results.


4. Spatial uncertainty linked to DT03. As previoulsy explaines, this dataset doesn’t have a national coverage. This means that if a habitat is not recorded at a specific spatial location, it is either not occurring at that particular location, or occurring but the location has not been surveyed yet. Unfortunately, it was not possible to take into account this mapping uncertainty in our results as we could not find any appropriate proxy to represent it.


<!--# Describe the main reasons why you might not trust 100% the final value to represent the given ecosystem characteristic for the give ecosystem type. What aspects of this condition metric are most sensitive to subjective decitions? Is the spatial sampling sufficient? Does the choice of modeling framework affect the values? Is the metric relevant do describe ecosystem condition for the give ecosystem type?  -->


### 5.2 Reported uncertainties
For each municipality in each of the extent accounts we compiled, we calculated a qualitative indicator of uncertainty associated to the results. This indicator is based on the first two sources of uncertainty described in the previous section (i.e., prediction uncertainty of DT02 and temporal discrepancy between datasets) and has been designed as a traffic light system whereby:

-	Green: “Certain”
- Yellow: “Rather certain”
- Orange: “Rather uncertain”
- Red: “Uncertain”

Note that although we use “certain” in our Likert scale, no results can truly be qualified as “certain” with scientific robustness due to the numerous sources of uncertainty we listed previously. This was done to make the interpretation of our indicator more user-friendly.

The complete description of how this indicator has been compiled can be found in our NINA Report: Clappe, S., Barton D.N., Spielhofer, R. & Venter, Z. 2026.  Arealregnskap for friluftsliv - Hvor stort område gikk tapt på grunn av inngrep og hvilke andre naturverdier hadde de? NINA Rapport 2754. Norsk institutt for naturforskning. ISBN: 978-82-426-5592-9


<!-- Describe the nature of the errors you (hopefully) provide alongside the point estimate for the condition metric. It is especially important to clarify if the errors are inferential (referring to the uncertainty of the point estimate) or descriptive (referreing to the data itself). Here' an example text: "What we report as uncertainties in @tbl-finalTable are inferential and refer to the uncertainty around the point estimates. It does not directly describe variation in the data itself. 
However, the variation in the mean is defined by the spatial variation only, and quartiles are obtained by bootstrapping the set of localities within a region, and calculating the indicator for each of 9999 such samples".   -->



## 6. References

NRK, 2025a. NRK avslører europeisk naturtap: Norge topper verstingliste [online]. NRK. URL https://www.nrk.no/dokumentar/xl/europa-fra-gront-til-gratt_-nordmenn-bygger-ned-mest-natur_-viser-europeisk-kartlegging-1.17577504 (accessed 2.20.26).

Venter, Z., 2024a. Viser nedbygging av naturen ved hjelp av fugleperspektiv og kunstig intel-ligens [online]. Norsk institutt for naturforskning. URL https://www.nina.no/Om-NINA/Aktuelt/Nyheter/article/viser-nedbygging-av-naturen-ved-hjelp-av-fugleperspektiv-og-kunstig-intelligens (accessed 2.20.26).

Venter, Z., Mads, N.S., Ruben, S., Anne Linn, K.-E., Su Thet, M., 2024b. Map of built-up ex-pansion (“nedbygging”) over Norway 2017-2022. https://doi.org/10.5281/zenodo.10566644

Venter, Z.S., Czúcz, B., Stange, E., Nowell, M.S., Simensen, T., Immerzeel, B., Barton, D.N., 2024c. ‘Uncertainty audit’ for ecosystem accounting: Satellite-based ecosystem extent is biased without design-based area estimation and accuracy assessment. Ecosystem Services 66: 101599. https://doi.org/10.1016/j.ecoser.2024.101599

Venter, Z., 2025. Map of built-up expansion ("nedbygging") over Norway 2017-2024 with year of loss. Zenodo. https://doi.org/10.5281/zenodo.15130992



<!--# You can add references manually or use a citation manager and add intext citations as with crossreferencing and hyperlinks. See https://quarto.org/docs/authoring/footnotes-and-citations.html -->

## 7. Import and prepare datasets
In this section, we show how to import the data through API. We also prepare them  for analyses. We first set up the path that will be used during the project. Users can change it to match their own path.
```{r path_setup1}
#| eval: true
#| code-summary: "Path setup"

# Path to the project folder
project_path <- here::here()

# Path to the functions folder within the project folder
functions_path <- here::here(project_path, "src/functions")
```

<!--# All the data import and preparation should be done in this chapter, before the analysis chapter. As a minimum, add code for importing data, and describe or show the data structure. You may also explain certain variables, what they represent, and their distributions -->

### 7.1. Encroachment areas - DT01
#### 7.1.1 Download DT01 from Zenodo through an API
We created a function that users can use to retrieve the downloadable *content* link to the data. We could have used `zen4R` package but we decided to create our own function to read the data from Zenodo directly instead of downloading them locally before reading them.
```{r enc_link}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT01 API download"

# Call for the function that will read the data through an API
source(paste0(functions_path,"/api_zenodo.R"))

# Retrieve downloadable link to the data
encroach_link <- api_zenodo(zenodo_url = "https://zenodo.org/records/15130992")

```

#### 7.1.2 Read DT01
Now that we have the downloadable link, we can read the data in R and clean the R environment.
```{r enc_read}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT01 - read"

encroach_dt <- st_read(encroach_link)

```

#### 7.1.3 Clean and format the data for analyses
All columns are of relevance to our work but we need to clean the topological and geometrical errors if any.

```{r clean_enc}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "DT01 - clean"

# Check for topological and geometrical errors
which(st_is_valid(encroach_dt) == FALSE) %>% length()
which(st_is_empty(encroach_dt) == TRUE) %>% length()

# Clean R environment
rm(list = setdiff(ls(), 
                  c("encroach_dt",
                    "project_path",
                    "functions_path")))

```

The data are all clear and do not need further corrections.


### 7.2 Map of municipalities in Norway - DT02
#### 7.2.1 Download & read DT02 from GISCO through an API
For DT03, the API process is much simpler as a package has been developed to read data from GISCO via an API: `giscoR`. We just need to use one function to cache the dataset in a temporary directory and read it in R.
```{r lau_down_data}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT02 API download"

# Create a temporary directory to store the cache
temp_dir <- tempdir()

# Download the dataset
lau_no <- gisco_get_lau(year = "2024",
                        epsg = "3035",
                        cache = TRUE,
                        update_cache = TRUE,
                        cache_dir = temp_dir,
                        verbose = FALSE,
                        country = "NO",
                        gisco_id = NULL)

```

#### 7.2.2 Clean and format the data for analyses
We select the columns of relevance to our work and clean the topological and geometrical errors if any. Note that we also need to clean the `GISCO_ID`, which is the municipality number. 

```{r clean_lau}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "DT02 - general clean and format"

# We select the columns of interests
lau_no <- lau_no %>%
            select(GISCO_ID,
                   LAU_NAME) %>%
            mutate(GISCO_ID = sub("NO_", "", .$GISCO_ID)) %>%
            st_transform(., st_crs(encroach_dt)) # project to ESPG : 25832

# Check for topological and geometrical errors
which(st_is_valid(lau_no) == FALSE) %>% length()
which(st_is_empty(lau_no) == TRUE) %>% length()

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path")))
```


### 7.3 Naturtyper - verdsatte - DT03
#### 7.3.1 Download DT03 for Geonorge through an API
Downloading datasets from Geornorge using an API is a bit subtle and we created a function for it. The full step-by-step process is explained in the encroachment_recreation_areas.qmd document of this project (see README).
```{r import_natur_api1}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT03 API download"
# Call for the function that will read the data through an API
source(paste0(functions_path,"/api_geonorge.R"))

# Retrieve and download the data
natur_link <- api_geonorge(geonorge_url = "https://kartkatalog.geonorge.no/metadata/naturtyper-verdsatte/64cbb884-a19d-4356-a114-380cfe4a7314",
                              email_address = "john.doe@nina.no")

```

#### 7.3.2 Read DT03
Now that the dataset has been downloaded in a temporary file, we can read it in R and clean the R environment. Once the analyses are finished, we will remove the temporary file containing the data.
```{r read_nature}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT03 - read the data"

# Read data
naturtyper <- st_read(natur_link[grepl("Naturtyper", natur_link)])
```

#### 7.3.3 Clean and format the data for analyses
We first need to select the column of relevance for our work and clean the topological and geometrical errors if any. The two major issues we have with these data are the spatial and temporal coverages. Spatial coverage because the data do not cover the entire country of Norway. This means that it's not because we don't record one a nature type as occurring that it is indeed not occurring, it might just not have been sampled yet. Also, the nature types have been recorded only once for a wide temporal coverage: from 1905 to 2025. Given that our encroachment data cover 2017-2024, we decided to keep only the data that have been recorded in 2015 or after. We consider that a record in 2025 doesn't mean that the type was not present before. We also assume that nature types are relatively stable and have not change since 2015. We also removed all nature types whose year of report was not stated. This made us keep 72% of the original dataset, which is still very good for analyses.

In our report, we define "specially important nature" as the nature types within DT03 that are classified as "stor" and "svært stor" verdi. So we need to filter for them.

*Note: we restrict the coverage to 2025, which means that we filter out the polygons from 2026. This was done to be consistent with the figures in the NINA report we published but can of course be removed!*

```{r clean_natur}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "DT03 - clean and format"

# Select columns of relevance
natur_types <- naturtyper %>%
                mutate(year_natur_recorded = substr(.$registreringsdato, 1, 4)) %>%
                filter(verdikategori %in% c("Svært stor verdi",
                                            "Stor verdi")) %>%
                select(naturtype,
                       year_natur_recorded,
                       kommunenummer)

# Check for topological and geometrical errors
which(st_is_valid(natur_types) == FALSE) %>% length()
which(st_is_empty(natur_types) == TRUE) %>% length()

# Correct invalid geometries
natur_cln <- st_make_valid(natur_types)

# Check correction worked
which(st_is_valid(natur_cln) == FALSE) %>% length()
which(st_is_empty(natur_cln) == TRUE) %>% length()

# Clean typos in years
natur_cln$year_natur_recorded[natur_cln$year_natur_recorded == "2102"] <- "2012"
natur_cln$year_natur_recorded[natur_cln$year_natur_recorded == "2913"] <- "2013"

# Remove nature types that have been recorded before 2015
natur_cln <- natur_cln %>%
              filter(year_natur_recorded > 2009) %>%
              filter(year_natur_recorded < 2026) %>%
              filter(!is.na(year_natur_recorded))
  

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_cln",
                    "project_path",
                    "functions_path")))

```

## 8. Spatial units
All of our accounts have been compiled for Norway (Ecosystem Accounting Area) and broken down at municipal and county levels. As explained in section 3.1, the "ecosystem" of interest in ENCACC_001 is "land take". This means that our Basic Spatial Units are "land take" polygons in ENCACC_001, but "specially important nature" polygons in ENCACC_002. The ecosystem Asset is "land take" for ENCACC_001 and "specially important nature types" for ENCACC_002.

<!--# Describe the spatial units that you rely on in your analyses. Highlight the spatial units (the resolution) that the indicator values should be interpreted at. Potential spatial delineation data (maksing layers etc.) should be introduced under chapert 7. We recommend using the SEEA EA terminology of Basic Spatial Units (BSU), Ecosystem Asses (EA) and Ecosystem Accounting Area (EAA). -->

## 9. Analyses
### 9.1 ENCACC_001
To compile ENCACC_001, we need to perform a spatial intersections between DT01 and DT02. Note that we re-calculate the areas of the encroachment polygons after the intersection has been performed.

*Note: some municipalities had no encroachments detected in certain years. In these cases, the year where 0 encroachment was detected did not not appear. We thus had to add it manually for all municipalities. In these cases, we reported the uncertainty of the year of encroachment.*

```{r encrec001234_analyses}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "ENCACC_001 - intersection"

# Check the CRS of both DT01 and DT02
st_crs(lau_no) == st_crs(encroach_dt)

# Perform spatial intersection
komenc_raw <- st_intersection(lau_no, encroach_dt) %>%
                mutate(area_inter_m2 = unclass(st_area(.)), .after = area)

# Add missing years for accounting. Set the area as 0 for the missing year.
komenc <- komenc_raw %>%
            st_drop_geometry() %>%
            complete(nesting(GISCO_ID, LAU_NAME), year = 2018:2024, fill = list(id = "NA",
                                                                                area = 0,
                                                                                area_inter_m2 = 0))

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_cln",
                    "project_path",
                    "functions_path",
                    "komenc_raw",
                    "komenc")))

```
### 9.2 ENCACC_002
For ENCACC_002, we need a spatial intersection between DT03 and `komenc` as we are interested in the nature types that have been lost due to encroachment within municipalities. Here we have to note though that the only 4,2% of the specially important nature occur in the encroachment of municipalities.

```{r encrec0056_analyses}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "ENCACC_002 - intersection"

# Check the CRS of both encrec and DT04
st_crs(komenc_raw) == st_crs(natur_cln)

# Re-project natur_cln into ESPG : 25832
natur_32 <- st_transform(natur_cln, crs = st_crs(komenc_raw))

# Check CRS of both datasets are the same now
st_crs(natur_32) == st_crs(komenc_raw)

# Perform spatial intersection
encnatur_raw <- komenc_raw %>%
                  select(GISCO_ID,
                         LAU_NAME,
                         id,
                         year,
                         area_inter_m2) %>%
                  rename(area_enc = area_inter_m2,
                         year_enc = year) %>%
                  st_intersection(., natur_32) %>%
                  mutate(area_naturtyper_m2 = unclass(st_area(.)))

# Add missing years for accounting reporting
encnatur <- encnatur_raw %>%
            st_drop_geometry() %>%
            select(!c(kommunenummer, area_enc, id)) %>%
            complete(nesting(GISCO_ID, LAU_NAME, naturtype), 
                     year_enc = 2018:2024,
                     fill = list(area = 0,
                                 area_naturtyper_m2 = 0)) %>%
            mutate(year_natur_recorded = coalesce(as.integer(year_natur_recorded), year_enc))

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "natur_32",
                    "komenc",
                    "encnatur",
                    "encnatur_raw",
                    "komenc_raw")))

```
### 9.3 Uncertainty analyses
Our traffic-light uncertainty required to steps:

1. Attribute a traffic-light uncertainty at the polygon level. For this we used our lookup table developed in our NINA Rapport 2754; 

2. Derive a municipal / county traffic-light uncertainty based on the frequency of the traffic-light uncertainty at polygon-level (step 1 above). We basically created a frequency table of all the traffic-light color uncertainties from the polygons and derived the municipal / county traffic-light color uncertainty based the most predominant polygons' traffic-light colors. See NINA Rapport 2754 for more information.

First we attributed a colored-uncertainty to all sources of uncertainty.
```{r encacc_uncertainty_1}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "Step 1 - attribute colored-uncertainty to all polygons"

# Load necessary functions
source(paste0(functions_path,"/add_uncertainty_individual.R"))

# Uncertainty ENCACCC_001
unc_001_idv <- komenc %>%
                st_drop_geometry() %>%
                select(GISCO_ID, year) %>%
                mutate(uncertainty_source1 = apply(., 1, function(x) {add_uncertainty_individual(data = x, source1 = TRUE, yrsource1 = "year")}))


# Uncertainty ENCACCC_002
unc_002_idv <- encnatur %>%
                st_drop_geometry() %>%
                select(GISCO_ID, year_enc, year_natur_recorded, naturtype) %>%
                mutate(uncertainty_source2 = apply(., 1, function(x) {add_uncertainty_individual(data = x, source2 = TRUE, yrsource2 = c("year_enc", "year_natur_recorded"))}))


# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "natur_32",
                    "komenc",
                    "encnatur",
                    "unc_001_idv",
                    "unc_002_idv")))

```

The second step is to attribute ONE "global" color-based uncertainty to each municipality for each accounting year and change between 2018 and 2024. For this we need to create a vector of all the color-based uncertainties per municipality per year. After that we look at the frequency of each of the colors and we take the two most predominant. In case of ties, we take the three most predominant. In case the four color-based uncertainty were all equal in their frequency, we decided to classify the "global" uncertainty as "Orange". To see the "global" color-based uncertainty we assign with each combination of two or three predominant colors of uncertainty, you can refer to our NINA Rapport 2754. For this we first need to create a vector where we concatenate all the individual color-based uncertainties, and then create a function to produce the "global" uncertainty. As we haven't calculated all the metrics yet: cumulative areas and change area, we only summarise and concatenate all the color uncertainties for each municipality for each year. When we will calculate our metrics, we will then compile the final uncertainty going with each metric. We wrote a function to concatenate all the colored-uncertainty into one column per municipality per accounting year. *Note: the "accounting year" is the year of encroachment as we are compiling encroachment extent accounts. Only the area of accounting is changing among the accounts*.


```{r encacc_uncertainty_2}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "Pre-Step 2 - concatenate colored uncertainty from Step 1"

# Call function to concatenate
source(paste0(functions_path,"/concatenate_individual_uncertainty.R"))

# Concatenate individual colored-uncertainties
unc_001_sum <- concatenate_fn(unc_001_idv,
                              col_column_names = "uncertainty_source1", 
                              kom_num = "GISCO_ID",
                              accounting_year = "year")

unc_002_sum <- concatenate_fn(unc_002_idv,
                              col_column_names = "uncertainty_source2",
                              kom_num = "GISCO_ID",
                              habitat_col = "naturtype")

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "natur_32",
                    "komenc",
                    "encnatur",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum")))

```

<!--# 

Use this header for documenting the analyses. Put code in separate code chunks, and annotate the code in between using normal text (i.e. between the chunks, and try to avoid too many hashed out comments inside the code chunks). Add subheaders as needed. 

Code folding is activated, meaning the code will be hidden by default in the html (one can click to expand it).

Caching is also activated (from the top YAML), meaning that rendering to html will be quicker the second time you do it. This will create a folder inside you project folder (called INDICATORID_cache). Sometimes caching created problems because some operations are not rerun when they should be rerun. Try deleting the cash folder and try again.

-->

## 10. Results - Thematic Extent Accounts: land take in municipalities
Now we can built the accounting tables and maps based on the spatial intersection we performed in the previous section. We chose to build the thematic accounts over the 2018-2024 period. Maps will be created only for the closing (2024) year of the total accounting period 2018-2024.

### 10.1 Accounting tables
#### 10.1.1 ENCACC_001: Extent accounts of land take in municipalities
Here we present the extent accounts table for land take areas between 2018 and 2024 within each municipality and county in Norway. These accounts present both the yearly area of land take and the cumulative area of land take over the years. We also present a total for Norway. We added the uncertainty as color-based indicators presented in 5.2 Reported uncertainties. We first present the table per municipalities and then per county.

*Note: the function calculating the change compiles it as area_closing_year - area_opening_year. This is because we expect an increase in land take so we want to show how we gain in built-in areas.*

*Note: We consider the uncertainty of the change in area between the opening and closing accounting years, the same as the cumulative uncertainty for the closing accounting year. This is because when we calculate a change we refer to all the accounting period as the cumulative area for the last accounting year is calculated based on all the accounting years. In our code, add_change_encroachment_area() automatically duplicate the uncertainty of the last accounting year for the change in area.*

```{r res_acc_tables1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "ENCACC_001 - Create accounting table per municipalities"

## Call function to add change in areas 2018-2024
source(paste0(functions_path, "/add_change_encroachment_area.R"))

## Call function to add fylke
source(paste0(functions_path, "/add_fylke.R"))

## Call function to compile individual uncertainty
source(paste0(functions_path,"/add_uncertainty_individual.R"))

## Call function to compile global uncertainty
source(paste0(functions_path, "/add_uncertainty_global.R"))

## ENCACC_001 accounting table
# Accounts for all municipalities
encacc_001_kom <- komenc %>%
                    st_drop_geometry() %>%
                    select(GISCO_ID,
                           year,
                           LAU_NAME,
                           area_inter_m2) %>%
                    mutate(fylke = add_fylke(data = ., kommune_nummer = "GISCO_ID")[[1]], .before = GISCO_ID) %>%
                    group_by(fylke, GISCO_ID, LAU_NAME, year) %>%
                    summarise(area_dekar_yearly = round2(sum(area_inter_m2)/1000, 2)) %>%
                    mutate(area_dekar_cumulative = round2(cumsum(area_dekar_yearly), 2)) %>%
                    mutate(indicator_id = "ENCACC_001", .before = "fylke") 

# Add uncertainties to encacc_001_kom
encacc_001_kom_un <- encacc_001_kom %>%
                      left_join(., unc_001_sum, by = c("GISCO_ID", "year")) %>%
                      add_uncertainty_global(., 
                                             uncertainty_col_name = "unc_concat_full", 
                                             kom_num = "GISCO_ID", 
                                             accounting_year = "year") %>%
                      select(!unc_concat_full)


# Total across all municipalities
encacc_001_tot <- encacc_001_kom %>%
                    group_by(year) %>%
                    summarise(area_dekar_yearly = sum(area_dekar_yearly)) %>%
                    mutate(area_dekar_cumulative = round2(cumsum(area_dekar_yearly), 2)) %>%
                    mutate(indicator_id = "ENCACC_001",
                           fylke = "00",
                           GISCO_ID = "0000",
                           LAU_NAME = "Norge", .before = year)

# Add uncertainty to encacc_001_tot
encacc_001_tot_un <-encacc_001_tot  %>%
                      mutate(uncertainty_source1 = apply(., 
                                                         1, 
                                                         function(x){add_uncertainty_individual(data = x, source1 = TRUE, yrsource1 = "year")})) %>%
                      
                      add_uncertainty_global(., 
                                             uncertainty_col_name = "uncertainty_source1", 
                                             kom_num = "GISCO_ID", 
                                             accounting_year = "year") %>%
                      select(!uncertainty_source1)

# Add change to all tables
encacc_001_kom_un$year <- as.character(encacc_001_kom_un$year)

encacc_001_kom_un <- encacc_001_kom_un %>%
                      group_split(GISCO_ID) %>% 
                      map(~ add_change_encroachment_area(., 
                                                         year_column = "year",
                                                         area1 = "area_dekar_cumulative")) %>% 
                      compact() %>%
                      bind_rows()

encacc_001_tot_un$year <- as.character(encacc_001_tot_un$year)

encacc_001_tot_un <- encacc_001_tot_un %>%
                      add_change_encroachment_area(., 
                                                   year_column = "year",
                                                   area1 = "area_dekar_cumulative") %>% 
                      compact() %>%
                      bind_rows()

# Correct uncertainty for global_yearly_uncertainty when accounting year is "Change 2018-2024"
encacc_001_kom_un[which(encacc_001_kom_un$year == "Change 2018-2024"), "global_yearly_uncertainty"] <- "NA"
encacc_001_tot_un[which(encacc_001_tot_un$year == "Change 2018-2024"), "global_yearly_uncertainty"] <- "NA"

# Translate "Change 2018-2024" in Norwegian
encacc_001_kom_un[which(encacc_001_kom_un$year == "Change 2018-2024"), "year"] <- "Endring 2018-2024"
encacc_001_tot_un[which(encacc_001_tot_un$year == "Change 2018-2024"), "year"] <- "Endring 2018-2024"

# Final accounting table               
encacc_001_acc <- bind_rows(encacc_001_kom_un,
                            encacc_001_tot_un)

encacc_001_acc$fylke[which(encacc_001_acc$fylke == "00")] <- "Norge"
encacc_001_acc$GISCO_ID[which(encacc_001_acc$GISCO_ID == "0000")] <- "Norge"

colnames(encacc_001_acc) <- c("Indikatorens navn",
                              "Fylke nummer",
                              "Kommune nummer",
                              "Kommune navn",
                              "Regnskapsår",
                              "Årlig areal i dekar",
                              "Kumulativ areal i dekar",
                              "Usikkerhet for årlig areal",
                              "Usikkerhet for kumulativ areal")


# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "add_change_encroachment_area",
                    "add_fylke",
                    "add_uncertainty_individual",
                    "add_uncertainty_global",
                    "encacc_001_acc",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum")))


# Display the table dynamically
datatable(encacc_001_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list( dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                          columnDefs = list(list(targets = c(which(colnames(encacc_001_acc) == "Usikkerhet for årlig areal"),
                                                             which(colnames(encacc_001_acc) == "Usikkerhet for kumulativ areal")),
                                                 visible = FALSE))),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Tabell 1 – Tematiske arealregnskap: areal av nedbygging i kommunner i perioden 2018–2024."))) %>%
    formatCurrency(
    columns = c("Årlig areal i dekar", "Kumulativ areal i dekar"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",") %>%
  ## Example: color Årlig areal i dekar by var1
  formatStyle(
    "Årlig areal i dekar",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red", "NA"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A", "")
    ),
    valueColumns = "Usikkerhet for årlig areal"           # column used for the intervals
  ) %>%
  ## Example: color Kumulativ areal i dekar by var2
  formatStyle(
    "Kumulativ areal i dekar",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A")
    ),
    valueColumns = "Usikkerhet for kumulativ areal"           # column used for the intervals
  )

```


```{r res_acc_tables2}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "ENCACC_001 - Create accounting table per county"

# Call function to concatenate
source(paste0(functions_path,"/concatenate_individual_uncertainty.R"))

# Create the accounting table based on encacc_001_acc
encacc_001_fy <- encacc_001_acc %>%
                  filter(!Regnskapsår == "Endring 2018-2024") %>%
                  filter(!`Kommune navn` == "Norge") %>%
                  group_by(`Indikatorens navn`, `Fylke nummer`, Regnskapsår) %>%
                  summarise(yearly_area = sum(`Årlig areal i dekar`)) %>%
                  mutate(cumulative_area = cumsum(yearly_area))

# Add uncertainties to encacc_001_fy
encacc_001_fy_un <-  unc_001_idv %>%
                      mutate(fylke = add_fylke(data = ., kommune_nummer = "GISCO_ID")[[1]], .before = GISCO_ID) %>%
                      select(!GISCO_ID) %>%
                      concatenate_fn(.,
                                     col_column_names = c("uncertainty_source1"),
                                     kom_num = "fylke",
                                     accounting_year = "year") %>%
                      add_uncertainty_global(., 
                                                      kom_num = "fylke",
                                                      uncertainty_col_name = "unc_concat_full",
                                                      accounting_year = "year") %>%
                      select(!unc_concat_full) %>%
                      mutate(year = as.character(year)) %>%
                      left_join(encacc_001_fy, ., join_by(`Fylke nummer` == "fylke", "Regnskapsår" == "year")) %>%
                      ungroup() %>%
                      mutate(fylke_name = add_fylke(data = ., fylke_nummer = "Fylke nummer"), .after = `Fylke nummer`)


# Add change to encacc_001_fy_un
encacc_001_fy_unc <- encacc_001_fy_un %>%
                      
                      group_split(`Fylke nummer`) %>% 
                      map(~ add_change_encroachment_area(., 
                                                         year_column = "Regnskapsår",
                                                         area1 = "cumulative_area")) %>% 
                      compact() %>%
                      bind_rows()

# Correct uncertainty for global_yearly_uncertainty when accounting year is "Change 2018-2024"
encacc_001_fy_unc[which(encacc_001_fy_unc$Regnskapsår == "Change 2018-2024"), "global_yearly_uncertainty"] <- "NA"

# Translate "Change 2018-2024" in Norwegian
encacc_001_fy_unc[which(encacc_001_fy_unc$Regnskapsår == "Change 2018-2024"), "Regnskapsår"] <- "Endring 2018-2024"

# Modify colnames
colnames(encacc_001_fy_unc) <- c("Indikatorens navn",
                                 "Fylke nummer",
                                 "Fylke navn",
                                 "Regnskapsår",
                                 "Årlig areal i dekar",
                                 "Kumulativ areal i dekar",
                                 "Usikkerhet for årlig areal",
                                 "Usikkerhet for kumulativ areal")
                                        



# Add total across fylke
encacc_001_fy_tot <- encacc_001_acc %>%
                        filter(`Kommune navn` == "Norge") %>% 
                        rename("Fylke navn" = "Kommune navn") %>%
                        select(!`Kommune nummer`)
                        
  
encacc_001_fy_acc <- bind_rows(encacc_001_fy_unc,
                               encacc_001_fy_tot)

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "add_change_encroachment_area",
                    "add_fylke",
                    "add_uncertainty_individual",
                    "add_uncertainty_global",
                    "concatenate_fn",
                    "encacc_001_acc",
                    "encacc_001_fy_acc",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum")))


# Display the table dynamically
datatable(encacc_001_fy_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list( dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                          columnDefs = list(list(targets = c(which(colnames(encacc_001_fy_acc) == "Usikkerhet for årlig areal"),
                                                             which(colnames(encacc_001_fy_acc) == "Usikkerhet for kumulativ areal")),
                                                 visible = FALSE))),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Tabell 2 – Tematiske arealregnskap: areal av nedbygging i fylker i perioden 2018–2024."))) %>%
    formatCurrency(
    columns = c("Årlig areal i dekar", "Kumulativ areal i dekar"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",") %>%
  ## Example: color Årlig areal i dekar by var1
  formatStyle(
    "Årlig areal i dekar",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red", "NA"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A", "")
    ),
    valueColumns = "Usikkerhet for årlig areal"           # column used for the intervals
  ) %>%
  ## Example: color Kumulativ areal i dekar by var2
  formatStyle(
    "Kumulativ areal i dekar",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A")
    ),
    valueColumns = "Usikkerhet for kumulativ areal"           # column used for the intervals
  )

```

#### 10.1.2 ENCACC_002: Extent accounts of all specially important nature lost due to land take in municipalities
Here we present the extent accounts table for specially important nature lost to land take between 2018 and 2024 within each municipality in Norway. These accounts present both the yearly area of land take and the cumulative area of land take over the years. We also present a total for Norway. As previously, we first present the results per municipality sand then aggregated at county's level.

As the loss of the nature types did not happen homogeneously within municipalities, these accounts only present both the total area lost by nature types over the 2018-2024 period. This is to avoid confusion and present the bigger picture, instead of having different years for different municipalities and nature types. We also present a total across all municipalities in Norway which mapped their recreation areas.


```{r res_acc_tables3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "ENCACC_002 - Create accounting table per municpalities"

## ENCACC_002 accounting table
# Accounts for all municipalities
encacc_002_kom <- encnatur %>%
                    st_drop_geometry() %>%
                    select(GISCO_ID,
                           LAU_NAME,
                           naturtype,
                           year_enc,
                           area_naturtyper_m2) %>%
                    mutate(fylke = add_fylke(data = ., kommune_nummer = "GISCO_ID")[[1]], .before = GISCO_ID) %>%
                    group_by(fylke, GISCO_ID, LAU_NAME, naturtype) %>%
                    summarise(total_loss_dekar = round2(sum(area_naturtyper_m2)/1000, 2)) %>%
                    mutate(indicator_id = "ENCACC_002", .before = "fylke") %>%
                    mutate(accounting_period = "2018-2024", .before = "total_loss_dekar")

# Add uncertainties to encacc_002_kom
encacc_002_kom_un <- encacc_002_kom %>%
                      left_join(., unc_002_sum, by = c("GISCO_ID", "naturtype"))  %>%
                      add_uncertainty_global(., 
                                             uncertainty_col_name = "unc_concat_full", 
                                             kom_num = "GISCO_ID", 
                                             habitat_accounting = "naturtype") %>%
                      select(!unc_concat_full) %>%
                      rename(global_period_uncertainty = global_habitat_uncertainty)

# National total across all municipalities
encacc_002_tot <- encacc_002_kom %>%
                    group_by(indicator_id, accounting_period) %>%
                    summarise(total_loss_dekar = sum(total_loss_dekar)) %>%
                    mutate(fylke = "00",
                           GISCO_ID = "0000",
                           LAU_NAME = "Norge", .before = accounting_period) 

# Add uncertainties to encacc_002_tot
encacc_002_tot_un <-  unc_002_idv %>%
                        mutate(national = "0000",
                               accounting_period = "2018-2024") %>%
                        concatenate_fn(.,
                              col_column_names = "uncertainty_source2",
                              kom_num = "national",
                              accounting_year = "accounting_period") %>%
                        add_uncertainty_global(., 
                                               kom_num = "national",
                                               uncertainty_col_name = "unc_concat_full",
                                               accounting_year = "accounting_period") %>%
                      ungroup() %>%

                      select(!c(unc_concat_full, 
                                global_cumulative_uncertainty,
                                national,
                                accounting_period)) %>%
  
                      cbind(encacc_002_tot, .) %>%
                      mutate(naturtype = "Alle", .after = LAU_NAME) %>%
                      rename(global_period_uncertainty = global_yearly_uncertainty)


# Final accounting table               
encacc_002_acc <- bind_rows(encacc_002_kom_un,
                            encacc_002_tot_un)

encacc_002_acc$fylke[which(encacc_002_acc$fylke == "00")] <- "Norge"
encacc_002_acc$GISCO_ID[which(encacc_002_acc$GISCO_ID == "0000")] <- "Norge"


colnames(encacc_002_acc) <- c("Indikatorens navn",
                              "Fylke nummer",
                              "Kommune nummer",
                              "Kommune navn",
                              "Verdsatte naturtype",
                              "Regnskapsperiode",
                              "Areal verdsatte naturtyper gått tapt (dekar)",
                              "Usikkerhet for areal i regnskapsperiode")


# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "add_change_encroachment_area",
                    "add_fylke",
                    "add_uncertainty_individual",
                    "add_uncertainty_global",
                    "concatenate_fn",
                    "encacc_001_acc",
                    "encacc_001_fy_acc",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum",
                    "encacc_002_acc")))


# Display the table dynamically
datatable(encacc_002_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list( dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                          columnDefs = list(list(targets = c(which(colnames(encacc_002_acc) == "Usikkerhet for areal i regnskapsperiode")),
                                                 visible = FALSE))),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Tabel 3 - Tematiske arealregnskap: areal av verdsatte natur tapt på grunn av nedbygging i kommuner i perioden 2018–2024."))) %>%
    formatCurrency(
    columns = c("Areal verdsatte naturtyper gått tapt (dekar)"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",") %>%
  ## Example: color Årlig areal i dekar by var1
  formatStyle(
    "Areal verdsatte naturtyper gått tapt (dekar)",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red", "NA"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A", "")
    ),
    valueColumns = "Usikkerhet for areal i regnskapsperiode"           # column used for the intervals
  )

```


```{r res_acc_tables4}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "ENCACC_002 - Create accounting table per county"

# Create the accounting table based on encacc_001_acc
encacc_002_fy <- encacc_002_acc %>%
                  filter(!`Kommune navn` == "Norge") %>%
                  group_by(`Indikatorens navn`, `Fylke nummer`, `Verdsatte naturtype`, Regnskapsperiode) %>%
                  summarise(total_area = sum(`Areal verdsatte naturtyper gått tapt (dekar)`))

# Add uncertainties to encacc_001_fy
encacc_002_fy_un <-  unc_002_idv %>%
                      mutate(fylke = add_fylke(data = ., kommune_nummer = "GISCO_ID")[[1]]) %>%
                      concatenate_fn(.,
                                     col_column_names = "uncertainty_source2",
                                     kom_num = "fylke",
                                     habitat_col = "naturtype") %>%
                      add_uncertainty_global(., 
                                                      kom_num = "fylke",
                                                      uncertainty_col_name = "unc_concat_full",
                                                      habitat_accounting = "naturtype") %>%
                      select(!unc_concat_full) %>%
                      left_join(encacc_002_fy, ., join_by(`Fylke nummer` == "fylke", "Verdsatte naturtype" == "naturtype"))  %>%
                      ungroup() %>%
                      mutate(fylke_name = add_fylke(data = ., fylke_nummer = "Fylke nummer"), .after = `Fylke nummer`)

# Modify colnames
colnames(encacc_002_fy_un) <- c("Indikatorens navn",
                                 "Fylke nummer",
                                 "Fylke navn",
                                 "Verdsatte naturtype",
                                 "Regnskapsperiode",
                                 "Areal verdsatte naturtyper gått tapt (dekar)",
                                 "Usikkerhet for areal i regnskapsperiode")


# Add total across fylke
encacc_002_fy_tot <- encacc_002_acc %>%
                        filter(`Kommune navn` == "Norge") %>% 
                        rename("Fylke navn" = "Kommune navn") %>%
                        select(!`Kommune nummer`)
  
encacc_002_fy_acc <- bind_rows(encacc_002_fy_un,
                               encacc_002_fy_tot)

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "add_change_encroachment_area",
                    "add_fylke",
                    "add_uncertainty_individual",
                    "add_uncertainty_global",
                    "concatenate_fn",
                    "encacc_001_acc",
                    "encacc_001_fy_acc",
                    "encacc_002_acc",
                    "encacc_002_fy_acc",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum")))


# Display the table dynamically
datatable(encacc_002_fy_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list( dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                          columnDefs = list(list(targets = c(which(colnames(encacc_002_fy_acc) == "Usikkerhet for areal i regnskapsperiode")),
                                                 visible = FALSE))),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Tabel 4 - Tematiske arealregnskap: areal av verdsatte natur tapt på grunn av nedbygging i fylker i perioden 2018–2024."))) %>%
    formatCurrency(
    columns = c("Areal verdsatte naturtyper gått tapt (dekar)"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",") %>%
      
  formatStyle(
    "Areal verdsatte naturtyper gått tapt (dekar)",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red", "NA"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A", "")
    ),
    valueColumns = "Usikkerhet for areal i regnskapsperiode" )

```


### 10.2 Maps
#### 10.2.1 ENCACC_001: Map of the total land take area in municipalities over 2018-2024
We first prepare the dataset for mapping then map it both using an interactive and static map. Due to memory issues in the html, we only display the static map.
```{r map_enc001_1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Prepare dataset for mapping"

## Import LAU dataset to then join the name of the municipalities
kom_navn <- lau_no %>%
              select(GISCO_ID)

## LAU dataset grouped by fylke
fylke_geo <- lau_no %>%
              select(GISCO_ID) %>%
              mutate(fylke = add_fylke(data = ., kommune_nummer = "GISCO_ID")[[1]]) %>%
              group_by(fylke) %>%
              summarise()

## Prepare dataset
# Municipalities
encacc_001_map_mun <- encacc_001_acc %>%
                        filter(!c(`Kommune navn` == "Norge" | Regnskapsår == "Endring 2018-2024")) %>%
                        filter(Regnskapsår == 2024) %>%
                        left_join(., kom_navn, join_by( `Kommune nummer` == "GISCO_ID")) %>%
                        select("Fylke nummer",
                               "Kommune navn",
                               "Kumulativ areal i dekar",
                               "Usikkerhet for kumulativ areal",
                               "geometry") %>%
                        rename("Kumulativ areal av nedbygging i dekar" = "Kumulativ areal i dekar") %>%
                        st_as_sf()

# Fylke
encacc_001_map_fy <- encacc_001_fy_acc %>%
                        filter(!c(`Fylke navn` == "Norge" | Regnskapsår == "Endring 2018-2024")) %>%
                        filter(Regnskapsår == 2024) %>%
                        left_join(., fylke_geo, join_by( `Fylke nummer` == "fylke")) %>%
                        select("Fylke nummer",
                               "Fylke navn",
                               "Kumulativ areal i dekar",
                               "Usikkerhet for kumulativ areal",
                               "geometry") %>%
                        rename("Kumulativ areal av nedbygging i dekar" = "Kumulativ areal i dekar") %>%
                        st_as_sf()


# Replace labels of uncertainty
encacc_001_map_mun$`Usikkerhet for kumulativ areal`[which(encacc_001_map_mun$`Usikkerhet for kumulativ areal` == "Green")]<- "Sikker"

encacc_001_map_mun$`Usikkerhet for kumulativ areal`[which(encacc_001_map_mun$`Usikkerhet for kumulativ areal` == "Yellow")]<- "Ganske sikker"

encacc_001_map_mun$`Usikkerhet for kumulativ areal`[which(encacc_001_map_mun$`Usikkerhet for kumulativ areal` == "Orange")]<- "Ganske usikker"

encacc_001_map_mun$`Usikkerhet for kumulativ areal`[which(encacc_001_map_mun$`Usikkerhet for kumulativ areal` == "Red")]<- "Usikker"


encacc_001_map_fy$`Usikkerhet for kumulativ areal`[which(encacc_001_map_fy$`Usikkerhet for kumulativ areal` == "Green")]<- "Sikker"

encacc_001_map_fy$`Usikkerhet for kumulativ areal`[which(encacc_001_map_fy$`Usikkerhet for kumulativ areal` == "Yellow")]<- "Ganske sikker"

encacc_001_map_fy$`Usikkerhet for kumulativ areal`[which(encacc_001_map_fy$`Usikkerhet for kumulativ areal` == "Orange")]<- "Ganske usikker"

encacc_001_map_fy$`Usikkerhet for kumulativ areal`[which(encacc_001_map_fy$`Usikkerhet for kumulativ areal` == "Red")]<- "Usikker"

```

```{r map_enc001_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Interactive map of total encroachment area in municipalities over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/interactive_map.R"))

# Run function
interactive_map(data1 = encacc_001_map_mun, 
                data2 = encacc_001_map_mun, 
                fill_variable1 = "Kumulativ areal av nedbygging i dekar", 
                fill_variable2 = "Usikkerhet for kumulativ areal", 
                legend_title1 = "Totalt areal av nedbygging (dekar)", 
                legend_title2 = "Usikkerhet",
                style_type_map1 = "fixed",
                interval_breaks_map1 = c(0, 500, 1000, 1500, 2000, 2500),
                color_values_map2 = c("Sikker" = "#94F4B6",
                                      "Ganske sikker" = "#FFF085", 
                                      "Ganske usikker" = "#FFC68A",
                                      "Usikker" = "#FF8A8A"),
                map_title1 = "Totalt areal av nedbygging (dekar) i kommuner over 2018-2024", 
                map_title2 = "Usikkerhet knyttet til det totale arealet av nedbygging over 2018-2024")

```

```{r map_enc001_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of total encroachment area in municipalities over 2018-2024"

## Static maps
# Call for the function to do static mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = encacc_001_map_mun, 
           data2 = encacc_001_map_mun, 
           fill_variable1 = "Kumulativ areal av nedbygging i dekar", 
           fill_variable2 = "Usikkerhet for kumulativ areal", 
           legend_title1 = "Totalt areal av nedbygging (dekar)", 
           legend_title2 = "Usikkerhet",
           style_type_map1 = "fixed",
           interval_breaks_map1 = c(0, 500, 1000, 1500, 2000, 5700),
           color_values_map2 = c("Sikker" = "#94F4B6",
                                 "Ganske sikker" = "#FFF085", 
                                 "Ganske usikker" = "#FFC68A",
                                 "Usikker" = "#FF8A8A"),
           map_title1 = "Totalt areal av nedbygging (dekar) i kommuner over 2018-2024", 
           map_title2 = "Usikkerhet til det totale arealet av nedbygging over 2018-2024")

```


```{r map_enc001_4}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of total encroachment area in counties over 2018-2024"

## Static maps
# Call for the function to do static mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = encacc_001_map_fy, 
           data2 = encacc_001_map_fy, 
           fill_variable1 = "Kumulativ areal av nedbygging i dekar", 
           fill_variable2 = "Usikkerhet for kumulativ areal", 
           legend_title1 = "Totalt areal av nedbygging (dekar)", 
           legend_title2 = "Usikkerhet",
           style_type_map1 = "fixed",
           interval_breaks_map1 = c(0, 10000, 20000, 30000, 40000, 54000),
           color_values_map2 = c("Sikker" = "#94F4B6",
                                 "Ganske sikker" = "#FFF085", 
                                 "Ganske usikker" = "#FFC68A",
                                 "Usikker" = "#FF8A8A"),
           map_title1 = "Totalt areal av nedbygging (dekar) i fylker over 2018-2024", 
           map_title2 = "Usikkerhet til det totale arealet av nedbygging over 2018-2024")


```

```{r map_enc001_5}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Clean R environment"

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "add_change_encroachment_area",
                    "add_fylke",
                    "add_uncertainty_individual",
                    "add_uncertainty_global",
                    "concatenate_fn",
                    "encacc_001_acc",
                    "encacc_001_fy_acc",
                    "encacc_002_acc",
                    "encacc_002_fy_acc",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum",
                    "fylke_geo",
                    "encacc_001_map_mun",
                    "encacc_001_map_fy")))
```

#### 10.2.2 ENCACC_001: Share of the total land take area in municipalities over 2018-2024
We first prepare the dataset for mapping then map it both using an interactive and static map. Due to memory issues in the html, we only display the static map. 

```{r map_enc001_perc_1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Prepare data for mapping"

## LAU dataset with municipal areas
kom_navn <- lau_no %>%
              select(GISCO_ID) %>%
              mutate(area_kom = round2(unclass(st_area(.))/1000, 2))


## Calculate total area per fylke - attention!! only take the area of the municipalities that have encroachment...
fylke_total_area <- encacc_001_acc %>%
                      filter(!is.na(`Fylke nummer`)) %>%
                      left_join(., kom_navn, join_by(`Kommune nummer` == "GISCO_ID")) %>%
                      st_drop_geometry() %>%
                      select(`Fylke nummer`, area_kom) %>%
                      group_by(`Fylke nummer`) %>%
                      summarise(area_fylke = sum(area_kom)) %>%
                      left_join(., fylke_geo, join_by("Fylke nummer" == "fylke"))
  



## Prepare dataset
encacc_001_perc_map_mun <- encacc_001_acc %>%
                              filter(!c(`Kommune navn` == "Norge" | Regnskapsår == "Endring 2018-2024")) %>%
                              filter(Regnskapsår == 2024) %>%
                              left_join(., kom_navn, join_by( `Kommune nummer` == "GISCO_ID")) %>%
                              select("Fylke nummer",
                                     "Kommune navn",
                                     "area_kom",
                                     "Kumulativ areal i dekar",
                                     "Usikkerhet for kumulativ areal",
                                     "geometry") %>%
                              mutate("Prosentandel (%) av nedbygging" = round2(`Kumulativ areal i dekar`/area_kom * 100, 2), .before = "Usikkerhet for kumulativ areal") %>%
                              select(!c("Kumulativ areal i dekar", area_kom)) %>%
                              st_as_sf()

encacc_001_perc_map_fy <- encacc_001_fy_acc %>%
                              filter(!c(`Fylke navn` == "Norge" | Regnskapsår == "Endring 2018-2024")) %>%
                              filter(Regnskapsår == 2024) %>%
                              select("Fylke nummer",
                                     "Fylke navn",
                                     "Kumulativ areal i dekar",
                                     "Usikkerhet for kumulativ areal") %>%
                              left_join(., fylke_total_area, by = "Fylke nummer") %>%
                              mutate("Prosentandel (%) av nedbygging" = round2(`Kumulativ areal i dekar`/area_fylke * 100, 2), .before = "Usikkerhet for kumulativ areal") %>%
                              select(!c("Kumulativ areal i dekar", area_fylke)) %>%
                              st_as_sf()


# Replace labels of uncertainty
encacc_001_perc_map_mun$`Usikkerhet for kumulativ areal`[which(encacc_001_perc_map_mun$`Usikkerhet for kumulativ areal` == "Green")]<- "Sikker"

encacc_001_perc_map_mun$`Usikkerhet for kumulativ areal`[which(encacc_001_perc_map_mun$`Usikkerhet for kumulativ areal` == "Yellow")]<- "Ganske sikker"

encacc_001_perc_map_mun$`Usikkerhet for kumulativ areal`[which(encacc_001_perc_map_mun$`Usikkerhet for kumulativ areal` == "Orange")]<- "Ganske usikker"

encacc_001_perc_map_mun$`Usikkerhet for kumulativ areal`[which(encacc_001_perc_map_mun$`Usikkerhet for kumulativ areal` == "Red")]<- "Usikker"


encacc_001_perc_map_fy$`Usikkerhet for kumulativ areal`[which(encacc_001_perc_map_fy$`Usikkerhet for kumulativ areal` == "Green")]<- "Sikker"

encacc_001_perc_map_fy$`Usikkerhet for kumulativ areal`[which(encacc_001_perc_map_fy$`Usikkerhet for kumulativ areal` == "Yellow")]<- "Ganske sikker"

encacc_001_perc_map_fy$`Usikkerhet for kumulativ areal`[which(encacc_001_perc_map_fy$`Usikkerhet for kumulativ areal` == "Orange")]<- "Ganske usikker"

encacc_001_perc_map_fy$`Usikkerhet for kumulativ areal`[which(encacc_001_perc_map_fy$`Usikkerhet for kumulativ areal` == "Red")]<- "Usikker"

```

```{r map_enc001_perc_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Interactive map of the share of encroachments in municipalities over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/interactive_map.R"))

# Run function
interactive_map(data1 = encacc_001_perc_map_mun, 
                data2 = encacc_001_perc_map_mun, 
                fill_variable1 = "Prosentandel (%) av nedbygging", 
                fill_variable2 =  "Usikkerhet for kumulativ areal", 
                legend_title1 = "Prosentandel (%) av nedbygging", 
                legend_title2 = "Usikkerhet",
                style_type_map1 = "fixed",
                interval_breaks_map1 = c(0, 0.05, 0.1, 0.2, 0.5, 2.6),
                color_values_map2 = c("Sikker" = "#94F4B6",
                                      "Ganske sikker" = "#FFF085", 
                                      "Ganske usikker" = "#FFC68A",
                                      "Usikker" = "#FF8A8A"),
                map_title1 = "Prosentandel (%) av nedbygging innenfor det totale kommunale arealet over 2018-2024", 
                map_title2 = "Usikkerhet om prosentandelen av nedbygging over 2018-2024")

```

```{r map_enc001_perc_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of the the share of encroachments in counties over 2018-2024"

## Static maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = encacc_001_perc_map_fy, 
           data2 = encacc_001_perc_map_fy, 
           fill_variable1 = "Prosentandel (%) av nedbygging", 
           fill_variable2 =  "Usikkerhet for kumulativ areal", 
           legend_title1 = "Prosentandel (%) av nedbygging", 
           legend_title2 = "Usikkerhet",
           style_type_map1 = "fixed",
           interval_breaks_map1 = c(0, 0.01, 0.02, 0.03, 0.07),
           color_values_map2 = c("Sikker" = "#94F4B6",
                                 "Ganske sikker" = "#FFF085", 
                                 "Ganske usikker" = "#FFC68A",
                                 "Usikker" = "#FF8A8A"),
           map_title1 = "Prosentandel (%) av nedbygging innenfor det totale fylke arealet over 2018-2024", 
           map_title2 = "Usikkerhet om prosentandelen av nedbygging over 2018-2024")

```

```{r map_enc001_perc_4}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of the share of encroachments in municipalities over 2018-2024"

## Static maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = encacc_001_perc_map_mun, 
           data2 = encacc_001_perc_map_mun, 
           fill_variable1 = "Prosentandel (%) av nedbygging", 
           fill_variable2 =  "Usikkerhet for kumulativ areal", 
           legend_title1 = "Prosentandel (%) av nedbygging", 
           legend_title2 = "Usikkerhet",
           style_type_map1 = "fixed",
           interval_breaks_map1 = c(0, 0.05, 0.1, 0.2, 0.5, 2.6),
           color_values_map2 = c("Sikker" = "#94F4B6",
                                 "Ganske sikker" = "#FFF085", 
                                 "Ganske usikker" = "#FFC68A",
                                 "Usikker" = "#FF8A8A"),
           map_title1 = "Prosentandel (%) av nedbygging innenfor det totale kommunale arealet over 2018-2024", 
           map_title2 = "Usikkerhet om prosentandelen av nedbygging over 2018-2024")

```

```{r map_enc001_perc_5}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Clean R environment"

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "add_change_encroachment_area",
                    "add_fylke",
                    "add_uncertainty_individual",
                    "add_uncertainty_global",
                    "concatenate_fn",
                    "encacc_001_acc",
                    "encacc_001_fy_acc",
                    "encacc_002_acc",
                    "encacc_002_fy_acc",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum",
                    "fylke_geo",
                    "encacc_001_map_mun",
                    "encacc_001_map_fy",
                    "encacc_001_perc_map_mun",
                    "encacc_001_perc_map_fy")))

```

#### 10.2.3 ENCACC_002: Map of the total area of specially important nature lost due to land take in municipalities over 2018-2024
We first prepare the dataset for mapping then map it both using an interactive and static map. Due to memory issues in the html, we only display the static map. 
```{r map_enc002_1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Prepare dataset for mapping"

## Prepare dataset
# Municipalities
encacc_002_map_tot <- encacc_002_acc %>%
                        filter(!`Kommune navn` == "Norge") %>%
                        group_by(`Fylke nummer`, `Kommune nummer`, `Kommune navn`) %>%
                        summarise("Totalt areal verdsatte naturtyper gått tapt (dekar)" = sum(`Areal verdsatte naturtyper gått tapt (dekar)`))


encacc_002_map_un <- unc_002_idv %>%
                        mutate(accounting_period = "2018-2024") %>%
                        concatenate_fn(.,
                                       col_column_names = "uncertainty_source2",
                                       kom_num = "GISCO_ID",
                                       accounting_year = "accounting_period") %>%
                        add_uncertainty_global(., 
                                               kom_num = "GISCO_ID",
                                               uncertainty_col_name = "unc_concat_full",
                                               accounting_year = "accounting_period") %>%
                      ungroup() %>%
                      select(!c(unc_concat_full, 
                                global_cumulative_uncertainty,
                                GISCO_ID,
                                accounting_period)) %>%
                      cbind(encacc_002_map_tot, .) %>%
                      rename(global_period_uncertainty = global_yearly_uncertainty) %>%
                    left_join(., kom_navn, join_by("Kommune nummer" == "GISCO_ID")) %>%
                    select("Fylke nummer",
                           "Kommune nummer",
                           "Kommune navn",
                           "Totalt areal verdsatte naturtyper gått tapt (dekar)",
                           "global_period_uncertainty",
                           "geometry") %>%
                    rename("Usikkerhet av totalt areal" = "global_period_uncertainty") %>%
                    st_as_sf()


# Fylke
encacc_002_map_tot_fy <- encacc_002_fy_acc %>%
                          filter(!`Fylke navn` == "Norge") %>%
                          group_by(`Fylke nummer`, `Fylke navn`) %>%
                          summarise("Totalt areal verdsatte naturtyper gått tapt (dekar)" = sum(`Areal verdsatte naturtyper gått tapt (dekar)`))


encacc_002_map_un_fy <- unc_002_idv %>%
                        mutate(fylke_nr = add_fylke(., kommune_nummer = "GISCO_ID")[[1]],
                               accounting_period = "2018-2024") %>%
                        concatenate_fn(.,
                                       col_column_names = "uncertainty_source2",
                                       kom_num = "fylke_nr",
                                       accounting_year = "accounting_period") %>%
                        add_uncertainty_global(., 
                                               kom_num = "fylke_nr",
                                               uncertainty_col_name = "unc_concat_full",
                                               accounting_year = "accounting_period") %>%
                      ungroup() %>%
                      select(!c(unc_concat_full, 
                                global_cumulative_uncertainty,
                                fylke_nr,
                                accounting_period)) %>%
                      cbind(encacc_002_map_tot_fy, .) %>%
                      rename(global_period_uncertainty = global_yearly_uncertainty) %>%
                    left_join(., fylke_geo, join_by("Fylke nummer" == "fylke")) %>%
                    select("Fylke nummer",
                           "Fylke navn",
                           "Totalt areal verdsatte naturtyper gått tapt (dekar)",
                           "global_period_uncertainty",
                           "geometry") %>%
                    rename("Usikkerhet av totalt areal" = "global_period_uncertainty") %>%
                    st_as_sf()


# Replace labels of uncertainty
encacc_002_map_un$`Usikkerhet av totalt areal`[which(encacc_002_map_un$`Usikkerhet av totalt areal` == "Green")]<- "Sikker"

encacc_002_map_un$`Usikkerhet av totalt areal`[which(encacc_002_map_un$`Usikkerhet av totalt areal` == "Yellow")]<- "Ganske sikker"

encacc_002_map_un$`Usikkerhet av totalt areal`[which(encacc_002_map_un$`Usikkerhet av totalt areal` == "Orange")]<- "Ganske usikker"

encacc_002_map_un$`Usikkerhet av totalt areal`[which(encacc_002_map_un$`Usikkerhet av totalt areal` == "Red")]<- "Usikker"


encacc_002_map_un_fy$`Usikkerhet av totalt areal`[which(encacc_002_map_un_fy$`Usikkerhet av totalt areal` == "Green")]<- "Sikker"

encacc_002_map_un_fy$`Usikkerhet av totalt areal`[which(encacc_002_map_un_fy$`Usikkerhet av totalt areal` == "Yellow")]<- "Ganske sikker"

encacc_002_map_un_fy$`Usikkerhet av totalt areal`[which(encacc_002_map_un_fy$`Usikkerhet av totalt areal` == "Orange")]<- "Ganske usikker"

encacc_002_map_un_fy$`Usikkerhet av totalt areal`[which(encacc_002_map_un_fy$`Usikkerhet av totalt areal` == "Red")]<- "Usikker"

```

```{r map_enc002_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Interactive map of total area of specially important nature lost due to encroachment in municipalities over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/interactive_map.R"))

# Run function
interactive_map(data1 = encacc_002_map_un, 
                data2 = encacc_002_map_un, 
                fill_variable1 =  "Totalt areal verdsatte naturtyper gått tapt (dekar)", 
                fill_variable2 =  "Usikkerhet av totalt areal", 
                legend_title1 = "Areal av verdsatte natur gått tapt", 
                legend_title2 = "Usikkerhet",
                style_type_map1 = "fixed",
                interval_breaks_map1 = c(0, 1, 5, 15, 30, 100, 400),
                color_values_map2 = c("Sikker" = "#94F4B6",
                                      "Ganske sikker" = "#FFF085", 
                                      "Ganske usikker" = "#FFC68A",
                                      "Usikker" = "#FF8A8A"),
                map_title1 = "Totalt areal av verdsatte natur gått tapt (dekar) på nedbygging i kommuner over 2018-2024",
                map_title2 = "Usikkerhet om totalt areal av nedbygging over 2018-2024")
```

```{r map_enc002_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of map of total area of specially important nature lost due to encroachment in municipalities over 2018-2024"

## Static maps
# Call for the function to do static mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = encacc_002_map_un, 
           data2 = encacc_002_map_un, 
           fill_variable1 =  "Totalt areal verdsatte naturtyper gått tapt (dekar)", 
           fill_variable2 =  "Usikkerhet av totalt areal", 
           legend_title1 = "Areal av verdsatte natur gått tapt", 
           legend_title2 = "Usikkerhet",
           style_type_map1 = "fixed",
           interval_breaks_map1 = c(0, 1, 5, 15, 30, 100, 400),
           color_values_map2 = c("Sikker" = "#94F4B6",
                                 "Ganske sikker" = "#FFF085", 
                                 "Ganske usikker" = "#FFC68A",
                                 "Usikker" = "#FF8A8A"),
           map_title1 = "Totalt areal av verdsatte gått tapt (dekar) på nedbygging i kommuner over 2018-2024",
           map_title2 = "Usikkerhet om totalt areal av nedbygging over 2018-2024")


```

```{r map_enc002_4}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of map of total area of specially important nature lost due to encroachment in counties over 2018-2024"

## Static maps
# Call for the function to do static mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = encacc_002_map_un_fy, 
           data2 = encacc_002_map_un_fy, 
           fill_variable1 =  "Totalt areal verdsatte naturtyper gått tapt (dekar)", 
           fill_variable2 =  "Usikkerhet av totalt areal", 
           legend_title1 = "Areal av verdsatte natur gått tapt", 
           legend_title2 = "Usikkerhet",
           style_type_map1 = "fixed",
           interval_breaks_map1 = c(0, 100, 300, 500, 1000),
           color_values_map2 = c("Sikker" = "#94F4B6",
                                 "Ganske sikker" = "#FFF085", 
                                 "Ganske usikker" = "#FFC68A",
                                 "Usikker" = "#FF8A8A"),
           map_title1 = "Totalt areal av verdsatte natur gått tapt (dekar) på nedbygging i fylker over 2018-2024",
           map_title2 = "Usikkerhet om totalt areal av nedbygging over 2018-2024")

```

```{r map_enc002_5}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Clean R environment"

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "add_change_encroachment_area",
                    "add_fylke",
                    "add_uncertainty_individual",
                    "add_uncertainty_global",
                    "concatenate_fn",
                    "encacc_001_acc",
                    "encacc_001_fy_acc",
                    "encacc_002_acc",
                    "encacc_002_fy_acc",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum",
                    "fylke_geo",
                    "encacc_001_map_mun",
                    "encacc_001_map_fy",
                    "encacc_001_perc_map_mun",
                    "encacc_001_perc_map_fy",
                    "encacc_002_map_un",
                    "encacc_002_map_un_fy")))
```

#### 10.2.4 ENCACC_002: Share of land take that was specially important nature
We first prepare the dataset for mapping then map it both using an interactive and static map. Due to memory issues in the html, we only display the static map.
```{r map_enc002_perc_1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Prepare dataset for mapping"

## Prepare dataset
# Municipalities
## Prepare dataset
# Total area recreation in municipalities
kom_total_enc_area <- komenc %>%
                          st_drop_geometry() %>%
                          group_by(GISCO_ID) %>%
                          summarise(total_enc_area = round2(sum(area_inter_m2)/1000, 2))

# Total area recreation in Fylke
fylke_total_enc_area <- komenc %>%
                          st_drop_geometry() %>%
                          mutate(fylkesnr = add_fylke(data = ., kommune_nummer = "GISCO_ID")[[1]], .before = GISCO_ID) %>%
                          group_by(fylkesnr) %>%
                          summarise(total_enc_area = round2(sum(area_inter_m2)/1000, 2))

# Datasets for mapping
encacc_002_perc_map <- encacc_002_map_un %>%
                        ungroup() %>%
                        left_join(., kom_total_enc_area, join_by("Kommune nummer" == "GISCO_ID")) %>%
                        mutate("Prosentandel (%) av nedbygging som er verdsatte natur" = round2(`Totalt areal verdsatte naturtyper gått tapt (dekar)` / `total_enc_area` * 100, 2)) %>%
                        select("Fylke nummer",
                               "Kommune nummer",
                               "Kommune navn",
                               "total_enc_area",
                               "Totalt areal verdsatte naturtyper gått tapt (dekar)",
                               "Prosentandel (%) av nedbygging som er verdsatte natur",
                               "Usikkerhet av totalt areal",
                               "geometry") %>%
                        rename("Usikkerhet" =  "Usikkerhet av totalt areal",
                               "Totalt areal av nedbygging" = "total_enc_area")


encacc_002_perc_map_fy <- encacc_002_map_un_fy %>%
                            ungroup() %>%
                            left_join(., fylke_total_enc_area, join_by("Fylke nummer" == "fylkesnr")) %>%
                            mutate("Prosentandel (%) av nedbygging som er verdsatte natur" = round2(`Totalt areal verdsatte naturtyper gått tapt (dekar)` / `total_enc_area` * 100, 2)) %>%
                            select("Fylke nummer",
                                   "Fylke navn",
                                   "total_enc_area",
                                   "Totalt areal verdsatte naturtyper gått tapt (dekar)",
                                   "Prosentandel (%) av nedbygging som er verdsatte natur",
                                   "Usikkerhet av totalt areal",
                                   "geometry") %>%
                            rename("Usikkerhet" =  "Usikkerhet av totalt areal",
                                   "Totalt areal av nedbygging" = "total_enc_area")

# Replace labels of uncertainty
encacc_002_perc_map$`Usikkerhet`[which(encacc_002_perc_map$`Usikkerhet av totalt areal` == "Green")]<- "Sikker"

encacc_002_perc_map$`Usikkerhet`[which(encacc_002_perc_map$`Usikkerhet` == "Yellow")]<- "Ganske sikker"

encacc_002_perc_map$`Usikkerhet`[which(encacc_002_perc_map$`Usikkerhet` == "Orange")]<- "Ganske usikker"

encacc_002_perc_map$`Usikkerhet`[which(encacc_002_perc_map$`Usikkerhet` == "Red")]<- "Usikker"


encacc_002_perc_map_fy$`Usikkerhet`[which(encacc_002_perc_map_fy$`Usikkerhet` == "Green")]<- "Sikker"

encacc_002_perc_map_fy$`Usikkerhet`[which(encacc_002_perc_map_fy$`Usikkerhet` == "Yellow")]<- "Ganske sikker"

encacc_002_perc_map_fy$`Usikkerhet`[which(encacc_002_perc_map_fy$`Usikkerhet` == "Orange")]<- "Ganske usikker"

encacc_002_perc_map_fy$`Usikkerhet`[which(encacc_002_perc_map_fy$`Usikkerhet` == "Red")]<- "Usikker"

```


```{r map_enc002_perc_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Interactive map of the share of land take that are specially important nature in municipalities over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/interactive_map.R"))

# Run function
interactive_map(data1 = encacc_002_perc_map, 
                data2 = encacc_002_perc_map, 
                fill_variable1 = "Prosentandel (%) av nedbygging som er verdsatte natur", 
                fill_variable2 = "Usikkerhet",
                legend_title1 = "Prosentandel (%)", 
                legend_title2 = "Usikkerhet", 
                style_type_map1 = "fixed",
                interval_breaks_map1 = c(0, 0.5, 2, 4, 10, 50),
                color_values_map2 = c("Sikker" = "#94F4B6",
                                      "Ganske sikker" = "#FFF085",
                                      "Ganske usikker" = "#FFC68A",
                                      "Usikker" = "#FF8A8A"),
                map_title1 = "Prosentandel (%) av nedbygging som er verdsatte naturtyper i kommuner over 2018-\n2024", 
                map_title2 = "Usikkerhet av prosentandel over 2018-2024")


```

```{r map_enc002_perc_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of the share of nebygging that are specialy important nature in municipalities over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = encacc_002_perc_map, 
           data2 = encacc_002_perc_map, 
           fill_variable1 = "Prosentandel (%) av nedbygging som er verdsatte natur", 
           fill_variable2 = "Usikkerhet",
           legend_title1 = "Prosentandel (%)", 
           legend_title2 = "Usikkerhet", 
           style_type_map1 = "fixed",
           interval_breaks_map1 = c(0, 0.5, 2, 4, 10, 50),
           color_values_map2 = c("Sikker" = "#94F4B6",
                                 "Ganske sikker" = "#FFF085",
                                 "Ganske usikker" = "#FFC68A",
                                 "Usikker" = "#FF8A8A"),
           map_title1 = "Prosentandel (%) av nedbygging som er verdsatte naturtyper i kommuner over 2018-\n2024", 
           map_title2 = "Usikkerhet av prosentandel over 2018-2024")

```

```{r map_enc006_perc_4}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of the share of nebygging that are specially important nature in counties over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = encacc_002_perc_map_fy, 
           data2 = encacc_002_perc_map_fy, 
           fill_variable1 = "Prosentandel (%) av nedbygging som er verdsatte natur", 
           fill_variable2 = "Usikkerhet",
           legend_title1 = "Prosentandel (%)", 
           legend_title2 = "Usikkerhet", 
           style_type_map1 = "fixed",
           interval_breaks_map1 = c(0, 1.5, 2.5, 3, 5.5),
           color_values_map2 = c("Sikker" = "#94F4B6",
                                 "Ganske sikker" = "#FFF085",
                                 "Ganske usikker" = "#FFC68A",
                                 "Usikker" = "#FF8A8A"),
           map_title1 = "Prosentandel (%) av nedbygging som er verdsatte naturtyper i fylker over 2018-\n2024", 
           map_title2 = "Usikkerhet av prosentandel over 2018-2024")


```
<!--# 

Repeat the final results here. Typically this is a map or table of indicator values.

This is typically where people will harvest data from, so make sure to include all relevant output here, but don't clutter this section with too much output either.

-->

## 11. Export file
### 11.1 Tables
```{r export_tab}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Export tables"

# Accounting tables
write.table(st_drop_geometry(encacc_001_acc), 
            paste0(project_path,
                   "/result/tables/accounts/ENCACC001_accounting_table_kommune.csv"), 
            sep = ";",
            fileEncoding = "UTF-8",
            row.names = FALSE)

write.table(st_drop_geometry(encacc_001_fy_acc), 
            paste0(project_path,
                   "/result/tables/accounts/ENCACC001_accounting_table_fylke.csv"), 
            sep = ";",
            fileEncoding = "UTF-8",
            row.names = FALSE)

write.table(st_drop_geometry(encacc_002_acc), 
            paste0(project_path,
                   "/result/tables/accounts/ENCACC002_accounting_table_kommune.csv"), 
            sep = ";",
            fileEncoding = "UTF-8",
            row.names = FALSE)

write.table(st_drop_geometry(encacc_002_fy_acc), 
            paste0(project_path,
                   "/result/tables/accounts/ENCACC002_accounting_table_fylke.csv"), 
            sep = ";",
            fileEncoding = "UTF-8",
            row.names = FALSE)


# Tables associated to maps
write.table(st_drop_geometry(encacc_001_map_mun), 
            paste0(project_path,
                   "/result/tables/map_tables/ENCACC001_map_kommune.csv"), 
            sep = ";",
            fileEncoding = "UTF-8",
            row.names = FALSE)

write.table(st_drop_geometry(encacc_001_map_fy), 
            paste0(project_path,
                   "/result/tables/map_tables/ENCACC001_map_fylke.csv"), 
            sep = ";",
            fileEncoding = "UTF-8",
            row.names = FALSE)

write.table(st_drop_geometry(encacc_001_perc_map_mun), 
            paste0(project_path,
                   "/result/tables/map_tables/ENCACC001-share_map_kommune.csv"), 
            sep = ";",
            fileEncoding = "UTF-8",
            row.names = FALSE)


write.table(st_drop_geometry(encacc_001_perc_map_fy), 
            paste0(project_path,
                   "/result/tables/map_tables/ENCACC001-share_map_fylke.csv"), 
            sep = ";",
            fileEncoding = "UTF-8",
            row.names = FALSE)

write.table(st_drop_geometry(encacc_002_map_un), 
            paste0(project_path,
                   "/result/tables/map_tables/ENCACC002_map_kommune.csv"), 
            sep = ";",
            fileEncoding = "UTF-8",
            row.names = FALSE)

write.table(st_drop_geometry(encacc_002_map_un_fy), 
            paste0(project_path,
                   "/result/tables/map_tables/ENCACC002_map_fylke.csv"), 
            sep = ";",
            fileEncoding = "UTF-8",
            row.names = FALSE)

write.table(st_drop_geometry(encacc_002_perc_map), 
            paste0(project_path,
                   "/result/tables/map_tables/ENCACC002-share_map_kommune.csv"), 
            sep = ";",
            fileEncoding = "UTF-8",
            row.names = FALSE)

write.table(st_drop_geometry(encacc_002_perc_map_fy), 
            paste0(project_path,
                   "/result/tables/map_tables/ENCACC002-share_map_fylke.csv"), 
            sep = ";",
            fileEncoding = "UTF-8",
            row.names = FALSE)

```

### 11.2 Maps
We create a geopackage with all the different maps created above: (i) total area of land take; (ii) share of land take; (iii) total area of specially important nature lost to land take; (iv) share of land take that was specially important nature. All of these indicators were display by municipalities and fylke.

We first need to prepare the datasets because to display the maps on maps.nina.no, we need to create a variable for coloring.

```{r export_1}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Prepare datasets"
## Call function to format the map to NINA's standard
source(paste0(functions_path,"/map_nina_format.R"))

## Layer 1 - Total area of nedbygging in municipalities
nedbygging_kommune <- map_nina_format(encacc_001_map_mun, 
                                      mapping_variable = "Kumulativ areal av nedbygging i dekar", 
                                      mapping_groups = 5, 
                                      groups_thresholds = c(500, 1000, 1500, 2000, 2500))


## Layer 2 - Total area of nedbygging in fylke
nedbygging_fylke <- map_nina_format(encacc_001_map_fy, 
                                    mapping_variable = "Kumulativ areal av nedbygging i dekar", 
                                    mapping_groups = 5, 
                                    groups_thresholds = c(10000, 20000, 30000, 40000, 54000))


## Layer 3 - Share of nedbygging in municipalities
share_nedbygging_kommune <- map_nina_format(encacc_001_perc_map_mun,
                                            mapping_variable = "Prosentandel (%) av nedbygging", 
                                            mapping_groups = 5, 
                                            groups_thresholds = c(0.05, 0.1, 0.2, 0.5, 2.6))


## Layer 4 - Share of nedbygging in fylke
share_nedbygging_fylke <- map_nina_format(encacc_001_perc_map_fy,
                                          mapping_variable = "Prosentandel (%) av nedbygging", 
                                          mapping_groups = 5, 
                                          groups_thresholds = c(0.01, 0.02, 0.03, 0.07))

## Layer 5 - Total area of verdatte natur lost to nedbygging in municipalities
viktignatur_kommune <- map_nina_format(encacc_002_map_un, 
                                       mapping_variable = "Totalt areal verdsatte naturtyper gått tapt (dekar)", 
                                       mapping_groups = 5, 
                                       groups_thresholds = c(1, 5, 15, 30, 100, 400))



## Layer 6 - Total area of verdsatte natur lost to nedbygging in fylke
viktignatur_fylke <- map_nina_format(encacc_002_map_un_fy, 
                                     mapping_variable = "Totalt areal verdsatte naturtyper gått tapt (dekar)", 
                                     mapping_groups = 5, 
                                     groups_thresholds = c(100, 300, 500, 1000))

## Layer 7 - Share of nedbygging that was verdsatte natur in municipalities
share_viktignatur_kommune <- map_nina_format(encacc_002_perc_map, 
                                     mapping_variable = "Prosentandel (%) av nedbygging som er verdsatte natur", 
                                     mapping_groups = 5, 
                                     groups_thresholds = c(0.5, 2, 4, 10, 50))



## Layer 8 - Share of nedbygging that was verdsatte natur in fylke
share_viktignatur_fylke <- map_nina_format(encacc_002_perc_map_fy, 
                                           mapping_variable = "Prosentandel (%) av nedbygging som er verdsatte natur", 
                                           mapping_groups = 5, 
                                           groups_thresholds = c(1.5, 2.5, 3, 5.5))


```

Now we can export all the layers in the same geopackage. *Note: we need to export to the scratch folder because the Mounts cause problems when writing geopackages.*
```{r export_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Export datasets as geopackage"

# Path to the result geopackage
gpkg_path <- paste0(tempdir(), "/encroachment_accounts.gpkg")

st_write(nedbygging_kommune,
         dsn   = gpkg_path,
         layer = "nedbygging_kommune",
         append = FALSE)   

st_write(nedbygging_fylke,
         dsn   = gpkg_path,
         layer = "nedbygging_fylke",
         append = TRUE)    # append the new layer to same gpkg

st_write(share_nedbygging_kommune,
         dsn   = gpkg_path,
         layer = "prosent_nedbygging_kommune",
         append = TRUE)

st_write(share_nedbygging_fylke,
         dsn   = gpkg_path,
         layer = "prosent_nedbygging_fylke",
         append = TRUE)

st_write(viktignatur_kommune,
         dsn   = gpkg_path,
         layer = "natur_nedbygging_kommune",
         append = TRUE)

st_write(viktignatur_fylke,
         dsn   = gpkg_path,
         layer = "natur_nedbygging_fylke",
         append = TRUE)

st_write(share_viktignatur_kommune,
         dsn   = gpkg_path,
         layer = "natur_prosent_nedbygging_kommune",
         append = TRUE)

st_write(share_viktignatur_fylke,
         dsn   = gpkg_path,
         layer = "natur_prosent_nedbygging_fylke",
         append = TRUE)


```

<!--# 

Optional: Display the code (don't execute it) or the workflow for exporting the indicator values to file. Ideally the indicator values are exported as a georeferenced shape or raster file with indicators values, reference values and errors. You can also chose to export the raw (un-normalised or unscaled variable) as a separate product. You should not save large sptaial output data on GitHub. You can use eval=FALSE to avoid code from being executed (example below - delete if not relevant) 

-->

```{r export}
#| eval: false
```

<!-- You can leave this last part as it is, unless you have a good reason to alter it. -->

::: {.callout-tip collapse="true"}
## Session Info

This information is mainly aimed at developers that need to recreate the R environment that produced the html.
Important note: this information is about the R session that was running when the html was rendered, and if some of the code chunks were not evaluated when the html was rendered, they may have been executed in a different environment.  

```{r sessionInfo}
#| echo: false
sessionInfo()
```

:::

<!-- ```{r highest_habitat_lost} -->

<!-- habitat_summmary <- encnatur %>% -->
<!--                       group_by(naturtype) %>% -->
<!--                       summarise(total_area = round2(sum(area_naturtyper_m2)/1000, 2)) -->

<!-- habitat_summmary_un <- unc_002_idv %>% -->
<!--                         mutate(accounting_period = "2018-2024") %>% -->
<!--                         concatenate_fn(., -->
<!--                                        col_column_names = "uncertainty_source2", -->
<!--                                        kom_num = "naturtype", -->
<!--                                        accounting_year = "accounting_period") %>% -->
<!--                         add_uncertainty_global(., -->
<!--                                                kom_num = "naturtype", -->
<!--                                                uncertainty_col_name = "unc_concat_full", -->
<!--                                                accounting_year = "accounting_period") %>% -->
<!--                       ungroup() %>% -->
<!--                       select(!c(unc_concat_full, -->
<!--                                 global_cumulative_uncertainty, -->
<!--                                 naturtype, -->
<!--                                 accounting_period)) %>% -->
<!--                       cbind(habitat_summmary, .) %>% -->
<!--                       rename(global_period_uncertainty = global_yearly_uncertainty) -->

<!-- sum(habitat_summmary_un$total_area) -->

<!-- hab_max <- habitat_summmary_un %>% -->
<!--             as.data.frame() %>% -->
<!--             arrange(desc(.$total_area)) %>% -->
<!--             filter(total_area <0.21) -->








<!-- max_kom <- encacc_002_perc_map %>% -->
<!--             as.data.frame() %>% -->
<!--             arrange(desc(.$`Prosentandel (%) av nedbygging som er ''særlig viktig natur''`)) %>% -->
<!--             filter(`Prosentandel (%) av nedbygging som er ''særlig viktig natur''`< 10) -->

<!-- max_kom <- encacc_002_map_un %>% -->
<!--             as.data.frame() %>% -->
<!--             arrange(desc(.$`Totalt areal særlig viktige naturtyper gått tapt (dekar)`)) %>% -->
<!--             filter(`Totalt areal særlig viktige naturtyper gått tapt (dekar)`< 100) -->

<!-- tot_enc <- encacc_001_acc %>% -->
<!--             as.data.frame() %>% -->
<!--             filter(!is.na(`Årlig areal i dekar`)) %>% -->
<!--             filter(Regnskapsår == 2024) %>% -->
<!--   filter(`Kommune navn` %in% c("Nesbyen", "Flesberg", "Gausdal", "Sigdal", "Flå", -->
<!--                                          "Sandnes", "Skien", "Sunnfjord", "Tønsberg", "Kongsberg" -->
<!-- )) %>% -->
<!--   select("Kommune nummer", -->
<!--          "Kommune navn", -->
<!--          "Kumulativ areal i dekar") -->

<!-- essai <- left_join(max_kom, tot_enc, by = c("Kommune navn", "Kommune nummer")) %>% -->
<!--           st_drop_geometry() %>% -->
<!--           mutate(perc = `Kumulativ areal av nedbygging i dekar` / `Kumulativ areal i dekar` * 100) %>% -->
<!--             arrange(desc(.$`perc`)) %>% -->
<!--             filter(`perc`> 1211) -->

<!-- ``` -->
