---
title: "Thematic accounts - encroachment extent accounts in recreation areas"
subtitle: |
  [ENCREC_001] (Extent accounts of encroachment in all recreation areas) \
  [ENCREC_002] (Extent accounts of encroachment in very important (A) and important (B) recreation areas) \
  [ENCREC_003] (Extent accounts of all recreation areas) \
  [ENCREC_004] (Extent accounts of very important (A) and important (B) recreation areas) \
  [ENCREC_005] (Extent accounts of sværlig vikitig natur lost due to encroachments in all recreation areas) \
  [ENCREC_006] (Extent accounts of sværlig vikitig natur lost due to encroachments in very important (A) and important (B) recreation areas)
# delete the line with the second indicatorID (line 5) 
# if it's not relevant for you.
# In the parentheses you can add a word to distinguish multiple indicators 
# from each other, for example the ecosystem type that they cover
indicatorID: 
  - ENCREC_001
  - ENCREC_002
  - ENCREC_003
  - ENCREC_004
  - ENCREC_005
  - ENCREC_006 # delete whole line if not needed
# If you are presenting multiple indicators, and have multiple 
# indicator IDs, then add them as a list here, and make sure you add
# the same number of list entries, 
# and in the same order that corresponds to the indicator IDs,  
# to the meta data fields where the indicators differ. 
# For example, if the indicators cover different ecosystem types, 
# add the ecosystems as a list under 'Ecosystems', but if they are in the
# same realm, same country etc., then you only need to write the realm and
# country once. 
indicatorName: 
  - Extent accounts of encroachment in all recreation areas # add more by adding another line starting '-'
  - Extent accounts of encroachment in very important (A) and important (B) recreation areas
  - Extent accounts of all recreation areas
  - Extent accounts of very important (A) and important (B) recreation areas
  - Extent accounts of sværlig vikitig natur lost due to encroachments in all recreation areas
  - Extent accounts of sværlig vikitig natur lost due to encroachments in very important (A) and important (B) recreation areas
folderName: code # name of the folder under 'indicators/'. 
# If it's a single indicator being reported, the folder name should be the same as the indicatorID. 
# If there are multiple ID's then the folder name could look something like XX_YYY_ZZZ-ÅÅÅ.
country: Norway
continent: Europe
# ECT: write ECT here # e.g. A2 - Chemical State characteristics
Realm: 
  - write realm here # e.g. Freshwater (F)
Biome: 
  - write biome here # e.g. F1 Rivers and streams biome
Ecosystem: 
  - write ecosystem here # e.g. F1.3 Freeze-thaw streams
yearAdded: 2025 # the year when the indicator was first submitted to ecRxiv
yearLastUpdate: 2025
status: incomplete # options are complete, incomplete, or deprecated
Version: "000.001" # Beta-versions are numbers incrementally with values below 0. Operational versions can be given the version numbers above 0. Use this terms similar as for software versioning,
VersionComment: This is the first version of the thematic encroachemnt extent accounts on recreation encroachment being compiled # Describe briefly how this version differs from earlier versions 
url: https://github.com/NINAnor/recereation_encroachment_accounts
# Leave the url as is, but replace XX_YYYY_ZZ with the name of the root folder for the current indicator(s). It should be the same as the indicatorID, but if there are more indicator IDs presented in the same folder, the folder name might be different, e.g. XX_YYYY_AAA-BBB 
Normalised: No # Yes or no depending on whether the metric has been normliased or not.
SpatialAggregationPathway: Describe the aggregation using the standrdaise terminology # e.g. Scale - Transform - Aggregate (length-weighted arithmetic mean) - Truncate
format: 
  html:
    embed-resources: true
    code-fold: true
    toc: true
    toc-title: Contents
    toc-depth: 3
    smooth-scroll: true
execute: 
  cache: true
author:
  - name: Sylvie Clappe             # Enter name
    email: sylvie.clappe@nina.no  # Enter email
    affiliations:
      - id: myID
        name: Norwegian Institute for Nature Research  # Enter affiliations
  - name: David Barton              #  Enter subsequent authors like this, or remove if not relevant
    affiliations:
      - ref: myID               # To reuse affiliations reference the id like this
  - name: Reto Spielhofer             #  Enter subsequent authors like this, or remove if not relevant
    affiliations:
  - ref: myID 
AuthorList: Clappe, S., Barton, D.N., Spielhofer, R.  
date: January 22, 2026 # Enter date
data_availability: none
code_reproducibility: none
open_science_badge: none
callout-icon: false
lightbox: true
css: ../../../style.css
code-links:
      - text: Add a review
        icon: github
        href: https://github.com/NINAnor/ecRxiv
hide: TRUE # Change to FALSE
---

<!--# This is a template for how to document the indicator analyses. Make sure also to not change the order, or modify, the headers, unless you really need to. This is because it easier to read if all the indicators are presented using the same layout. If there is one header where you don't have anything to write, just leave the header as is, and don't write anything below it. If you are providing code, be careful to annotate and comment on every step in the analysis. Before starting it is recommended to fill in as much as you can in the metadata file. This file will populate the initial table in your output.-->

<!--# Load all you dependencies here -->

```{r setup}
#| include: false
#| cache: false
library(knitr)
library(tidyverse)
library(kableExtra)
library(here)
library(anybadger)
library(yaml)
library(tibble)
library(conflicted)
library(httr2)
library(giscoR)
library(zen4R)
library(jsonlite)
library(sf)
library(purrr)
library(numform)
library(DT)
library(tmap)

# weird workarounds for cache issues
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::lag)
conflicts_prefer(dplyr::select)
pull<-dplyr::pull

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Remove dplyr and messages
options(dplyr.summarise.inform = FALSE)
```

```{r source}
#| echo: false
#| cache: false
source(here::here("_common.R"))
```

```{r working_directory}
#| include: false
#| cache: false

# Workaround to force the working directory
# Needs to be set after {r source} otherwise
# {_common.R is not found}
opts_knit$set(root.dir = here:::here())
```


<!--# The following parts are autogenerated. Do not edit. -->

::: {layout-ncol="5"}

```{r statusBadge}
#| echo: false
status_badge(as.character(st))
```

```{r versionBadge}
#| echo: false
version_badge(my_version_number = as.character(version[[1]]),
  folder_name = folder_name)
```

```{r dataBadge}
#| echo: false
data_badge(as.character(badges[[1]]))
```

```{r codeBadge}
#| echo: false
code_badge(as.character(badges[[2]]))
```

```{r openScienceBadge}
#| echo: false
open_science_badge(as.character(badges[[3]]))
```
::: 
 

> **Recommended citation**: `r paste(auth, " ", year, ". ", name, " (ID(s): ", folder_name, ") ", "v. ", version, ". ", url, sep="")`



::: {.callout-tip collapse="true"}
## Metadata
```{r tbl-meta2}
#| tbl-cap: 'Indicator metadata'
#| echo: false
#| warning: false
meta |>
  dplyr::filter(Variable %in% c(
    "Indicator ID",
    "Indicator Name",
    "Continent",
    "Country",
    "Ecosystem Condition Typology Class", 
    "Realm",
    "Biome",
    "Ecosystem",
    "Year added", 
    "Last update",
    "Version",
    "Version comment",
    "Normalised",
    "Spatial aggregation pathway")) |>   
  kable()

```
:::

::: {.callout-tip collapse="true"}

## Log

<!--# Update this logg with short messages for each update -->
- 22 January 2026 - Analyses completed
- XXX - First draft version
:::


<hr />

<!--# Document you work below.  -->

## 1. Summary

<!--# 
With a maximum of 300 words, describe the indicator in general terms as you would to a non-expert. Think of this as a kind of common language summary. Follow the four numbers sub-headers below. under point 2, it is possible under  include a bullet point list of the specific steps in the workflow. 
 -->

1. **Background and rationale**:
<!-- What does the metric represent? Why is this relevant for describing ecosystem condition in this ecosystem?
What are the main anthropogenic impact factors? -->
2. **Methods**:
<!-- What kind of data is used? How is the data customized (modified, estimated, integrated) to fit its purpose as an indicator? -->
3. **Interpretation**:
<!-- Add a very short description of how the indicator values should be interpreted. Here's an example for a bumblebee indicator: An indicator value of 1 signifies an intact bumblebee community. Values decreasing toward zero reflect negative ecological change, specifically the decline of species expected to be present. This metric does not account for potentially positive changes, such as increases in species' abundance or the arrival of new species. Tip: Add text on the line directly following the subheader, and don't introduce blank lines. This way the whole paragraph renders as a block quote. -->
An indicator value of 1 signifies ...
Values decreasing toward zero reflect...

4. **Results and development needs**:
<!-- Briefly, what does the indicator say about the ecosystem condition. What is the current status of  the metric (can it be used or is it still in development), and what immediate tasks remain to improve the indicator or to make the indicator operational -->



## 2. About the underlying data

<!--# Give a general introduction the datasets you have used, and provide more specific details in the sub-chapters.  -->

### 2.1 Spatial and temporal resolution and extent

<!--# Describe the temporal and spatial resolution and extent of the data used -->

### 2.2 Sampling design

<!-- How was the data sampled? Describe potential sampling biases -->

### 2.3 Original units

<!--# What are the original units for the most relevant  variables in the data-->


### 2.3 Instructions for citing, using and accessing data

<!--# Is the data openly available? If not, how can one access it? What are the key references to the datasets?   -->

### 2.4 Additional comments about the dataset

<!--# Optional. Text here -->

## 3. Indicator properties

### 3.1 Collinearities with other indicators

<!--# Describe known collinearities with other metrics (indicators or variables) that could become problematic if they were also included in the same Ecosystem Condition Assessment as the indicator described here. -->


### 3.2 Impact factors

<!--# Describe the main natural and anthropogenic factors that affecst the metric -->

### 3.3 Spatial aggregation

<!-- Describe in word the same aggregation pathway presented in the top yaml (the metadata). If you have scaled against the reference value, it is especially important to know if you did any avergaing or summations etc. beforehand. An example text for a rather simple case could be: "The variable is scaled at the initial polygon level and the point estimate is a area-weighted arithmetic mean of all polygins. There is no transformation, and truncation is not needed as the indicator becomes naturally bound between 0 and 1".  -->


## 5. Uncertainties
### 5.1 General uncertainties
<!--# Describe the main reasons why you might not trust 100% the final value to represent the given ecosystem characteristic for the give ecosystem type. What aspects of this condition metric are most sensitive to subjective decitions? Is the spatial sampling sufficient? Does the choice of modeling framework affect the values? Is the metric relevant do describe ecosystem condition for the give ecosystem type?  -->

Be careful: we did not add any restoration: we just have additions and not "reductions" in the extent of the encroachment. THis is a bias, it might be that the figures are smaller in reality.

In cumulative accounts, we assume emcroachment add over the years but in fact some might have disappeared (restoration as mentioned above)


### 5.2 Reported uncertainties
There are different sources of uncertainty associated with our results (see 5.1 General uncertainties). We chose to create a traffic light system based on 2 of these sources of uncertainty: (i) the prediction of encroachment; (ii) the temporal discrepancy between datasets. This was done to make the interpretation of the results easier for the decision-makers.

We created a lookup table based on these two sources of uncertainty. This lookup table allowed to assign an uncertainty, following a traffic-light color, at polygon level. We would then aggregate the polygon-uncertainty to the municipal or fylke level and derive the final traffic-light color based on the frequency of polygon-uncertainty. The full documentation of this system can be found in our [NINA report XXXX](**<LINK>**).

Despite making things easier fr decision-makers, our approach has several limitations. One of them uncertainty of DT03 at national level and not polygon level

It also have to be noted that, despite having a quantitative uncertainty for DT03, this uncertainty was quantified at national level and does not apply at municipal level. We however developed an application that will help in quantify the uncertainty in the prediction of encroachment at fylke level. You want to help? Just go [here](**<LINK>**)! Description on how this qualitative indicator can be compiled is available in section 5.2 Reported uncertainties.



<!-- Describe the nature of the errors you (hopefully) provide alongside the point estimate for the condition metric. It is especially important to clarify if the errors are inferential (referring to the uncertainty of the point estimate) or descriptive (referreing to the data itself). Here' an example text: "What we report as uncertainties in @tbl-finalTable are inferential and refer to the uncertainty around the point estimates. It does not directly describe variation in the data itself. 
However, the variation in the mean is defined by the spatial variation only, and quartiles are obtained by bootstrapping the set of localities within a region, and calculating the indicator for each of 9999 such samples".   -->



## 6. References

<!--# You can add references manually or use a citation manager and add intext citations as with crossreferencing and hyperlinks. See https://quarto.org/docs/authoring/footnotes-and-citations.html -->

## 7. Import and prepare datasets
In this section, we show how to import the data through API. We also prepare them  for analyses. We first set up the path that will be used during the project. Users can change it to match their own path.
```{r path_setup1}
#| eval: true
#| code-summary: "Path setup"

# Path to the project folder
project_path <- here::here()

# Path to the functions folder within the project folder
functions_path <- here::here(project_path, "src/functions")
```

<!--# All the data import and preparation should be done in this chapter, before the analysis chapter. As a minimum, add code for importing data, and describe or show the data structure. You may also explain certain variables, what they represent, and their distributions -->

### 7.1 Map of municipalities in Norway - DT01
#### 7.1.1 Download & read DT03 from GISCO through an API
For DT03, the API process is much simpler as a package has been developed to read data from GISCO via an API: `giscoR`. We just nee dto use one function to cache the dataset in a temporary directory and read it in R.
```{r lau_down_data}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT01 API download"

# Create a temporary directory to store the cache
temp_dir <- tempdir()

# Download the dataset
lau_no <- gisco_get_lau(year = "2024",
                        epsg = "3035",
                        cache = TRUE,
                        update_cache = TRUE,
                        cache_dir = temp_dir,
                        verbose = FALSE,
                        country = "NO",
                        gisco_id = NULL)

```

#### 7.1.2 Clean and format the data for analyses
We select the columns of relevance to our work and clean the topological and geometrical errors if any. Note that we also need to clean the `GISCO_ID`, which is the municipality number. 

```{r clean_lau}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "DT03 - general clean and format"

# We select the columns of interests
lau_no <- lau_no %>%
            select(GISCO_ID,
                   LAU_NAME) %>%
            mutate(GISCO_ID = sub("NO_", "", .$GISCO_ID)) %>%
            st_transform(., crs = "EPSG:25832") # project to ESPG : 25832

# Check for topological and geometrical errors
which(st_is_valid(lau_no) == FALSE) %>% length()
which(st_is_empty(lau_no) == TRUE) %>% length()

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "project_path",
                    "functions_path")))
```

### 7.2 Recreation areas - DT02
#### 7.2.1 Download DT02 for Geonorge through an API
We download DT02 through an API from [Geonorge](https://kartkatalog.geonorge.no/metadata/friluftslivsomraader-kartlagte/91e31bb7-356f-4478-bcba-d5c2de6e91bc?search=friluftsliv). Downloading through API in Geonorge is not an easy process, hence we fully decompose the API download request to facilitate the download of other datasets by users if they wish to. Note however that datasets not publicly accessible cannot be downloaded via an API.

1. **Create a  <u>base url</u>**. The *base url* is the basic url that needs to be used when downloading data. It will be the same for all datasets. In Geonorge, users should go the their datsaset of interest and scroll to the bottom of the page to the section *Distribusjon*. There, there are several subsections called *Distribusjonstype*. They display the *base url* depending on where the users want to download the data from. In our case, we chose to download our data from Geonorge whose *base url* is https://geoapi.p.aks.miljodirektoratet.no/geonorge. Here we removed the last part of the url (/capabilities) as it refers to another concept explained in the point 3 below.

```{r import_rec_areas_api1}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT02 API download - base url"

# base url for downloading dataset in Geonorge through API
base_url <- "https://geoapi.p.aks.miljodirektoratet.no/geonorge"
```


2. **Identify the <u>dataset ID </u>**. The dataset ID is called an uuid. It can be found by going to the dataset's page on Geonorge, it is the long series of numbers and letters in the url of the page: ".../metadata/dataset_name/uuid"
```{r import_rec_areas_api2}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT02 API download - dataset ID"

# dataset specific ID from Geonorge
dataset_uuid <- "91e31bb7-356f-4478-bcba-d5c2de6e91bc"
```

3. **Request the dataset's <u>capabilities</u>**. When downloading data through an API, users have to choose what is called "capabilities". These refer to the settings that should be used when downloading the data such as geographic areas, raster or vector type, format etc. We need to first build an API request to know what these capabilities are before downloading the data.
```{r import_rec_areas_api3}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT02 API download - list capabilities"

# httr2:::req_perform() sends API request for dataset's capabilities
dt_cap <- request(sprintf("%s/capabilities/%s", base_url, dataset_uuid)) %>% # create request
            req_headers("Accept" = "application/json") %>% # create request
            req_perform()


# Read the JSON response of the server listing the capabilities
cap_all <- resp_body_json(dt_cap, simplifyVector = TRUE)
```


4. **Choose the dataset's <u>capabilities</u> of relevance**. Now that we have the list of all the dataset's capabilities, we can choose those we want to download the dataset with. We are interested in three capabilities: projection, format, area and order. *Note: (i) order is the url to be used to order the dataset to the server; (ii) in the code below, the "rel" column just indicates what the link "href" refers to, meaning project, format, area, order etc. The url to use are in the "href" column; (iii) four projections can be chosen but in reality, only two are available for Norway as a country: ESPG:25833 and WGS84. Trying the others codes will lead to an error. (iv) we just read "http://rel.geonorge.no/download/area" as it has the area together withe project and format. So we don't need to read them all separately*
```{r import_rec_areas_api4}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT02 API download - choose capabilities"

# Extract the capablities into a dataframe so it's easier to choose from
cap_choice <- request(cap_all$`_links`[cap_all$`_links`$rel == "http://rel.geonorge.no/download/area", "href"]) %>%
                req_perform() %>%
                resp_body_json()

# Separate each capabilities: project, format and area
projection <- cap_choice[[1]]$projections[[1]] # ESPG:25833

format <- cap_choice[[1]]$formats[[2]]$name # FILEGDB

area <- cap_choice[[1]] # Norway with Svalbard

order_url <- cap_all$`_links`[cap_all$`_links`$rel == "http://rel.geonorge.no/download/order", "href"]

```


5. **Create an <u>order</u>**. Here we create the API request to order the data from the server. In our case, the server responds with a downloadable link that we can then use. Every server has its own JSON request structure, it is important to go on the [API help page of Geonorge](https://nedlasting.geonorge.no/Help) to use the correct structure. As we use a POST API order, we need to go to the [POST API oder page](https://nedlasting.geonorge.no/Help/Api/POST-api-order) to figure it out.
```{r import_rec_areas_api5}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT02 API download - order the data"

# Create the order by listing the dataset id and capabilities of relevance
order_body <- list(
  downloadAsBundle = FALSE,
  email = "sylvie.clappe@nina.no",
  orderLines = list(
    list(
      areas = list(
        list(
          code = area$code,  
          name = area$name,
          type = area$type
        )
      ),
      formats = list(
        list(name = format)
      ),
      metadataUuid = dataset_uuid,
      projections = list(
        list(
          code = projection$code,
          name = projection$name,
          codespace = projection$codespace
        )
      )
    )
  )
)

# Send the order to the server
order_req <- request(order_url) %>% # create request
                req_body_json(order_body) %>% # create request
                req_headers("Accept" = "application/json") %>% # create request
                req_perform() # send request

# Read the order information
order_info <- resp_body_json(order_req)
```

6. **Download the data**. Now that we have the downloadable link, we just have the use it to download the dataset in R. *Note: we download the data in a temporary file which can then be deleted.*
```{r import_rec_areas_api6}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT02 API download - download and read the data"

# URL to download the data
down_url <- order_info$files[[1]]$downloadUrl

# Create a temporary directory to store the dataset .zip and unzipped
temp_dir <- tempdir()

# Create a temporary file to store the dataset .zip
destination_tmp <- tempfile(fileext = ".zip", tmpdir = temp_dir)

# Download the .zip file containing the dataset
download.file(down_url, destfile = destination_tmp, mode = "wb")

# Unzip the .zip
unzip(destination_tmp, exdir = temp_dir)

# Use "\\.gdb$" pattern to identify the file to read
gdb_file <- list.files(temp_dir, pattern = "\\.gdb$", full.names = TRUE)

```


#### 7.2.2 Read DT02
Now that the dataset has been downloaded in a temporary file, we can read it in R and clean the R environment. Once the analyses are finished, we will remove the temporary file containing the data.
```{r read_rec_area}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT02 - read the data"

# Read data
rec_no <- st_read(gdb_file)

# Clean R environment
rm(list = setdiff(ls(), 
                  c("rec_no",
                    "lau_no",
                    "project_path",
                    "functions_path")))
```

#### 7.2.3 Clean and format the data for analyses
We first need to select the column of relevance for our work and clean the topological and geometrical errors if any.

```{r clean_rec_area1}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "DT02 - clean"

# Select columns of relevance
rec_areas <- rec_no %>%
              select(fylkesnr,
                     kommune,
                     kartleggingsår,
                     kartlagtStatus,
                     identifikasjon_lokalid,
                     områdeverdi)

# Check for topological and geometrical errors
which(st_is_valid(rec_areas) == FALSE) %>% length()
which(st_is_empty(rec_areas) == TRUE) %>% length()

# Correct invalid geometries
rec_cln <- st_make_valid(rec_areas)

# Check correction worked
which(st_is_valid(rec_cln) == FALSE) %>% length()
which(st_is_empty(rec_cln) == TRUE) %>% length()

# Clean a typo
rec_cln$kartleggingsår[rec_cln$kartleggingsår == "2921"] <- "2021"

```

DT02 includes both terrestrial and marine outdoor recreation areas. That is why we first did a spatial intersection between DT02 and DT01 to restrict the analyses to terrestrial outdoor recreation areas only. The reason for this is that DT03 only encompasses encroachment within terrestrial areas. Note that we had to first dissolve municipal boundaries of DT01 as we are only interested in correcting the coastal boundaries of outdoor recreation areas.

```{r clean_rec_area2}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "DT02 - restrict to terrestrial recreation areas only"

# Check the CRS of both DT01 and DT02
st_crs(lau_no) == st_crs(rec_cln)


# Re-project rec_cln into ESPG : 25832
rcl_32 <- st_transform(rec_cln, crs = st_crs(lau_no))

# Perform spatial intersection to have terrestrial recreation areas only
rcl_32_tr <- lau_no %>%
              st_union() %>%
              st_intersection(rcl_32, .) %>%
              mutate(area_rec_m2 = unclass(st_area(.)), .after = områdeverdi)

```

The last step of formatting for the recreation data is to correct the year of mapping of municipalities with several and only report the most recent year. Except for Oslo, where the different district mapped their recreation areas in different years, the other municipalities should only have one year. This is in fact not the case in the recreation dataset. 196 municipalities have polygons appearing for more than one year. This is because they updated their polygons by adding new ones. So we need to take into account this year of update as the final mapping year.


```{r clean_rec_area3}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "DT02 - correct years of mapping"

# Select municipalities have more than one year
several_yr_mun <- rcl_32_tr %>%
                    st_drop_geometry() %>%
                    select(kommune,
                           kartleggingsår) %>%
                    filter(!kommune == "0301") %>%
                    group_by(kommune) %>%
                    summarise(count_yr = n_distinct(kartleggingsår)) %>%
                    filter(count_yr >1) %>%
                    ungroup() 

# Select the most recent year of update as the mapping year
update_yr_mun <- rcl_32_tr %>%
                    st_drop_geometry() %>%
                    select(kommune,
                           kartleggingsår) %>%
                    filter(kommune %in% as.vector(several_yr_mun$kommune)) %>%
                    group_by(kommune) %>%
                    arrange(kartleggingsår, .by_group = TRUE) %>% 
                    slice_tail(n = 1) %>%
                    ungroup() %>%
                    rename(kartleggingsår_updated = kartleggingsår)
                    
# Replace all the mapping years of the 103 municipalities in update_yr_mun by their last year of update
rcl_final <- rcl_32_tr %>%
                left_join(., update_yr_mun, by = "kommune") %>%
                mutate(kartleggingsår = if_else(!is.na(kartleggingsår_updated), 
                                                kartleggingsår_updated, 
                                                kartleggingsår)) %>%
                select(!kartleggingsår_updated)


# Clean R environment
rm(list = setdiff(ls(), 
                  c("rcl_32_tr",
                    "lau_no",
                    "rcl_final",
                    "project_path",
                    "functions_path")))


```


### 7.3. Encroachment areas - DT03
#### 7.3.1 Download DT03 from Zenodo through an API
The process is very similar than for Geonorge. Instead of going through all the steps above, we created a function that users can use to retrieve the downloadable *content* link to the data. We could have used `zen4R` package but we decided to create our own function to read the data from Zenodo directly instead of downloading them locally before reading them.
```{r enc_link}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT03 API download"

# Call for the function that will read the data through an API
source(paste0(functions_path,"/api_zenodo.R"))

# Retrieve downloadable link to the data
encroach_link <- api_zenodo(zenodo_url = "https://zenodo.org/records/15130992")

```

#### 7.2.2 Read DT03
Now that we have the downloadable link, we can read the data in R and clean the R environment.
```{r enc_read}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT03 - read"

encroach_dt <- st_read(encroach_link)

```

#### 7.3.3 Clean and format the data for analyses
All columns are of relevance to our work but we need to clean the topological and geometrical errors if any.

```{r clean_enc}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "DT03 - clean"

# Check for topological and geometrical errors
which(st_is_valid(encroach_dt) == FALSE) %>% length()
which(st_is_empty(encroach_dt) == TRUE) %>% length()

# Clean R environment
rm(list = setdiff(ls(), 
                  c("rcl_32_tr",
                    "lau_no",
                    "rcl_final",
                    "encroach_dt",
                    "project_path",
                    "functions_path")))

```

The data are all clear and do not need further corrections.


### 7.4 Naturtyper - verdsatte - DT04
#### 7.4.1 Download DT04 for Geonorge through an API
We use the same protocol as for DT01. This time, we created a function for it.
```{r import_natur_api1}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT04 API download"
# Call for the function that will read the data through an API
source(paste0(functions_path,"/api_geonorge.R"))

# Retrieve and download the data
natur_link <- api_geonorge(geonorge_url = "https://kartkatalog.geonorge.no/metadata/naturtyper-verdsatte/64cbb884-a19d-4356-a114-380cfe4a7314",
                              email_address = "sylvie.clappe@nina.no")

```

#### 7.4.2 Read DT04
Now that the dataset has been downloaded in a temporary file, we can read it in R and clean the R environment. Once the analyses are finished, we will remove the temporary file containing the data.
```{r read_nature}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT04 - read the data"

# Read data
naturtyper <- st_read(natur_link[grepl("Naturtyper", natur_link)])
```

#### 7.4.3 Clean and format the data for analyses
We first need to select the column of relevance for our work and clean the topological and geometrical errors if any. The two major issues we have with these data are the spatial and temporal coverages. Spatial coverage because the data do not cover the entire country of Norway. This means that it's not because we don't record one a nature type as occurring that it is indeed not occurring, it might just not have been sampled yet. Also, the nature types have been recorded only once for a wide temporal coverage: from 1905 to 2025. Given that our encroachment data cover 2017-2024, we decided to keep only the data that have been recorded in 2015 or after. We consider that a record in 2025 doesn't mean that the type was not present before. We also assume that nature types are relatively stable and have not change since 2017. We also removed all nature types whose year of report was not stated. This made us keep 69.5% of the original dataset, which is still very good for analyses.

```{r clean_natur}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "DT04 - clean and format"

# Select columns of relevance
natur_types <- naturtyper %>%
                mutate(year_natur_recorded = substr(.$registreringsdato, 1, 4)) %>%
                select(naturtype,
                       year_natur_recorded,
                       kommunenummer)

# Check for topological and geometrical errors
which(st_is_valid(natur_types) == FALSE) %>% length()
which(st_is_empty(natur_types) == TRUE) %>% length()

# Correct invalid geometries
natur_cln <- st_make_valid(natur_types)

# Check correction worked
which(st_is_valid(natur_cln) == FALSE) %>% length()
which(st_is_empty(natur_cln) == TRUE) %>% length()

# Clean typos in years
natur_cln$year_natur_recorded[natur_cln$year_natur_recorded == "2102"] <- "2012"
natur_cln$year_natur_recorded[natur_cln$year_natur_recorded == "2913"] <- "2013"


# Remove nature types that have been recorded before 2010
natur_cln <- natur_cln %>%
              filter(year_natur_recorded > 2009) %>%
              filter(!is.na(year_natur_recorded))
  

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "rcl_32_tr",
                    "rcl_final",
                    "encroach_dt",
                    "natur_cln",
                    "lau_no",
                    "project_path",
                    "functions_path")))

```

## 8. Spatial units

<!--# Describe the spatial units that you rely on in your analyses. Highlight the spatial units (the resolution) that the indicator values should be interpreted at. Potential spatial delineation data (maksing layers etc.) should be introduced under chapert 7. We recommend using the SEEA EA terminology of Basic Spatial Units (BSU), Ecosystem Asses (EA) and Ecosystem Accounting Area (EAA). -->

## 9. Analyses
### 9.1 ENCREC_001 to ENCREC_0004
Both the analyses for ENCREC_001 to ENCREC_004 rely on spatial intersections between DT02 and DT03. Also, note that we re-calculate the areas of the encroachment polygons after the intersection with DT02 has been performed.
```{r encrec001234_analyses}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "ENCREC_001 and 002 - intersection"

# Check the CRS of both DT02 and DT03
st_crs(rcl_final) == st_crs(encroach_dt)

# Perform spatial intersection
encrec <- st_intersection(rcl_final, encroach_dt) %>%
            mutate(area_inter_m2 = unclass(st_area(.)))

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_cln",
                    "project_path",
                    "functions_path",
                    "rcl_final",
                    "encrec")))

```
### 9.2 ENCREC_005 and ENCREC_006
For ENCREC_005 and ENCRREC_006, we need a spatial intersection between DT04 and `encrec` as we are interested in the nature types that have been lost due to encroachment within recreation areas. Here we have to note though that the only 1,5% of the "særlig viktig natur" occur in the encroachment of outdoor recreation areas.

```{r encrec0056_analyses}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "ENCREC_003 - intersection"

# Check the CRS of both encrec and DT04
st_crs(encrec) == st_crs(natur_cln)

# Re-project natur_cln into ESPG : 25832
natur_32 <- st_transform(natur_cln, crs = st_crs(encrec))

# Check CRS of both datasets are the same now
st_crs(natur_32) == st_crs(encrec)

# Perform spatial intersection
encnatur <- encrec %>%
              select(fylkesnr,
                     kommune,
                     kartleggingsår,
                     områdeverdi,
                     id,
                     year,
                     area_inter_m2) %>%
              rename(area_encrec = area_inter_m2,
                     year_encrec = year) %>%
              st_intersection(., natur_32) %>%
              mutate(area_naturtyper_m2 = unclass(st_area(.)))

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "rcl_final",
                    "natur_32",
                    "encrec",
                    "encnatur")))

```

### 9.3 Uncertainty analyses
Our traffic-light uncertainty required to steps:

1. Attribute a traffic-light uncertainty at the polygon level. For this we used our lookup table developed in our [NINA Rapport XXXX](**<LINK>**); 

2. Derive a municipal / fylke traffic-light uncertainty based on the frequency of the traffic-light uncertainty at polygon-level (step 1 above). We basically created a frequency table of all the traffic-light color uncertainties from the polygons and derived the municipal / fylke traffic-light color uncertainty based the most predominant polygons' traffic-light colors. See[NINA Rapport XXXX](**<LINK>**) for more information.


First we attributed a colored-uncertainty to all polygons. `unc_0013_idv` and `unc_0024_idv`serve for both RECACC_001 and RECACC_003, and RECACC_002 and RECACC_004. This is because the same information is presented differently in both RECACC_001/REC_ACC_003 and RECACC_002/RECAA_004.

```{r recacc_uncertainty_1}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "Step 1 - attribute colored-uncertainty to all polygons"

# Load necessary functions
source(paste0(functions_path,"/add_uncertainty_individual.R"))

# Uncertainty RECACC_001
unc_0013_idv <- encrec %>%
                st_drop_geometry() %>%
                select(kommune, year, kartleggingsår) %>%
                mutate(uncertainty_source2 = apply(., 1, function(x) {add_uncertainty_individual(data = x, source2 = TRUE, yrsource2 = c("year", "kartleggingsår"))}))


unc_0024_idv <- encrec %>%
                st_drop_geometry() %>%
                    filter(områdeverdi %in% c("viktigFriluftslivsområde", 
                                              "sværtViktigFriluftslivsområde")) %>%
                select(kommune, year, kartleggingsår) %>%
                mutate(uncertainty_source2 = apply(., 1, function(x) {add_uncertainty_individual(data = x, source2 = TRUE, yrsource2 = c("year", "kartleggingsår"))}))



unc_0024_idv2 <- encrec %>%
                st_drop_geometry()  %>%
                    filter(områdeverdi %in% c("viktigFriluftslivsområde", 
                                              "sværtViktigFriluftslivsområde")) %>%
                group_by(kommune, year, kartleggingsår)  %>%
                summarise(n_poly = n()) %>%
                select(!n_poly) %>%
                ungroup() %>%
                mutate(uncertainty_source2 = apply(., 1, function(x) {add_uncertainty_individual(data = x, source2 = TRUE, yrsource2 = c("year", "kartleggingsår"))}))

# Uncertainty ENCACCC_005
unc_005_idv <- encnatur %>%
                st_drop_geometry() %>%
                select(kommune, year_encrec, year_natur_recorded, naturtype) %>%
                mutate(uncertainty_source2 = apply(., 1, function(x) {add_uncertainty_individual(data = x, source2 = TRUE, yrsource2 = c("year_encrec", "year_natur_recorded"))}))

# Uncertainty ENCACCC_006
unc_006_idv <- encnatur %>%
                st_drop_geometry() %>%
                filter(områdeverdi %in% c("viktigFriluftslivsområde", 
                                          "sværtViktigFriluftslivsområde")) %>%
                select(kommune, year_encrec, year_natur_recorded, naturtype) %>%
                mutate(uncertainty_source2 = apply(., 1, function(x) {add_uncertainty_individual(data = x, source2 = TRUE, yrsource2 = c("year_encrec", "year_natur_recorded"))}))


unc_006_idv2 <- encnatur %>%
                st_drop_geometry() %>%
                group_by(kommune, year_encrec, year_natur_recorded, naturtype) %>%
                summarise(n_poly = n()) %>%
                select(!n_poly) %>%
                ungroup() %>%
                mutate(uncertainty_source2 = apply(., 1, function(x) {add_uncertainty_individual(data = x, source2 = TRUE, yrsource2 = c("year_encrec", "year_natur_recorded"))}))


# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "rcl_final",
                    "natur_32",
                    "encrec",
                    "encnatur",
                    "unc_0013_idv",
                    "unc_0024_idv",
                    "unc_005_idv",
                    "unc_006_idv",
                    "unc_006_idv2",
                    "unc_0024_idv2")))

```

Now we can concatenate the traffic-light uncertainty from polygon level to municipal and fylke levels. As we haven't calculated the extent accounts yet, we won't derive the the traffic-light uncertainty at municipal / fylke level. We prefer to wait until we build the accounts as we need to add changes in extent as well. It is easier to do it in the end.

*Note: for "unc_005_sum" and "unc_006_sum", we only concatenate per habitat types as the accounts presnet a total extent of the "særlig viktig natur" lost over the total accounting period 2018-2024.*

```{r encacc_uncertainty_2}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "Pre-Step 2 - concatenate the polygin uncertainty from Step 1 at municipal / fylke levels"

# Call function to concatenate
source(paste0(functions_path,"/concatenate_individual_uncertainty.R"))

# Concatenate for RECACC_001 and RECACC_003
unc_0013_sum <- concatenate_fn(unc_0013_idv,
                              col_column_names = "uncertainty_source2", 
                              kom_num = "kommune",
                              accounting_year = "year")

# Concatenate for RECACC_002 and RECACC_004
unc_0024_sum <- concatenate_fn(unc_0024_idv,
                              col_column_names = "uncertainty_source2", 
                              kom_num = "kommune",
                              accounting_year = "year")

unc_0024_sum2 <- concatenate_fn(unc_0024_idv2,
                              col_column_names = "uncertainty_source2", 
                              kom_num = "kommune",
                              accounting_year = "year")

# Concatenate for RECACC_005
unc_005_sum <- concatenate_fn(unc_005_idv,
                              col_column_names = "uncertainty_source2", 
                              kom_num = "kommune",
                              habitat_col = "naturtype")

# Concatenate for RECACC_006
unc_006_sum <- concatenate_fn(unc_006_idv,
                              col_column_names = "uncertainty_source2",
                              kom_num = "kommune",
                              habitat_col = "naturtype")


unc_006_sum2 <- concatenate_fn(unc_006_idv2,
                              col_column_names = "uncertainty_source2",
                              kom_num = "kommune",
                              habitat_col = "naturtype")

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "rcl_final",
                    "natur_32",
                    "encrec",
                    "encnatur",
                    "unc_0013_idv",
                    "unc_0024_idv",
                    "unc_005_idv",
                    "unc_006_idv",
                    "unc_006_idv2",
                    "unc_0024_idv2",
                    "unc_0013_idv",
                    "unc_0024_idv",
                    "unc_0013_sum",
                    "unc_0024_sum",
                    "unc_005_sum",
                    "unc_006_sum",
                    "unc_0024_sum2",
                    "unc_006_sum2")))

```


<!--# 

Use this header for documenting the analyses. Put code in separate code chunks, and annotate the code in between using normal text (i.e. between the chunks, and try to avoid too many hashed out comments inside the code chunks). Add subheaders as needed. 

Code folding is activated, meaning the code will be hidden by default in the html (one can click to expand it).

Caching is also activated (from the top YAML), meaning that rendering to html will be quicker the second time you do it. This will create a folder inside you project folder (called INDICATORID_cache). Sometimes caching created problems because some operations are not rerun when they should be rerun. Try deleting the cash folder and try again.

-->

## 10. Results - Thematic Extent Accounts: encroachment in recreation areas
Now we can built the accounting tables and maps based on the spatial intersection we performed in the previous section. We chose to build the thematic accounts over the 2018-2024 period. Maps will be created only for the closing (2024) year of the total accounting period 2018-2024. **Only municipalities which have mapped their recreation areas are present in these accounts.**

All recreation areas have been mapped in different years ranging from 2010 to 2018. We decided to not take this difference in our analyses and the subsequent presentation of the accounts. This means that we consider the mapped recreation areas as homogeneous across the years. Our assumption for this is that the year of mapping does not indicate that the area mapped was not a recreation area before it was mapped. For example, we consider that recreation areas only mapped in 2025 should be part of the accounts as we assume that people would already use these areas to recreate in 2018, despite them not having been officially mapped in 2018 by municipalities. This assumption was used to <u> build the thematic accounts </u>. Nevertheless, we expect that a recreation areas mapped in 2010 might present some variations overtime and would present changes in its area in 2024. We took this possible temporal discrepancy into account in the uncertainty presented alongside the results.


### 10.1 Accounting tables
#### 10.1.1 ENCREC_001 and ENCREC_002: Extent accounts of encroachment in reacreation areas
Here we present the extent accounts table for encroachment areas between 2018 and 2024 for recreation areas within each municipality in Norway. We compile the extent of these encroachment areas in all recreation areas (ENCREC_001), but also only in recreation areas classified as important and very important (ENCREC_002). These accounts present both the yearly area of encroachment and the cumulative area of encroachment over the years. We also present a total across all municipalities in Norway which mapped their recreation areas.

Note that only the municipalities which have mapped their outdoors areas appear in the table below.

*Note: the function calculating the change compiles it as area_closing_year - area_opening_year. This is because we expect an increase in encroachment so we want to show how we gain in built-in areas.*
```{r res_acc_tables1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "ENCREC_001 and ENCREC_002- Create accounting table per municipality"

## Call function to add change in areas 2018-2024
source(paste0(functions_path, "/add_change_encroachment_area.R"))

## Call function to derive uncertainty at municipal / fylke level
source(paste0(functions_path, "/add_uncertainty_global.R"))

## Call function to concatenate
source(paste0(functions_path,"/concatenate_individual_uncertainty.R"))


## Import LAU dataset to then join the name of the municipalities
kom_navn <- lau_no %>%
              st_drop_geometry() %>%
              select(GISCO_ID,
                     LAU_NAME)

## ENCREC_001 and 002 accounting tables
# Accounts for all municipalities
encrec_001_kom <- encrec %>%
                    st_drop_geometry() %>%
                    select(fylkesnr,
                           kommune,
                           year,
                           area_inter_m2) %>%
                    group_by(fylkesnr, kommune, year) %>%
                    summarise(area_dekar_yearly = round2(sum(area_inter_m2)/1000, 2)) %>%
                    mutate(area_dekar_cumulative = round2(cumsum(area_dekar_yearly), 2)) %>%
                    left_join(., kom_navn, join_by("kommune" == "GISCO_ID")) %>%
                    relocate(c(LAU_NAME), .after = "kommune") %>%
                    mutate(indicator_id = "ENCREC_001", .before = "fylkesnr")

encrec_002_kom <- encrec %>%
                    st_drop_geometry() %>%
                    filter(områdeverdi %in% c("viktigFriluftslivsområde", 
                                              "sværtViktigFriluftslivsområde")) %>%
                    select(fylkesnr,
                           kommune,
                           year,
                           area_inter_m2) %>%
                    group_by(fylkesnr, kommune, year) %>%
                    summarise(area_dekar_yearly = round2(sum(area_inter_m2)/1000, 2)) %>%
                    mutate(area_dekar_cumulative = round2(cumsum(area_dekar_yearly), 2)) %>%
                    left_join(., kom_navn, join_by("kommune" == "GISCO_ID")) %>%
                    relocate(c(LAU_NAME), .after = "kommune") %>%
                    mutate(indicator_id = "ENCREC_002", .before = "fylkesnr")

# Add uncertainties to encacc_001_kom
encrec_001_kom_un <- encrec_001_kom %>%
                      left_join(., unc_0013_sum, by = c("kommune", "year")) %>%
                      add_uncertainty_global(., 
                                             uncertainty_col_name = "unc_concat_full", 
                                             kom_num = "kommune", 
                                             accounting_year = "year") %>%
                      select(!unc_concat_full)

encrec_002_kom_un <- encrec_002_kom %>%
                      left_join(., unc_0024_sum2, by = c("kommune", "year")) %>%
                      add_uncertainty_global(., 
                                             uncertainty_col_name = "unc_concat_full", 
                                             kom_num = "kommune", 
                                             accounting_year = "year") %>%
                      select(!unc_concat_full)


# Total across all municipalities
encrec_001_tot <- encrec_001_kom %>%
                    group_by(year) %>%
                    summarise(area_dekar_yearly = sum(area_dekar_yearly)) %>%
                    mutate(area_dekar_cumulative = round2(cumsum(area_dekar_yearly), 2)) %>%
                    mutate(indicator_id = "ENCREC_001",
                           fylkesnr = "00",
                           kommune = "0000",
                           LAU_NAME = "Total", .before = year)

encrec_002_tot <- encrec_002_kom %>%
                    group_by(year) %>%
                    summarise(area_dekar_yearly = sum(area_dekar_yearly)) %>%
                    mutate(area_dekar_cumulative = round2(cumsum(area_dekar_yearly), 2)) %>%
                    mutate(indicator_id = "ENCREC_002",
                           fylkesnr = "00",
                           kommune = "0000",
                           LAU_NAME = "Total", .before = year)
                     

# Add uncertainty to encacc_001_tot and encrec_002_tot
encrec_001_tot_un <-  unc_0013_idv %>%
                        mutate(national = "0000") %>%
                        concatenate_fn(.,
                                       col_column_names = "uncertainty_source2",
                                       kom_num = "national",
                                       accounting_year = "year") %>%
                        add_uncertainty_global(., 
                                               kom_num = "national",
                                               uncertainty_col_name = "unc_concat_full",
                                               accounting_year = "year") %>%
                      ungroup() %>%
                      select(!c(unc_concat_full, 
                                global_cumulative_uncertainty,
                                national,
                                year)) %>%
                      cbind(encrec_001_tot, .)

encrec_002_tot_un <-  unc_0024_idv %>%
                        mutate(national = "0000") %>%
                        concatenate_fn(.,
                                       col_column_names = "uncertainty_source2",
                                       kom_num = "national",
                                       accounting_year = "year") %>%
                        add_uncertainty_global(., 
                                               kom_num = "national",
                                               uncertainty_col_name = "unc_concat_full",
                                               accounting_year = "year") %>%
                      ungroup() %>%
                      select(!c(unc_concat_full, 
                                global_cumulative_uncertainty,
                                national,
                                year)) %>%
                      cbind(encrec_002_tot, .)


# Add change to all tables
encrec_001_kom_un$year <- as.character(encrec_001_kom_un$year)

encrec_001_kom_un <- encrec_001_kom_un %>%
                      group_split(kommune) %>% 
                      map(~ add_change_encroachment_area(., 
                                                         year_column = "year",
                                                         area1 = "area_dekar_cumulative")) %>% 
                      compact() %>%
                      bind_rows()

encrec_002_kom_un$year <- as.character(encrec_002_kom_un$year)

encrec_002_kom_un <- encrec_002_kom_un %>%
                      group_split(kommune) %>% 
                      map(~ add_change_encroachment_area(., 
                                                         year_column = "year",
                                                         area1 = "area_dekar_cumulative")) %>% 
                      compact() %>%
                      bind_rows()


encrec_001_tot_un$year <- as.character(encrec_001_tot_un$year)

encrec_001_tot_un <- add_change_encroachment_area(encrec_001_tot_un, 
                                                  year_column = "year",
                                                  area1 = "area_dekar_cumulative") %>% 
                      compact() %>%
                      bind_rows()


encrec_002_tot_un$year <- as.character(encrec_002_tot_un$year)

encrec_002_tot_un <- add_change_encroachment_area(encrec_002_tot_un, 
                                                  year_column = "year",
                                                  area1 = "area_dekar_cumulative") %>% 
                      compact() %>%
                      bind_rows()


# Correct uncertainty for global_yearly_uncertainty when accounting year is "Change 2018-2024"
encrec_001_kom_un[which(encrec_001_kom_un$year == "Change 2018-2024"), "global_yearly_uncertainty"] <- "NA"
encrec_002_kom_un[which(encrec_002_kom_un$year == "Change 2018-2024"), "global_yearly_uncertainty"] <- "NA"
encrec_001_tot_un[which(encrec_001_tot_un$year == "Change 2018-2024"), "global_yearly_uncertainty"] <- "NA"
encrec_002_tot_un[which(encrec_002_tot_un$year == "Change 2018-2024"), "global_yearly_uncertainty"] <- "NA"
                    
# Final accounting table               
encrec_0012_acc <- bind_rows(encrec_001_kom_un,
                             encrec_001_tot_un, 
                             encrec_002_kom_un,
                             encrec_002_tot_un)

encrec_0012_acc$fylkesnr[which(encrec_0012_acc$fylkesnr == "00")] <- "Total"
encrec_0012_acc$kommune[which(encrec_0012_acc$kommune == "0000")] <- "Total"

colnames(encrec_0012_acc) <- c("Indikatorens navn",
                               "Fylke nummer",
                               "Kommune nummer",
                               "Kommune navn",
                               "Regnskapsår",
                               "Årlig areal i dekar",
                               "Kumulativ areal i dekar",
                               "Usikkerhet for årlig areal",
                               "Usikkerhet for kumulativ areal")

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "rcl_final",
                    "natur_32",
                    "encrec",
                    "encnatur",
                    "unc_0013_idv",
                    "unc_0024_idv",
                    "unc_005_idv",
                    "unc_006_idv",
                    "unc_006_idv2",
                    "unc_0024_idv2",
                    "unc_0013_idv",
                    "unc_0024_idv",
                    "unc_0013_sum",
                    "unc_0024_sum",
                    "unc_005_sum",
                    "unc_006_sum",
                    "unc_0024_sum2",
                    "unc_006_sum2",
                    "kom_navn",
                    "add_change_encroachment_area",
                    "encrec_0012_acc")))


# Display the table dynamically
datatable(encrec_0012_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list( dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                          columnDefs = list(list(targets = c(which(colnames(encrec_0012_acc) == "Usikkerhet for årlig areal"),
                                                             which(colnames(encrec_0012_acc) == "Usikkerhet for kumulativ areal")),
                                                 visible = FALSE))),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Tabell 1 – Tematiske arealregnskap: areal av all nedbygging i all friluftslivomeråder i kommunner i perioden 2018–2024."))) %>%
    formatCurrency(
    columns = c("Årlig areal i dekar", "Kumulativ areal i dekar"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",") %>%
  ## Example: color Årlig areal i dekar by var1
  formatStyle(
    "Årlig areal i dekar",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red", "NA"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A", "")
    ),
    valueColumns = "Usikkerhet for årlig areal"           # column used for the intervals
  ) %>%
  ## Example: color Kumulativ areal i dekar by var2
  formatStyle(
    "Kumulativ areal i dekar",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A")
    ),
    valueColumns = "Usikkerhet for kumulativ areal"           # column used for the intervals
  )

```

#### 10.1.3 ENCREC_003 and ENCREC_004: Extent accounts of recreation areas
Here we present the extent accounts table for the total area of mapped recreation areas between 2018 and 2024 for all recreation areas within each municipality in Norway. We compile the extent of these area in all recreation areas (ENCREC_003), but also only in recreation areas classified as important and very important (ENCREC_004). These accounts present both the yearly area of encroachment and the cumulative area of encroachment over the years. We also present a total across all municipalities in Norway which mapped their recreation areas.

Note that only the municipalities which have mapped their outdoors areas appear in the table below.

*Note: the function calculating the change compiles it as area_opening_year - area_closing_year, as aligned with the System of Economic Environmental Accounting - Ecosystem Accounting Framework.*
```{r res_acc_tables3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "ENCREC_003 and ENCREC_004 - Create accounting table"

## Call function to add change in areas 2018-2024
source(paste0(functions_path, "/add_change_recreation_area.R"))

## Import LAU dataset to then join the name of the municipalities
kom_navn <- lau_no %>%
              st_drop_geometry() %>%
              select(GISCO_ID,
                     LAU_NAME)

## Compile the total area of recreation areas
rec_area_total <- rcl_32_tr %>%
                    st_drop_geometry() %>%
                    group_by(kommune) %>%
                    summarise(total_area_kom_dekar = round2(sum(area_rec_m2)/1000, 2))

recAB_area_total <- rcl_32_tr %>%
                    st_drop_geometry() %>%
                    filter(områdeverdi %in% c("viktigFriluftslivsområde", 
                                              "sværtViktigFriluftslivsområde")) %>%
                      group_by(kommune) %>%
                      summarise(total_area_kom_dekar = round2(sum(area_rec_m2)/1000, 2))
            

## ENCREC_003 and 004 accounting tables
# Accounts for all municipalities
encrec_003_kom <- encrec %>%
                    st_drop_geometry() %>%
                    select(fylkesnr,
                           kommune,
                           year,
                           area_inter_m2) %>%
                    group_by(fylkesnr, kommune, year) %>%
                    summarise(enc_dekar_yearly = round2(sum(area_inter_m2)/1000, 2)) %>%
                    mutate(enc_dekar_cumulative = round2(cumsum(enc_dekar_yearly), 2)) %>%
                    left_join(., rec_area_total, by = "kommune") %>%
                    mutate(area_dekar_yearly = total_area_kom_dekar - enc_dekar_yearly,
                           area_dekar_cumulative = total_area_kom_dekar - enc_dekar_cumulative) %>%
                    select(!c(enc_dekar_yearly, enc_dekar_cumulative, total_area_kom_dekar)) %>%
                    left_join(., kom_navn, join_by("kommune" == "GISCO_ID")) %>%
                    relocate(c(LAU_NAME), .after = "kommune") %>%
                    mutate(indicator_id = "ENCREC_003", .before = "fylkesnr")

encrec_004_kom <- encrec %>%
                    st_drop_geometry() %>%
                    filter(områdeverdi %in% c("viktigFriluftslivsområde", 
                                              "sværtViktigFriluftslivsområde")) %>%
                    select(fylkesnr,
                           kommune,
                           year,
                           area_inter_m2) %>%
                    group_by(fylkesnr, kommune, year) %>%
                    summarise(enc_dekar_yearly = round2(sum(area_inter_m2)/1000, 2)) %>%
                    mutate(enc_dekar_cumulative = round2(cumsum(enc_dekar_yearly), 2)) %>%
                    left_join(., recAB_area_total, by = "kommune") %>%
                    mutate(area_dekar_yearly = total_area_kom_dekar - enc_dekar_yearly,
                           area_dekar_cumulative = total_area_kom_dekar - enc_dekar_cumulative) %>%
                    select(!c(enc_dekar_yearly, enc_dekar_cumulative, total_area_kom_dekar)) %>%
                    left_join(., kom_navn, join_by("kommune" == "GISCO_ID")) %>%
                    relocate(c(LAU_NAME), .after = "kommune") %>%
                    mutate(indicator_id = "ENCREC_004", .before = "fylkesnr")


# Total across all municipalities
encrec_003_tot <- encrec_003_kom %>%
                    group_by(year) %>%
                    summarise(area_dekar_yearly = sum(area_dekar_yearly),
                              area_dekar_cumulative = round2(sum(area_dekar_cumulative), 2)) %>%
                    mutate(indicator_id = "ENCREC_003",
                           fylkesnr = "00",
                           kommune = "0000",
                           LAU_NAME = "Total", .before = year)

encrec_004_tot <- encrec_004_kom %>%
                    group_by(year) %>%
                    summarise(area_dekar_yearly = sum(area_dekar_yearly),
                              area_dekar_cumulative = round2(sum(area_dekar_cumulative), 2)) %>%
                    mutate(indicator_id = "ENCREC_004",
                           fylkesnr = "00",
                           kommune = "0000",
                           LAU_NAME = "Total", .before = year)
                    

# Add change to all tables
encrec_003_kom$year <- as.character(encrec_003_kom$year)

encrec_003_kom <- encrec_003_kom %>%
                      group_split(kommune) %>% 
                      map(~ add_change_encroachment_area(., 
                                                         year_column = "year",
                                                         area1 = "area_dekar_cumulative")) %>% 
                      compact() %>%
                      bind_rows()

encrec_004_kom$year <- as.character(encrec_004_kom$year)

encrec_004_kom <- encrec_004_kom %>%
                      group_split(kommune) %>% 
                      map(~ add_change_encroachment_area(., 
                                                         year_column = "year",
                                                         area1 = "area_dekar_cumulative")) %>% 
                      compact() %>%
                      bind_rows()


encrec_003_tot$year <- as.character(encrec_003_tot$year)

encrec_003_tot <- add_change_encroachment_area(encrec_003_tot, 
                                               year_column = "year",
                                               area1 = "area_dekar_cumulative") %>% 
                      compact() %>%
                      bind_rows()


encrec_004_tot$year <- as.character(encrec_004_tot$year)

encrec_004_tot <- add_change_encroachment_area(encrec_004_tot, 
                                               year_column = "year",
                                               area1 = "area_dekar_cumulative") %>% 
                      compact() %>%
                      bind_rows()

# Final accounting table               
encrec_0034_acc <- bind_rows(encrec_003_kom,
                             encrec_003_tot,
                             encrec_004_kom,
                             encrec_004_tot)

encrec_0034_acc$fylkesnr[which(encrec_0034_acc$fylkesnr == "00")] <- NA
encrec_0034_acc$kommune[which(encrec_0034_acc$kommune == "0000")] <- NA

colnames(encrec_0034_acc) <- c("Indikatorens navn",
                               "Fylke nummer",
                               "Kommune nummer",
                               "Kommune navn",
                               "Regnskapsår",
                               "Årlig areal i dekar",
                               "Kumulativ areal i dekar")

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "rcl_32_tr",
                    "rcl_final",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "add_change_recreation_area",
                    "rec_area_total",
                    "encrec_0012_acc",
                    "encrec_0034_acc")))


# Display the table dynamically
datatable(encrec_0034_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list( dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print')),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Table 2 - Arealregnskap av inngrep i alle friluftslivområder (ENCREC_003) eller i viktige og svært viktige friluftslivområder (ENCREC_004) over 2018-2024."))) %>%
    formatCurrency(
    columns = c("Årlig areal i dekar", "Kumulativ areal i dekar"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",")

```

#### 10.1.5 ENCREC_005 and ENCREC_006: Extent accounts of sværlig vikitig natur lost due to encroachments in recreation areas
Here we present the extent accounts table for the total area of "særlig viktig natur" lost due to encroachment in recreation areas between 2018 and 2024, within each municipality in Norway. We compile the loss of these area in all recreation areas (ENCREC_005), but also only in recreation areas classified as important and very important (ENCREC_006). As the loss of these nature types did not happen homogeneously within municipalities, these accounts only present both the total area lost by nature types over the 2018-2024 period. This is to avoid confusion and present the bigger picture, instead of having different years for different municipalities and nature types. We also present a total across all municipalities in Norway which mapped their recreation areas.

Note that only the municipalities which have mapped their outdoors areas and in which "særlig viktig natur" lost due to encroachment, appear in the table below.

```{r res_acc_tables56}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "ENCREC_005 and 006 - Create accounting table"


## ENCREC_005 and 006 accounting tables
# Accounts for all municipalities
encrec_005_kom <- encnatur %>%
                    st_drop_geometry() %>%
                    select(fylkesnr,
                           kommune,
                           naturtype,
                           year_encrec,
                           area_naturtyper_m2) %>%
                    group_by(fylkesnr, kommune, naturtype) %>%
                    summarise(total_nature_loss_dekar = round2(sum(area_naturtyper_m2)/1000, 2)) %>%
                    left_join(., kom_navn, join_by("kommune" == "GISCO_ID")) %>%
                    mutate(accounting_period = "2018-2024", .before = "total_nature_loss_dekar") %>%
                    relocate(c(LAU_NAME), .after = "kommune") %>%
                    mutate(indicator_id = "ENCREC_005", .before = "fylkesnr")

encrec_006_kom <- encnatur %>%
                    st_drop_geometry() %>%
                    filter(områdeverdi %in% c("viktigFriluftslivsområde", 
                                              "sværtViktigFriluftslivsområde")) %>%
                    select(fylkesnr,
                           kommune,
                           naturtype,
                           year_encrec,
                           area_naturtyper_m2) %>%
                    group_by(fylkesnr, kommune, naturtype) %>%
                    summarise(total_nature_loss_dekar = round2(sum(area_naturtyper_m2)/1000, 2)) %>%
                    left_join(., kom_navn, join_by("kommune" == "GISCO_ID")) %>%
                    mutate(accounting_period = "2018-2024", .before = "total_nature_loss_dekar") %>%
                    relocate(c(LAU_NAME), .after = "kommune") %>%
                    mutate(indicator_id = "ENCREC_006", .before = "fylkesnr")

# Total across all municipalities
encrec_005_tot <- encrec_005_kom %>%
                    group_by(naturtype) %>%
                    summarise(total_nature_loss_dekar = sum(total_nature_loss_dekar)) %>%
                    mutate(indicator_id = "ENCREC_005",
                           fylkesnr = "00",
                           kommune = "0000",
                           LAU_NAME = "Total", .before = naturtype) %>%
                    mutate(accounting_period = "2018-2024", .before = "total_nature_loss_dekar")

encrec_006_tot <- encrec_006_kom %>%
                    group_by(naturtype) %>%
                    summarise(total_nature_loss_dekar = sum(total_nature_loss_dekar)) %>%
                    mutate(indicator_id = "ENCREC_006",
                           fylkesnr = "00",
                           kommune = "0000",
                           LAU_NAME = "Total", .before = naturtype) %>%
                    mutate(accounting_period = "2018-2024", .before = "total_nature_loss_dekar")
                    
# Final accounting table               
encrec_0056_acc <- bind_rows(encrec_005_kom,
                             encrec_005_tot,
                             encrec_006_kom, 
                             encrec_006_tot)

encrec_0056_acc$fylkesnr[which(encrec_0056_acc$fylkesnr == "00")] <- NA
encrec_0056_acc$kommune[which(encrec_0056_acc$kommune == "0000")] <- NA

colnames(encrec_0056_acc) <- c("Indikatorens navn",
                               "Fylke nummer",
                               "Kommune nummer",
                               "Kommune navn",
                               "Natur typer",
                               "Regnskapsperiode",
                               "Areal særlig viktige naturtyper gått tapt (dekar)")

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "rcl_32_tr",
                    "rcl_final",
                    "encrec",
                    "encnatur",
                    "encrec_0012_acc",
                    "encrec_0033_acc",
                    "encrec_0056_acc")))


# Display the table dynamically
datatable(encrec_0056_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list(dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print')),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Table 3 - Arealregnskap av sværlig vikitig natur som går tapt på grunn av inngrep i alle friluftslivområder (ENCREC_005) eller i viktige og svært viktige friluftslivområder (ENCREC_006) over 2018-2024."))) %>%
    formatCurrency(
    columns = c("Areal særlig viktige naturtyper gått tapt (dekar)"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",")

```



### 10.2 Maps
#### 10.2.1 ENCREC_001: Map of total encroachment area in all recreation areas over 2018-2024
We first prepare the dataset for mapping then map it both using an interactive and static map. Due to memory issues in the html, we only display the static map.
```{r map_enc001_1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Prepare dataset for mapping"


## Import LAU dataset to then join the name of the municipalities
kom_navn <- lau_no %>%
              select(GISCO_ID,
                     LAU_NAME)

## Prepare dataset
encrec_001_map <- encrec_0012_acc %>%
                    filter(!c(`Kommune navn` == "Total" | Regnskapsår == "Change 2018-2024")) %>%
                    filter(Regnskapsår == 2024) %>%
                    filter(`Indikatorens navn` == "ENCREC_001") %>%
                    left_join(kom_navn, ., join_by("GISCO_ID" == "Kommune nummer")) %>%
                    select("Fylke nummer",
                           "Kommune navn",
                           "Kumulativ areal i dekar") %>%
                    rename("Kumulativ areal av inngrep i dekar" = "Kumulativ areal i dekar")

fylke_map <- encrec_001_map %>%
              group_by(`Fylke nummer`) %>%
              summarise("Kumulativ areal av inngrep i dekar" = sum(`Kumulativ areal av inngrep i dekar`))

```


```{r map_enc001_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Interactive map of total encroachment area in all recreation areas per municipality and fylke over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/interactive_map.R"))

# Run function
interactive_map(data1 = fylke_map, 
                data2 = encrec_001_map, 
                fill_variable1 = "Kumulativ areal av inngrep i dekar", 
                fill_variable2 =  "Kumulativ areal av inngrep i dekar", 
                legend_title1 = "Fylke - Areal av inngrep (dekar)", 
                legend_title2 = "Kommune - Areal av inngrep (dekar)", 
                map_title1 = "Totalt areal av inngrep (dekar) i friluftslivonsområder i fylker over 2018-2024", 
                map_title2 = "Totalt areal av inngrep (dekar) i friluftslivonsområder i kommuner over 2018-2024")

```

```{r map_enc001_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of total encroachment area in all recreation areas per municipality and fylke over 2018-2024"

## Static maps
# Call for the function to do static mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = fylke_map,
           data2 = encrec_001_map, 
           fill_variable1 = "Kumulativ areal av inngrep i dekar", 
           fill_variable2 =  "Kumulativ areal av inngrep i dekar", 
           legend_title1 = "Areal av inngrep (dekar)", 
           legend_title2 = "Areal av inngrep (dekar)",
           map_title1 = "Totalt areal av inngrep (dekar) \ni friluftslivonsområder \ni fylker i 2018-2024", 
           map_title2 = "Totalt areal av inngrep (dekar) \ni friluftslivonsområder \ni kommuner i 2018-\n2024")

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "rcl_32_tr",
                    "rcl_final",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "encrec_0012_acc",
                    "encrec_0034_acc",
                    "encrec_0056_acc")))

```

```{r map_enc001_4}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Clean R environment"

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "rcl_32_tr",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "encrec_0012_acc",
                    "encrec_0034_acc",
                    "encrec_0056_acc")))

```

#### 10.2.2 ENCREC_002: Map of total encroachment area in important and very important recreation areas over 2018-2024
We first prepare the dataset for mapping then map it both using an interactive and static map. Due to memory issues in the html, we only display the static map.
```{r map_enc002_1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Prepare dataset for mapping"

## Prepare dataset
encrec_002_map <- encrec_0012_acc %>%
                    filter(!c(`Kommune navn` == "Total" | Regnskapsår == "Change 2018-2024")) %>%
                    filter(Regnskapsår == 2024) %>%
                    filter(`Indikatorens navn` == "ENCREC_002") %>%
                    left_join(kom_navn, ., join_by("GISCO_ID" == "Kommune nummer")) %>%
                    select("Fylke nummer",
                           "Kommune navn",
                           "Kumulativ areal i dekar") %>%
                    rename("Kumulativ areal av inngrep i dekar" = "Kumulativ areal i dekar")

fylke_map <- encrec_002_map %>%
              group_by(`Fylke nummer`) %>%
              summarise("Kumulativ areal av inngrep i dekar" = sum(`Kumulativ areal av inngrep i dekar`))
```

```{r map_enc002_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Interactive map of total encroachment area in important and very important recreation areas per municipality and fylke over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/interactive_map.R"))

# Run function
interactive_map(data1 = fylke_map, 
                data2 = encrec_002_map, 
                fill_variable1 = "Kumulativ areal av inngrep i dekar", 
                fill_variable2 =  "Kumulativ areal av inngrep i dekar", 
                legend_title1 = "Fylke - Areal av inngrep (dekar)", 
                legend_title2 = "Kommune - Areal av inngrep (dekar)", 
                map_title1 = "Totalt areal av inngrep (dekar) i svært vikktig og viktig friluftslivonsområder i fylker over 2018-2024", 
                map_title2 = "Totalt areal av inngrep (dekar) i svært vikktig og viktig friluftslivonsområder i kommuner over 2018-2024")
```

```{r map_enc002_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of total encroachment area in important and very important recreation areas per municipality and fylke over 2018-2024"

## Static maps
# Call for the function to do static mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = fylke_map,
           data2 = encrec_002_map, 
           fill_variable1 = "Kumulativ areal av inngrep i dekar", 
           fill_variable2 =  "Kumulativ areal av inngrep i dekar", 
           legend_title1 = "Areal av inngrep (dekar)", 
           legend_title2 = "Areal av inngrep (dekar)",
           map_title1 = "Totalt areal av inngrep (dekar) \ni friluftslivonsområder \ni fylker i 2018-2024", 
           map_title2 = "Totalt areal av inngrep (dekar) \ni friluftslivonsområder \ni kommuner i 2018-\n2024")

```

```{r map_enc002_4}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Clean R environment"

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "rcl_32_tr",
                    "rcl_final",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "encrec_0012_acc",
                    "encrec_0034_acc",
                    "encrec_0056_acc")))

```

#### 10.2.3 Map of the share of the total recreation area lost to encroachment over 2018-2024
We first prepare the dataset for mapping then map it both using an interactive and static map. Due to memory issues in the html, we only display the static map.
```{r map_enc001_perc_1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Prepare data for mapping"


## Prepare dataset
rec_total_kom <- rcl_32_tr %>%
                  st_drop_geometry() %>%
                  mutate(area_dekar = round2(area_rec_m2/1000), 2) %>%
                  group_by(fylkesnr, kommune) %>%
                  summarise(total_rec_area = sum(area_dekar))

encrec_001_perc_map <- encrec_0012_acc %>%
                        filter(!c(`Kommune navn` == "Total" | Regnskapsår == "Change 2018-2024")) %>%
                        filter(Regnskapsår == 2024) %>%
                        filter(`Indikatorens navn` == "ENCREC_001") %>%
                        left_join(kom_navn, ., join_by("GISCO_ID" == "Kommune nummer")) %>%
                        left_join(., rec_total_kom, join_by("GISCO_ID" == "kommune")) %>%
                        mutate("Share (%) of recreation area lost" = round2(`Kumulativ areal i dekar`/total_rec_area * 100, 2)) %>%
                        select("Fylke nummer",
                               "Kommune navn",
                               "Share (%) of recreation area lost",
                               "total_rec_area",
                               "Kumulativ areal i dekar")

fylke_map <- encrec_001_perc_map %>%
              group_by(`Fylke nummer`) %>%
              summarise(fylke_rec_area = sum(total_rec_area),
                        "Kumulativ areal i dekar" = sum(`Kumulativ areal i dekar`),
                        "Share (%) of recreation area lost" = round2(`Kumulativ areal i dekar`/fylke_rec_area * 100, 2))
```

```{r map_enc001_perc_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Interactive map of the share of recreation area lost to encroachments in all recreation areas per municipality and fylke over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/interactive_map.R"))

# Run function
interactive_map(data1 = fylke_map, 
                data2 = encrec_001_perc_map, 
                fill_variable1 = "Share (%) of recreation area lost", 
                fill_variable2 =  "Share (%) of recreation area lost", 
                legend_title1 = "Fylke - Prosentandel (%)", 
                legend_title2 = "Kommune - Prosentandel (%)", 
                map_title1 = "Prosentandel (%) av totale friluftsområder som er gått tapt i fylker", 
                map_title2 = "Prosentandel (%) av totale friluftsområder som er gått tapt i kommuner")

```

```{r map_enc001_perc_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of the share of recreation area lost to encroachments in all recreation areas per municipality and fylke over 2018-2024"

## Static maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = fylke_map, 
           data2 = encrec_001_perc_map, 
           fill_variable1 = "Share (%) of recreation area lost", 
           fill_variable2 =  "Share (%) of recreation area lost", 
           legend_title1 = "Prosentandel (%)", 
           legend_title2 = "Prosentandel (%)", 
           map_title1 = "Prosentandel (%) av totale \n friluftsområder som \ner gått tapt \ni fylker over \n2018-2024", 
           map_title2 = "Prosentandel (%) av totale \n friluftsområder som \ner gått tapt \ni kommuner over \n2018-2024")


```

```{r map_enc003_4}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Clean R environment"

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "rcl_32_tr",
                    "rcl_final",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "encrec_0012_acc",
                    "encrec_0034_acc",
                    "encrec_0056_acc")))

```

#### 10.2.4 Map of the share of the total important and very important recreation area lost to encroachment over 2018-2024
We first prepare the dataset for mapping then map it both using an interactive and static map. Due to memory issues in the html, we only display the static map.
```{r map_enc002_perc_1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Prepare data for mapping"

## Prepare dataset
rec_total_kom <- rcl_32_tr %>%
                  st_drop_geometry() %>%
                  filter(områdeverdi %in% c("viktigFriluftslivsområde", 
                                              "sværtViktigFriluftslivsområde")) %>%
                  mutate(area_dekar = round2(area_rec_m2/1000), 2) %>%
                  group_by(fylkesnr, kommune) %>%
                  summarise(total_rec_area = sum(area_dekar))

encrec_002_perc_map <- encrec_0012_acc %>%
                        filter(!c(`Kommune navn` == "Total" | Regnskapsår == "Change 2018-2024")) %>%
                        filter(Regnskapsår == 2024) %>%
                        filter(`Indikatorens navn` == "ENCREC_002") %>%
                        left_join(kom_navn, ., join_by("GISCO_ID" == "Kommune nummer")) %>%
                        left_join(., rec_total_kom, join_by("GISCO_ID" == "kommune")) %>%
                        mutate("Share (%) of recreation area lost" = round2(`Kumulativ areal i dekar`/total_rec_area * 100, 2)) %>%
                        select("Fylke nummer",
                               "Kommune navn",
                               "Kumulativ areal i dekar",
                               "total_rec_area",
                               "Share (%) of recreation area lost")

fylke_map <- encrec_002_perc_map %>%
              group_by(`Fylke nummer`) %>%
              summarise(fylke_rec_area = sum(total_rec_area),
                        "Kumulativ areal i dekar" = sum(`Kumulativ areal i dekar`),
                        "Share (%) of recreation area lost" = round2(`Kumulativ areal i dekar`/fylke_rec_area * 100, 2))
                   
```

```{r map_enc002_perc_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Interactive map of the share of recreation area lost to encroachments in important and very important recreation areas per municipality and fylke over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/interactive_map.R"))

# Run function
interactive_map(data1 = fylke_map, 
                data2 = encrec_002_perc_map, 
                fill_variable1 = "Share (%) of recreation area lost", 
                fill_variable2 =  "Share (%) of recreation area lost", 
                legend_title1 = "Fylke - Prosentandel (%)", 
                legend_title2 = "Kommune -Prosentandel (%)", 
                map_title1 = "Prosentandel (%) av totale svært viktig og viktig friluftsområder som er gått tapt i fylker over 2018-2024", 
                map_title2 = "Prosentandel (%) av totale svært viktig og viktig friluftsområder som er gått tapt i kommuner over 2018-2024")
```

```{r map_enc002_perc_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of the share of recreation area lost to encroachments in important and very important recreation areas per municipality and fylke over 2018-2024"

## Static maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = fylke_map, 
           data2 = encrec_002_perc_map, 
           fill_variable1 = "Share (%) of recreation area lost", 
           fill_variable2 =  "Share (%) of recreation area lost", 
           legend_title1 = "Prosentandel (%)", 
           legend_title2 = "Prosentandel (%)", 
           map_title1 = "Prosentandel (%) av totale \nsvært viktig og viktig \nfriluftsområder \nsom er gått tapt \ni fylker over \n2018-2024", 
           map_title2 = "Prosentandel (%) av totale \nsvært viktig og viktig \nfriluftsområder \nsom er gått tapt \ni kommuner over \n2018-2024")



```

```{r map_enc004_4}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Clean R environment"

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "rcl_32_tr",
                    "rcl_final",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "encrec_0012_acc",
                    "encrec_0034_acc",
                    "encrec_0056_acc")))

```

#### 10.2.4 ENCREC_006 map of total "sæarlig viktig natur" due to encroachment in important and very important recreation areas over 2018-2024
We first prepare the dataset for mapping then map it both using an interactive and static map. Due to memory issues in the html, we only display the static map. As very few municipalities have "særlig viktig natur" in their encroachments of recreation areas, we do not provide summary statistics at Fylke Level.

```{r map_enc006_1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Prepare dataset for mapping"

## Import LAU dataset to then join the name of the municipalities
kom_navn <- lau_no %>%
              select(GISCO_ID,
                     LAU_NAME)

## Prepare dataset
encrec_006_map <- encrec_0056_acc %>%
                    filter(`Indikatorens navn` == "ENCREC_006") %>%
                    filter(!`Kommune navn` == "Total") %>%
                    group_by(`Fylke nummer`, `Kommune nummer`, `Kommune navn`) %>%
                    summarise("Areal særlig viktige naturtyper gått tapt (dekar)" = sum(`Areal særlig viktige naturtyper gått tapt (dekar)`)) %>%
                    left_join(kom_navn, ., join_by("GISCO_ID" == "Kommune nummer")) %>%
                    select("Fylke nummer",
                           "Kommune navn",
                           "Areal særlig viktige naturtyper gått tapt (dekar)")

```

```{r map_enc006_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Interactive map of total ''sæarlig viktig natur'' in important and very important recreation areas per municipality and fylke over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/interactive_map.R"))

# Run function
interactive_map(data1 = encrec_006_map,
                fill_variable1 = "Areal særlig viktige naturtyper gått tapt (dekar)",
                legend_title1 = "Areal særlig viktige naturtyper gått tapt (dekar)",
                map_title1 = "Areal særlig viktige naturtyper gått tapt (dekar) i svært viktige og viktige friluftsområder i kommuner over 2018-2024",
                style_type = "fixed",
                interval_breaks = c(0, 2, 15, 30, 50, 225))

```

```{r map_enc006_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of total ''sæarlig viktig natur'' in important and very important recreation areas per municipality and fylke over 2018-2024"

## Static maps
# Call for the function to do static mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = encrec_006_map,
           fill_variable1 = "Areal særlig viktige naturtyper gått tapt (dekar)",
           legend_title1 = "Areal særlig viktige naturtyper gått tapt (dekar)",
           map_title1 = "Areal særlig viktige naturtyper \ngått tapt (dekar) i svært \nviktige og viktige \nfriluftsområder i \nkommuner \nover 2018-\n2024",
           style_type = "fixed",
           interval_breaks = c(0, 2, 15, 30, 50, 225))

## Clean R environment
rm(list = setdiff(ls(),
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "rcl_32_tr",
                    "rcl_final",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "encrec_0021_acc",
                    "encrec_0034_acc",
                    "encrec_0056_acc")))

```


<!--# 

Repeat the final results here. Typically this is a map or table of indicator values.

This is typically where people will harvest data from, so make sure to include all relevant output here, but don't clutter this section with too much output either.

-->

## 11. Export file

<!--# 

Optional: Display the code (don't execute it) or the workflow for exporting the indicator values to file. Ideally the indicator values are exported as a georeferenced shape or raster file with indicators values, reference values and errors. You can also chose to export the raw (un-normalised or unscaled variable) as a separate product. You should not save large sptaial output data on GitHub. You can use eval=FALSE to avoid code from being executed (example below - delete if not relevant) 

-->

```{r export}
#| eval: false
```

<!-- You can leave this last part as it is, unless you have a good reason to alter it. -->

::: {.callout-tip collapse="true"}
## Session Info

This information is mainly aimed at developers that need to recreate the R environment that produced the html.
Important note: this information is about the R session that was running when the html was rendered, and if some of the code chunks were not evaluated when the html was rendered, they may have been executed in a different environment.  

```{r sessionInfo}
#| echo: false
sessionInfo()
```

:::
