---
title: "Thematic accounts - encroachment in recreation areas"
subtitle: |
  [ENCREC_001] (ENCroachments in all RECreation areas) \
  [ENCREC_002] (ENCroachments in very important (A) and important (B) RECreation areas)  
# delete the line with the second indicatorID (line 5) 
# if it's not relevant for you.
# In the parentheses you can add a word to distinguish multiple indicators 
# from each other, for example the ecosystem type that they cover
indicatorID: 
  - ENCREC_001
  - ENCREC_002 # delete whole line if not needed
# If you are presenting multiple indicators, and have multiple 
# indicator IDs, then add them as a list here, and make sure you add
# the same number of list entries, 
# and in the same order that corresponds to the indicator IDs,  
# to the meta data fields where the indicators differ. 
# For example, if the indicators cover different ecosystem types, 
# add the ecosystems as a list under 'Ecosystems', but if they are in the
# same realm, same country etc., then you only need to write the realm and
# country once. 
indicatorName: 
  - ENCroachments in all RECreation areas # add more by adding another line starting '-'
  - ENCroachments in very important (A) and important (B) RECreation areas
folderName: code # name of the folder under 'indicators/'. 
# If it's a single indicator being reported, the folder name should be the same as the indicatorID. 
# If there are multiple ID's then the folder name could look something like XX_YYY_ZZZ-ÅÅÅ.
country: Norway
continent: Europe
# ECT: write ECT here # e.g. A2 - Chemical State characteristics
Realm: 
  - write realm here # e.g. Freshwater (F)
Biome: 
  - write biome here # e.g. F1 Rivers and streams biome
Ecosystem: 
  - write ecosystem here # e.g. F1.3 Freeze-thaw streams
yearAdded: 2025 # the year when the indicator was first submitted to ecRxiv
yearLastUpdate: 2025
status: incomplete # options are complete, incomplete, or deprecated
Version: "000.001" # Beta-versions are numbers incrementally with values below 0. Operational versions can be given the version numbers above 0. Use this terms similar as for software versioning,
VersionComment: This is the first version of the thematic accounts on recreation encroachment being compiled # Describe briefly how this version differs from earlier versions 
url: https://github.com/NINAnor/recereation_encroachment_accounts
# Leave the url as is, but replace XX_YYYY_ZZ with the name of the root folder for the current indicator(s). It should be the same as the indicatorID, but if there are more indicator IDs presented in the same folder, the folder name might be different, e.g. XX_YYYY_AAA-BBB 
Normalised: No # Yes or no depending on whether the metric has been normliased or not.
SpatialAggregationPathway: Describe the aggregation using the standrdaise terminology # e.g. Scale - Transform - Aggregate (length-weighted arithmetic mean) - Truncate
format: 
  html:
    embed-resources: true
    code-fold: true
    toc: true
    toc-title: Contents
    toc-depth: 3
    smooth-scroll: true
execute: 
  cache: true
author:
  - name: Sylvie Clappe             # Enter name
    email: sylvie.clappe@nina.no  # Enter email
    affiliations:
      - id: myID
        name: Norwegian Institute for Nature Research  # Enter affiliations
  - name: David Barton              #  Enter subsequent authors like this, or remove if not relevant
    affiliations:
      - ref: myID               # To reuse affiliations reference the id like this
  - name: Reto Spielhofer             #  Enter subsequent authors like this, or remove if not relevant
    affiliations:
  - ref: myID 
AuthorList: Clappe, S., Barton, D.N., Spielhofer, R.  
date: January 7, 2026 # Enter date
data_availability: none
code_reproducibility: none
open_science_badge: none
callout-icon: false
lightbox: true
css: ../../../style.css
code-links:
      - text: Add a review
        icon: github
        href: https://github.com/NINAnor/ecRxiv
hide: TRUE # Change to FALSE
---

<!--# This is a template for how to document the indicator analyses. Make sure also to not change the order, or modify, the headers, unless you really need to. This is because it easier to read if all the indicators are presented using the same layout. If there is one header where you don't have anything to write, just leave the header as is, and don't write anything below it. If you are providing code, be careful to annotate and comment on every step in the analysis. Before starting it is recommended to fill in as much as you can in the metadata file. This file will populate the initial table in your output.-->

<!--# Load all you dependencies here -->

```{r setup}
#| include: false
#| cache: false
library(knitr)
library(tidyverse)
library(kableExtra)
library(here)
library(anybadger)
library(yaml)
library(tibble)
library(conflicted)
library(httr)
library(jsonlite)
library(sf)
# weird workarounds for cache issues
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::lag)
conflicts_prefer(dplyr::select)
pull<-dplyr::pull

knitr::opts_chunk$set(echo = TRUE)
```

```{r source}
#| echo: false
#| cache: false
source(here::here("_common.R"))
```

<!--# The following parts are autogenerated. Do not edit. -->

::: {layout-ncol="5"}

```{r statusBadge}
#| echo: false
status_badge(as.character(st))
```

```{r versionBadge}
#| echo: false
version_badge(my_version_number = as.character(version[[1]]),
  folder_name = folder_name)
```

```{r dataBadge}
#| echo: false
data_badge(as.character(badges[[1]]))
```

```{r codeBadge}
#| echo: false
code_badge(as.character(badges[[2]]))
```

```{r openScienceBadge}
#| echo: false
open_science_badge(as.character(badges[[3]]))
```
::: 
 

> **Recommended citation**: `r paste(auth, " ", year, ". ", name, " (ID(s): ", folder_name, ") ", "v. ", version, ". ", url, sep="")`



::: {.callout-tip collapse="true"}
## Metadata
```{r tbl-meta2}
#| tbl-cap: 'Indicator metadata'
#| echo: false
#| warning: false
meta |>
  dplyr::filter(Variable %in% c(
    "Indicator ID",
    "Indicator Name",
    "Continent",
    "Country",
    "Ecosystem Condition Typology Class", 
    "Realm",
    "Biome",
    "Ecosystem",
    "Year added", 
    "Last update",
    "Version",
    "Version comment",
    "Normalised",
    "Spatial aggregation pathway")) |>   
  kable()

```
:::

::: {.callout-tip collapse="true"}

## Log

<!--# Update this logg with short messages for each update -->
- 07 January 2026 - First draft version
:::


<hr />

<!--# Document you work below.  -->

## 1. Summary

<!--# 
With a maximum of 300 words, describe the indicator in general terms as you would to a non-expert. Think of this as a kind of common language summary. Follow the four numbers sub-headers below. under point 2, it is possible under  include a bullet point list of the specific steps in the workflow. 
 -->

1. **Background and rationale**:
<!-- What does the metric represent? Why is this relevant for describing ecosystem condition in this ecosystem?
What are the main anthropogenic impact factors? -->
2. **Methods**:
<!-- What kind of data is used? How is the data customized (modified, estimated, integrated) to fit its purpose as an indicator? -->
3. **Interpretation**:
<!-- Add a very short description of how the indicator values should be interpreted. Here's an example for a bumblebee indicator: An indicator value of 1 signifies an intact bumblebee community. Values decreasing toward zero reflect negative ecological change, specifically the decline of species expected to be present. This metric does not account for potentially positive changes, such as increases in species' abundance or the arrival of new species. Tip: Add text on the line directly following the subheader, and don't introduce blank lines. This way the whole paragraph renders as a block quote. -->
An indicator value of 1 signifies ...
Values decreasing toward zero reflect...

4. **Results and development needs**:
<!-- Briefly, what does the indicator say about the ecosystem condition. What is the current status of  the metric (can it be used or is it still in development), and what immediate tasks remain to improve the indicator or to make the indicator operational -->



## 2. About the underlying data

<!--# Give a general introduction the datasets you have used, and provide more specific details in the sub-chapters.  -->

### 2.1 Spatial and temporal resolution and extent

<!--# Describe the temporal and spatial resolution and extent of the data used -->

### 2.2 Sampling design

<!-- How was the data sampled? Describe potential sampling biases -->

### 2.3 Original units

<!--# What are the original units for the most relevant  variables in the data-->


### 2.3 Instructions for citing, using and accessing data

<!--# Is the data openly available? If not, how can one access it? What are the key references to the datasets?   -->

### 2.4 Additional comments about the dataset

<!--# Optional. Text here -->

## 3. Indicator properties

### 3.1 Collinearities with other indicators

<!--# Describe known collinearities with other metrics (indicators or variables) that could become problematic if they were also included in the same Ecosystem Condition Assessment as the indicator described here. -->


### 3.2 Impact factors

<!--# Describe the main natural and anthropogenic factors that affecst the metric -->

### 3.3 Spatial aggregation

<!-- Describe in word the same aggregation pathway presented in the top yaml (the metadata). If you have scaled against the reference value, it is especially important to know if you did any avergaing or summations etc. beforehand. An example text for a rather simple case could be: "The variable is scaled at the initial polygon level and the point estimate is a area-weighted arithmetic mean of all polygins. There is no transformation, and truncation is not needed as the indicator becomes naturally bound between 0 and 1".  -->


## 5. Uncertainties
### 5.1 General uncertainties
<!--# Describe the main reasons why you might not trust 100% the final value to represent the given ecosystem characteristic for the give ecosystem type. What aspects of this condition metric are most sensitive to subjective decitions? Is the spatial sampling sufficient? Does the choice of modeling framework affect the values? Is the metric relevant do describe ecosystem condition for the give ecosystem type?  -->

### 5.2 Reported uncertainties
<!-- Describe the nature of the errors you (hopefully) provide alongside the point estimate for the condition metric. It is especially important to clarify if the errors are inferential (referring to the uncertainty of the point estimate) or descriptive (referreing to the data itself). Here' an example text: "What we report as uncertainties in @tbl-finalTable are inferential and refer to the uncertainty around the point estimates. It does not directly describe variation in the data itself. 
However, the variation in the mean is defined by the spatial variation only, and quartiles are obtained by bootstrapping the set of localities within a region, and calculating the indicator for each of 9999 such samples".   -->



## 6. References

<!--# You can add references manually or use a citation manager and add intext citations as with crossreferencing and hyperlinks. See https://quarto.org/docs/authoring/footnotes-and-citations.html -->

## 7. Import and prepare datasets
In this section, we show how to import the data through API. We also prepare them  for analyses. We first set up the path that will be used during th eproject. Users can change it to match their own path.
```{r path_setup1}
#| eval: true
#| code-summary: "Path setup"

# Path to the project folder
project_path <- getwd()

# Path to the functions folder within the project folder
functions_path <- paste0(project_path, "/src/functions")
```

<!--# All the data import and preparation should be done in this chapter, before the analysis chapter. As a minimum, add code for importing data, and describe or show the data structure. You may also explain certain variables, what they represent, and their distributions -->

### 7.1 Recreation areas - DT01
#### 7.1.1 Download DT01 for Geonorge through an API
We download DT01 through an API from [Geonorge](https://kartkatalog.geonorge.no/metadata/friluftslivsomraader-kartlagte/91e31bb7-356f-4478-bcba-d5c2de6e91bc?search=friluftsliv). Downloading through API in Geonorge is not an easy process, hence we fully decompose the API download request to facilitate the download of other datasets by users if they wish to. Note however that datasets not publicly accessible cannot be downloaded via an API.

1. **Create a  <u>base url<u>**. The *base url* is the basic url that needs to be used when downloading data. It will be the same for all datasets. In Geonorge, users should go the their datsaset of interest and scroll to the bottom of the page to the section *Distribusjon*. There, there are several subsections called *Distribusjonstype*. They display the *base url* depending on where the users want to download the data from. In our case, we chose to download our data from Geonorge whose *base url* is https://geoapi.p.aks.miljodirektoratet.no/geonorge. Here we removed the last part of the url (/capabilities) as it refers to another concept explained in the point 3 below.

```{r import_rec_areas_api1}
#| eval: true
#| code-summary: "DT01 API download - base url"

# base url for downloading dataset in Geonorge through API
base_url <- "https://geoapi.p.aks.miljodirektoratet.no/geonorge"
```


2. **Identify the <u>dataset ID </u>**. The dataset ID is called an uuid. It can be found by going to the dataset's page on Geonorge, it is the long series of numbers and letters in the url of the page: ".../metadata/dataset_name/uuid"
```{r import_rec_areas_api2}
#| eval: true
#| code-summary: "DT01 API download - dataset ID"

# dataset specific ID from Geon
dataset_uuid <- "91e31bb7-356f-4478-bcba-d5c2de6e91bc"
```

3. **Request the dataset's <u>capabilities</u>**. When downloading data through an API, users have to choose what is called "capabilities". These refer to the settings that should be used when downloading the data such as geographic areas, raster or vector type, format etc. We need to first build an API request to know what these capabilities are before downloading the data.
```{r import_rec_areas_api3}
#| eval: true
#| code-summary: "DT01 API download - list capabilities"

# httr:::GET() sends API request for dataset's capabilities
dt_cap <- GET(
  url = sprintf("%s/capabilities/%s", base_url, dataset_uuid), # builds all the url
  accept_json() # signals the server we request JSON format
)

# check if the server's response is an error
stop_for_status(dt_cap)

# Read the JSON response of the server listing the capabilities
cap_all <- fromJSON(content(dt_cap, as = "text", encoding = "UTF-8"), simplifyVector = TRUE)
```


4. **Choose the dataset's <u>capabilities</u> of relevance**. Now that we have the list of all the dataset's capabilities, we can choose those we want to download the dataset with. We are interested in three capabilities: projection, format, area and order. *Note: (i) order is the url to be used to order the dataset to the server; (ii) in the code below, the "rel" column just indicates what the link "href" refers to, meaning project, format, area, order etc. The url to use are in the "href" column; (iii) four projections can be chosen but in reality, only two are available for Norway as a country: ESPG:25833 and WGS84. Trying the others codes will lead to an error.*
```{r import_rec_areas_api4}
#| eval: true
#| code-summary: "DT01 API download - choose capabilities"

# Extract the capablities into a dataframe so it's easier to choose from
projection <- GET(cap_all$`_links`[cap_all$`_links`$rel == "http://rel.geonorge.no/download/projection", "href"]) %>%
                content(., as = "text", encoding = "UTF-8") %>%
                fromJSON(.)

format <- GET(cap_all$`_links`[cap_all$`_links`$rel == "http://rel.geonorge.no/download/format", "href"]) %>%
                content(., as = "text", encoding = "UTF-8") %>%
                fromJSON(.)

area <- GET(cap_all$`_links`[cap_all$`_links`$rel == "http://rel.geonorge.no/download/area", "href"]) %>%
                content(., as = "text", encoding = "UTF-8") %>%
                fromJSON(.)

# Choose the code of interest for each capabilities
choice_proj <- projection[2,] # ESPG:25833

choice_format <- format$name[2] # FILEGDB

choice_area <- area[1,] # Norway with Svalbard

order_url <- cap_all$`_links`[cap_all$`_links`$rel == "http://rel.geonorge.no/download/order", "href"]

```


5. **Create an <u>order</u>**. Here we create the API request to order the data from the server. In our case, the server responds with a downloadable link that we can then use. Every server has its own JSON request structure, it is important to go on the [API help page of Geonorge](https://nedlasting.geonorge.no/Help) to use the correct structure. As we use a POST API order, we need to go to the [POST API oder page](https://nedlasting.geonorge.no/Help/Api/POST-api-order) to figure it out.
```{r import_rec_areas_api5}
#| eval: true
#| code-summary: "DT01 API download - order the data"

# Create the order by listing the dataset id and capabilities of relevance
order_body <- list(
  downloadAsBundle = FALSE,
  email = "sylvie.clappe@nina.no",
  orderLines = list(
    list(
      areas = list(
        list(
          code = choice_area$code,  
          name = choice_area$name,
          type = choice_area$type
        )
      ),
      formats = list(
        list(name = choice_format)
      ),
      metadataUuid = dataset_uuid,
      projections = list(
        list(
          code = choice_proj$code,
          name = choice_proj$name,
          codespace = choice_proj$codespace
        )
      )
    )
  )
)

# Send the order to the server
order_req <- POST(url = order_url,
                  body = order_body,
                  encode = "json",
                  accept_json())

# Check if the server returns an error
stop_for_status(order_req)


# Read the order information
order_info <- fromJSON(content(order_req, as = "text", encoding = "UTF-8"))

```

6. **Download the data**. Now that we have the downloadable link, we just have the use it to download the dataset in R. *Note: we download the data in a temporary file which can then be deleted.*
```{r import_rec_areas_api6}
#| eval: true
#| code-summary: "DT01 API download - download and read the data"

# URL to dowload the data
down_url <- order_info$files$downloadUrl

# Create a temporary directroy to store the dataset .zip and unzipped
temp_dir <- tempdir()

# Create a temporary file to store the dataset .zip
destination_tmp <- tempfile(fileext = ".zip", tmpdir = temp_dir)

# Download the .zip file containing the dataset
download.file(down_url, destfile = destination_tmp, mode = "wb")

# Unzip the .zip
unzip(destination_tmp, exdir = temp_dir)

# Use "\\.gdb$" pattern to identify the file to read
gdb_file <- list.files(temp_dir, pattern = "\\.gdb$", full.names = TRUE)

```


#### 7.1.2 Read DT01
Now that the dataset has been downloaded in a temporary file, we can read it in R and clean the R environment. Once the analyses are finished, we will remove the temporary file containing the data.
```{r read_rec_area}
#| eval: true
#| code-summary: "DT01 - read the data"

# Read data
rec_no <- st_read(gdb_file)

# Clean R environment
rm(list = setdiff(ls(), 
                  "rec_no"))
```

#### 7.1.3 Clean and format the data for analyses
We first need to select the column of relevance for our work and clean the topological and geometrical errors if any.

```{r clean_rec_area}
#| eval: true
#| code-summary: "DT01 - clean and format"

# Select columns of relevance
rec_areas <- rec_no %>%
              select(fylkesnr,
                     kommune,
                     kartleggingsår,
                     kartlagtStatus,
                     identifikasjon_lokalid,
                     områdeverdi)

# Check for topological and geometrical errors
which(st_is_valid(rec_areas) == FALSE) %>% length()
which(st_is_empty(rec_areas) == TRUE) %>% length()

# Correct invalid geometries
rec_cln <- st_make_valid(rec_areas)

# Check correction worked
which(st_is_valid(rec_cln) == FALSE) %>% length()
which(st_is_empty(rec_cln) == TRUE) %>% length()

# Clean R environment
rm(list = setdiff(ls(), 
                  "rec_cln"))

```

### 7.2. Encroachment areas - DT02
#### 7.2.1 Download DT01 from Zenodo through an API
The process is very similar than for Geonorge. Instead of going through all the steps above, we created a function that users can use to retrieve the downlodable link to the data.
```{r enc_link}
#| eval: true
#| code-summary: "DT02 API download"

# Call for the function that will read the data through an API
source(paste0(functions_path,"/api_zenodo.R"))

# Retrieve downloadable link to the data
encroach_link <- api_zenodo(zenodo_url = "https://zenodo.org/records/15130992")
```

#### 7.2.2 Read DT02
Now that we have the downloadable link, we can read the data in R and clean the R environment.
```{r enc_read}
#| eval: true
#| code-summary: "DT02 - read"

encroach_dt <- st_read(encroach_link)

```

#### 7.2.3 Clean and format the data for analyses
All columns are of relevance to our work but we need to clean the topological and geometrical errors if any.

```{r clean_rec_area}
#| eval: true
#| code-summary: "DT02 - clean"

# Check for topological and geometrical errors
which(st_is_valid(encroach_dt) == FALSE) %>% length()
which(st_is_empty(encroach_dt) == TRUE) %>% length()

# Clean R environment
rm(list = setdiff(ls(), 
                  "rec_cln",
                  "encroach_dt"))

```

The data are all clear and do not need further corrections.


## 8. Spatial units

<!--# Describe the spatial units that you rely on in your analyses. Highlight the spatial units (the resolution) that the indicator values should be interpreted at. Potential spatial delineation data (maksing layers etc.) should be introduced under chapert 7. We recommend using the SEEA EA terminology of Basic Spatial Units (BSU), Ecosystem Asses (EA) and Ecosystem Accounting Area (EAA). -->

## 9. Analyses
The analyses primarily rely on spatial intersections between DT01 and DT02. 

### 9.1 ENCREC_001 - Total encroachment in all recreation areas
For ENCREC_001, we do the intersection with DT01 and DT02 as they are.
```{r clean_rec_area}
#| eval: true
#| code-summary: "DT02 - clean"

# Check for topological and geometrical errors
which(st_is_valid(encroach_dt) == FALSE) %>% length()
which(st_is_empty(encroach_dt) == TRUE) %>% length()

# Clean R environment
rm(list = setdiff(ls(), 
                  "rec_cln",
                  "encroach_dt"))

```

### 9.2 ENCREC_002 - Total encroachment in important and very important recreation areas.
 For ENCREC_002, we first need to select the recreation areas that are considered as "important" and "very important", and then intersect them with DT02.

<!--# 

Use this header for documenting the analyses. Put code in separate code chunks, and annotate the code in between using normal text (i.e. between the chunks, and try to avoid too many hashed out comments inside the code chunks). Add subheaders as needed. 

Code folding is activated, meaning the code will be hidden by default in the html (one can click to expand it).

Caching is also activated (from the top YAML), meaning that rendering to html will be quicker the second time you do it. This will create a folder inside you project folder (called INDICATORID_cache). Sometimes caching created problems because some operations are not rerun when they should be rerun. Try deleting the cash folder and try again.

-->

## 10. Results

<!--# 

Repeat the final results here. Typically this is a map or table of indicator values.

This is typically where people will harvest data from, so make sure to include all relevant output here, but don't clutter this section with too much output either.

-->

## 11. Export file

<!--# 

Optional: Display the code (don't execute it) or the workflow for exporting the indicator values to file. Ideally the indicator values are exported as a georeferenced shape or raster file with indicators values, reference values and errors. You can also chose to export the raw (un-normalised or unscaled variable) as a separate product. You should not save large sptaial output data on GitHub. You can use eval=FALSE to avoid code from being executed (example below - delete if not relevant) 

-->

```{r export}
#| eval: false
```

<!-- You can leave this last part as it is, unless you have a good reason to alter it. -->

::: {.callout-tip collapse="true"}
## Session Info

This information is mainly aimed at developers that need to recreate the R environment that produced the html.
Important note: this information is about the R session that was running when the html was rendered, and if some of the code chunks were not evaluated when the html was rendered, they may have been executed in a different environment.  

```{r sessionInfo}
#| echo: false
sessionInfo()
```

:::
