---
title: "Thematic accounts - encroachment extent accounts in recreation areas"
subtitle: |
  [ENCACC_001] (Extent accounts of all encroachment in municipalities) \
  [ENCACC_005] (Extent accounts of all sværlig vikitig natur lost due to encroachments in municipalities)
# delete the line with the second indicatorID (line 5) 
# if it's not relevant for you.
# In the parentheses you can add a word to distinguish multiple indicators 
# from each other, for example the ecosystem type that they cover
indicatorID: 
  - ENCACC_001
  - ENCACC_002 # delete whole line if not needed
# If you are presenting multiple indicators, and have multiple 
# indicator IDs, then add them as a list here, and make sure you add
# the same number of list entries, 
# and in the same order that corresponds to the indicator IDs,  
# to the meta data fields where the indicators differ. 
# For example, if the indicators cover different ecosystem types, 
# add the ecosystems as a list under 'Ecosystems', but if they are in the
# same realm, same country etc., then you only need to write the realm and
# country once. 
indicatorName: 
  - Extent accounts of all encroachment in municipalities # add more by adding another line starting '-'
  - Extent accounts of all sværlig vikitig natur lost due to encroachments in municipalities
  
folderName: code # name of the folder under 'indicators/'. 
# If it's a single indicator being reported, the folder name should be the same as the indicatorID. 
# If there are multiple ID's then the folder name could look something like XX_YYY_ZZZ-ÅÅÅ.
country: Norway
continent: Europe
# ECT: write ECT here # e.g. A2 - Chemical State characteristics
Realm: 
  - write realm here # e.g. Freshwater (F)
Biome: 
  - write biome here # e.g. F1 Rivers and streams biome
Ecosystem: 
  - write ecosystem here # e.g. F1.3 Freeze-thaw streams
yearAdded: 2025 # the year when the indicator was first submitted to ecRxiv
yearLastUpdate: 2025
status: incomplete # options are complete, incomplete, or deprecated
Version: "000.001" # Beta-versions are numbers incrementally with values below 0. Operational versions can be given the version numbers above 0. Use this terms similar as for software versioning,
VersionComment: This is the first version of the thematic encroachment extent accounts in municipalities being compiled # Describe briefly how this version differs from earlier versions 
url: https://github.com/NINAnor/recereation_encroachment_accounts
# Leave the url as is, but replace XX_YYYY_ZZ with the name of the root folder for the current indicator(s). It should be the same as the indicatorID, but if there are more indicator IDs presented in the same folder, the folder name might be different, e.g. XX_YYYY_AAA-BBB 
Normalised: No # Yes or no depending on whether the metric has been normliased or not.
SpatialAggregationPathway: Describe the aggregation using the standrdaise terminology # e.g. Scale - Transform - Aggregate (length-weighted arithmetic mean) - Truncate
format: 
  html:
    embed-resources: true
    code-fold: true
    toc: true
    toc-title: Contents
    toc-depth: 3
    smooth-scroll: true
execute: 
  cache: true
author:
  - name: Sylvie Clappe             # Enter name
    email: sylvie.clappe@nina.no  # Enter email
    affiliations:
      - id: myID
        name: Norwegian Institute for Nature Research  # Enter affiliations
  - name: David Barton              #  Enter subsequent authors like this, or remove if not relevant
    affiliations:
      - ref: myID               # To reuse affiliations reference the id like this
  - name: Reto Spielhofer             #  Enter subsequent authors like this, or remove if not relevant
    affiliations:
  - ref: myID 
AuthorList: Clappe, S., Barton, D.N., Spielhofer, R.  
date: January 22, 2026 # Enter date
data_availability: none
code_reproducibility: none
open_science_badge: none
callout-icon: false
lightbox: true
css: ../../../style.css
code-links:
      - text: Add a review
        icon: github
        href: https://github.com/NINAnor/ecRxiv
hide: TRUE # Change to FALSE
---

<!--# This is a template for how to document the indicator analyses. Make sure also to not change the order, or modify, the headers, unless you really need to. This is because it easier to read if all the indicators are presented using the same layout. If there is one header where you don't have anything to write, just leave the header as is, and don't write anything below it. If you are providing code, be careful to annotate and comment on every step in the analysis. Before starting it is recommended to fill in as much as you can in the metadata file. This file will populate the initial table in your output.-->

<!--# Load all you dependencies here -->

```{r setup}
#| include: false
#| cache: false
library(knitr)
library(tidyverse)
library(kableExtra)
library(here)
library(anybadger)
library(yaml)
library(tibble)
library(conflicted)
library(httr2)
library(giscoR)
library(zen4R)
library(jsonlite)
library(sf)
library(purrr)
library(numform)
library(DT)
library(tmap)

# weird workarounds for cache issues
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::lag)
conflicts_prefer(dplyr::select)
pull<-dplyr::pull

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Remove dplyr and messages
options(dplyr.summarise.inform = FALSE)
```

```{r source}
#| echo: false
#| cache: false
source(here::here("_common.R"))
```

```{r working_directory}
#| include: false
#| cache: false

# Workaround to force the working directory
# Needs to be set after {r source} otherwise
# {_common.R is not found}
opts_knit$set(root.dir = here:::here())
```


<!--# The following parts are autogenerated. Do not edit. -->

::: {layout-ncol="5"}

```{r statusBadge}
#| echo: false
status_badge(as.character(st))
```

```{r versionBadge}
#| echo: false
version_badge(my_version_number = as.character(version[[1]]),
  folder_name = folder_name)
```

```{r dataBadge}
#| echo: false
data_badge(as.character(badges[[1]]))
```

```{r codeBadge}
#| echo: false
code_badge(as.character(badges[[2]]))
```

```{r openScienceBadge}
#| echo: false
open_science_badge(as.character(badges[[3]]))
```
::: 
 

> **Recommended citation**: `r paste(auth, " ", year, ". ", name, " (ID(s): ", folder_name, ") ", "v. ", version, ". ", url, sep="")`



::: {.callout-tip collapse="true"}
## Metadata
```{r tbl-meta2}
#| tbl-cap: 'Indicator metadata'
#| echo: false
#| warning: false
meta |>
  dplyr::filter(Variable %in% c(
    "Indicator ID",
    "Indicator Name",
    "Continent",
    "Country",
    "Ecosystem Condition Typology Class", 
    "Realm",
    "Biome",
    "Ecosystem",
    "Year added", 
    "Last update",
    "Version",
    "Version comment",
    "Normalised",
    "Spatial aggregation pathway")) |>   
  kable()

```
:::

::: {.callout-tip collapse="true"}

## Log

<!--# Update this logg with short messages for each update -->
- 22 January 2026 - Analyses completed
- XXX - First draft version
:::


<hr />

<!--# Document you work below.  -->

## 1. Summary

<!--# 
With a maximum of 300 words, describe the indicator in general terms as you would to a non-expert. Think of this as a kind of common language summary. Follow the four numbers sub-headers below. under point 2, it is possible under  include a bullet point list of the specific steps in the workflow. 
 -->

1. **Background and rationale**:

<!-- What does the metric represent? Why is this relevant for describing ecosystem condition in this ecosystem?
What are the main anthropogenic impact factors? -->
2. **Methods**:
<!-- What kind of data is used? How is the data customized (modified, estimated, integrated) to fit its purpose as an indicator? -->
3. **Interpretation**:
<!-- Add a very short description of how the indicator values should be interpreted. Here's an example for a bumblebee indicator: An indicator value of 1 signifies an intact bumblebee community. Values decreasing toward zero reflect negative ecological change, specifically the decline of species expected to be present. This metric does not account for potentially positive changes, such as increases in species' abundance or the arrival of new species. Tip: Add text on the line directly following the subheader, and don't introduce blank lines. This way the whole paragraph renders as a block quote. -->
An indicator value of 1 signifies ...
Values decreasing toward zero reflect...

4. **Results and development needs**:
<!-- Briefly, what does the indicator say about the ecosystem condition. What is the current status of  the metric (can it be used or is it still in development), and what immediate tasks remain to improve the indicator or to make the indicator operational -->



## 2. About the underlying data

<!--# Give a general introduction the datasets you have used, and provide more specific details in the sub-chapters.  -->

### 2.1 Spatial and temporal resolution and extent

<!--# Describe the temporal and spatial resolution and extent of the data used -->

### 2.2 Sampling design

<!-- How was the data sampled? Describe potential sampling biases -->

### 2.3 Original units

<!--# What are the original units for the most relevant  variables in the data-->


### 2.3 Instructions for citing, using and accessing data

<!--# Is the data openly available? If not, how can one access it? What are the key references to the datasets?   -->

### 2.4 Additional comments about the dataset

<!--# Optional. Text here -->

## 3. Indicator properties

### 3.1 Collinearities with other indicators

<!--# Describe known collinearities with other metrics (indicators or variables) that could become problematic if they were also included in the same Ecosystem Condition Assessment as the indicator described here. -->


### 3.2 Impact factors

<!--# Describe the main natural and anthropogenic factors that affecst the metric -->

### 3.3 Spatial aggregation

<!-- Describe in word the same aggregation pathway presented in the top yaml (the metadata). If you have scaled against the reference value, it is especially important to know if you did any avergaing or summations etc. beforehand. An example text for a rather simple case could be: "The variable is scaled at the initial polygon level and the point estimate is a area-weighted arithmetic mean of all polygins. There is no transformation, and truncation is not needed as the indicator becomes naturally bound between 0 and 1".  -->


## 5. Uncertainties
### 5.1 General uncertainties
<!--# Describe the main reasons why you might not trust 100% the final value to represent the given ecosystem characteristic for the give ecosystem type. What aspects of this condition metric are most sensitive to subjective decitions? Is the spatial sampling sufficient? Does the choice of modeling framework affect the values? Is the metric relevant do describe ecosystem condition for the give ecosystem type?  -->

Be careful: we did not add any restoration: we just have additions and not "reductions" in the extent of the encroachment. THis is a bias, it might be that the figures are smaller in reality.

In cumulative accounts, we assume emcroachment add over the years but in fact some might have disappeared (restoration as mentioned above)


### 5.2 Reported uncertainties
<!-- Describe the nature of the errors you (hopefully) provide alongside the point estimate for the condition metric. It is especially important to clarify if the errors are inferential (referring to the uncertainty of the point estimate) or descriptive (referreing to the data itself). Here' an example text: "What we report as uncertainties in @tbl-finalTable are inferential and refer to the uncertainty around the point estimates. It does not directly describe variation in the data itself. 
However, the variation in the mean is defined by the spatial variation only, and quartiles are obtained by bootstrapping the set of localities within a region, and calculating the indicator for each of 9999 such samples".   -->



## 6. References

<!--# You can add references manually or use a citation manager and add intext citations as with crossreferencing and hyperlinks. See https://quarto.org/docs/authoring/footnotes-and-citations.html -->

## 7. Import and prepare datasets
In this section, we show how to import the data through API. We also prepare them  for analyses. We first set up the path that will be used during th eproject. Users can change it to match their own path.
```{r path_setup1}
#| eval: true
#| code-summary: "Path setup"

# Path to the project folder
project_path <- here::here()

# Path to the functions folder within the project folder
functions_path <- here::here(project_path, "src/functions")
```

<!--# All the data import and preparation should be done in this chapter, before the analysis chapter. As a minimum, add code for importing data, and describe or show the data structure. You may also explain certain variables, what they represent, and their distributions -->

### 7.1. Encroachment areas - DT01
#### 7.1.1 Download DT01 from Zenodo through an API
The process is very similar than for Geonorge. Instead of going through all the steps above, we created a function that users can use to retrieve the downloadable *content* link to the data. We could have used `zen4R` package but we decided to create our own function to read the data from Zenodo directly instead of downloading them locally before reading them.
```{r enc_link}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT01 API download"

# Call for the function that will read the data through an API
source(paste0(functions_path,"/api_zenodo.R"))

# Retrieve downloadable link to the data
encroach_link <- api_zenodo(zenodo_url = "https://zenodo.org/records/15130992")

```

#### 7.1.2 Read DT01
Now that we have the downloadable link, we can read the data in R and clean the R environment.
```{r enc_read}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT01 - read"

encroach_dt <- st_read(encroach_link)

```

#### 7.1.3 Clean and format the data for analyses
All columns are of relevance to our work but we need to clean the topological and geometrical errors if any.

```{r clean_enc}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "DT01 - clean"

# Check for topological and geometrical errors
which(st_is_valid(encroach_dt) == FALSE) %>% length()
which(st_is_empty(encroach_dt) == TRUE) %>% length()

# Clean R environment
rm(list = setdiff(ls(), 
                  c("encroach_dt",
                    "project_path",
                    "functions_path")))

```

The data are all clear and do not need further corrections.


### 7.2 Map of municipalities in Norway - DT02
#### 7.2.1 Download & read DT02 from GISCO through an API
For DT03, the API process is much simpler as a package has been developed to read data from GISCO via an API: `giscoR`. We just need to use one function to cache the dataset in a temporary directory and read it in R.
```{r lau_down_data}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT02 API download"

# Create a temporary directory to store the cache
temp_dir <- tempdir()

# Download the dataset
lau_no <- gisco_get_lau(year = "2024",
                        epsg = "3035",
                        cache = TRUE,
                        update_cache = TRUE,
                        cache_dir = temp_dir,
                        verbose = FALSE,
                        country = "NO",
                        gisco_id = NULL)

```

#### 7.2.2 Clean and format the data for analyses
We select the columns of relevance to our work and clean the topological and geometrical errors if any. Note that we also need to clean the `GISCO_ID`, which is the municipality number. 

```{r clean_lau}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "DT02 - general clean and format"

# We select the columns of interests
lau_no <- lau_no %>%
            select(GISCO_ID,
                   LAU_NAME) %>%
            mutate(GISCO_ID = sub("NO_", "", .$GISCO_ID)) %>%
            st_transform(., st_crs(encroach_dt)) # project to ESPG : 25832

# Check for topological and geometrical errors
which(st_is_valid(lau_no) == FALSE) %>% length()
which(st_is_empty(lau_no) == TRUE) %>% length()

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "lau_no",
                    "project_path",
                    "functions_path")))
```


### 7.3 Naturtyper - verdsatte - DT03
#### 7.3.1 Download DT03 for Geonorge through an API
Downloading datasets from Geornorge using an API is a bit subtle and we created a function for it.
```{r import_natur_api1}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT03 API download"
# Call for the function that will read the data through an API
source(paste0(functions_path,"/api_geonorge.R"))

# Retrieve and download the data
natur_link <- api_geonorge(geonorge_url = "https://kartkatalog.geonorge.no/metadata/naturtyper-verdsatte/64cbb884-a19d-4356-a114-380cfe4a7314",
                              email_address = "sylvie.clappe@nina.no")

```

#### 7.3.2 Read DT03
Now that the dataset has been downloaded in a temporary file, we can read it in R and clean the R environment. Once the analyses are finished, we will remove the temporary file containing the data.
```{r read_nature}
#| eval: true
#| warning: false
#| message: false
#| code-summary: "DT03 - read the data"

# Read data
naturtyper <- st_read(natur_link[grepl("Naturtyper", natur_link)])
```

#### 7.3.3 Clean and format the data for analyses
We first need to select the column of relevance for our work and clean the topological and geometrical errors if any. The two major issues we have with these data are the spatial and temporal coverages. Spatial coverage because the data do not cover the entire country of Norway. This means that it's not because we don't record one a nature type as occurring that it is indeed not occurring, it might just not have been sampled yet. Also, the nature types have been recorded only once for a wide temporal coverage: from 1905 to 2025. Given that our encroachment data cover 2017-2024, we decided to keep only the data that have been recorded in 2015 or after. We consider that a record in 2025 doesn't mean that the type was not present before. We also assume that nature types are relatively stable and have not change since 2015. We also removed all nature types whose year of report was not stated. This made us keep 72% of the original dataset, which is still very good for analyses.

```{r clean_natur}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "DT03 - clean and format"

# Select columns of relevance
natur_types <- naturtyper %>%
                mutate(year_natur_recorded = substr(.$registreringsdato, 1, 4)) %>%
                select(naturtype,
                       year_natur_recorded,
                       kommunenummer)

# Check for topological and geometrical errors
which(st_is_valid(natur_types) == FALSE) %>% length()
which(st_is_empty(natur_types) == TRUE) %>% length()

# Correct invalid geometries
natur_cln <- st_make_valid(natur_types)

# Check correction worked
which(st_is_valid(natur_cln) == FALSE) %>% length()
which(st_is_empty(natur_cln) == TRUE) %>% length()

# Clean typos in years
natur_cln$year_natur_recorded[natur_cln$year_natur_recorded == "2102"] <- "2012"
natur_cln$year_natur_recorded[natur_cln$year_natur_recorded == "2913"] <- "2013"

# Remove nature types that have been recorded before 2015
natur_cln <- natur_cln %>%
              filter(year_natur_recorded > 2016) %>%
              filter(!is.na(year_natur_recorded))
  

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_cln",
                    "lau_no",
                    "project_path",
                    "functions_path")))

```

## 8. Spatial units

<!--# Describe the spatial units that you rely on in your analyses. Highlight the spatial units (the resolution) that the indicator values should be interpreted at. Potential spatial delineation data (maksing layers etc.) should be introduced under chapert 7. We recommend using the SEEA EA terminology of Basic Spatial Units (BSU), Ecosystem Asses (EA) and Ecosystem Accounting Area (EAA). -->

## 9. Analyses
### 9.1 ENCACC_001
To compile ENCACC_001, we need to perform a spatial intersections between DT01 and DT02. Note that we re-calculate the areas of the encroachment polygons after the intersection has been performed.

*Note: some municipalities had no encroachments detected in certain years. In these cases, the year where 0 encroachment was detected did not not appear. We thus had to add it manually for all municipalities*

```{r encrec001234_analyses}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "ENCACC_001 - intersection"

# Check the CRS of both DT01 and DT02
st_crs(lau_no) == st_crs(encroach_dt)

# Perform spatial intersection
komenc_raw <- st_intersection(lau_no, encroach_dt) %>%
                mutate(area_inter_m2 = unclass(st_area(.)), .after = area)

# Add missing years for accounting. Set the area as 0 for the missing year.
komenc <- komenc_raw %>%
            st_drop_geometry() %>%
            complete(nesting(GISCO_ID, LAU_NAME), year = 2018:2024, fill = list(id = "NA",
                                                                                area = 0,
                                                                                area_inter_m2 = 0))

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_cln",
                    "project_path",
                    "functions_path",
                    "komenc_raw",
                    "komenc")))

```
### 9.2 ENCACC_002
For ENCACC_005, we need a spatial intersection between DT04 and `komenc` as we are interested in the nature types that have been lost due to encroachment within municipalities. Here we have to note though that the only 4,2% of the "særlig viktig natur" occur in the encroachment of municipaplities.

```{r encrec0056_analyses}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "ENCACC_002 - intersection"

# Check the CRS of both encrec and DT04
st_crs(komenc_raw) == st_crs(natur_cln)

# Re-project natur_cln into ESPG : 25832
natur_32 <- st_transform(natur_cln, crs = st_crs(komenc_raw))

# Check CRS of both datasets are the same now
st_crs(natur_32) == st_crs(komenc_raw)

# Perform spatial intersection
encnatur_raw <- komenc_raw %>%
                  select(GISCO_ID,
                         LAU_NAME,
                         id,
                         year,
                         area_inter_m2) %>%
                  rename(area_enc = area_inter_m2,
                         year_enc = year) %>%
                  st_intersection(., natur_32) %>%
                  mutate(area_naturtyper_m2 = unclass(st_area(.)))

# Add missing years for accounting reporting
encnatur <- encnatur_raw %>%
            st_drop_geometry() %>%
            select(!c(kommunenummer, area_enc, id)) %>%
            complete(nesting(GISCO_ID, LAU_NAME, naturtype), 
                     year_enc = 2018:2024,
                     fill = list(area = 0,
                                 area_naturtyper_m2 = 0)) %>%
            mutate(year_natur_recorded = coalesce(as.integer(year_natur_recorded), year_enc))

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "natur_32",
                    "komenc",
                    "encnatur",
                    "encnatur_raw",
                    "komenc_raw")))

```
### 9.3 Uncertainty analyses
There are different sources of uncertainty associated with our results (see 5.1 General uncertainties). We could report **qualitatively** some of these uncertainties and we chose to design a color-coded uncertainty indicator to make the interpretation of the results from the municipalities easier. It also have to be noted that, despite having a quantitative uncertainty for DT01, this uncertainty was quantified at national level and does not apply at municipal level. That is why we could not quantify the uncertainty linked to DT01. We however developed an application that will help in quantify the uncertainty in the prediction of encroachment at municipal level. You want to help? Just go [here](**<LINK>**)! Description on how this qualitative indicator can be compiled is available in section 5.2 Reported uncertainties.

First we attributed a colored-uncertainty to all sources of uncertainty. In te case of ENCACC_001, we just have one source: DT01 prediction. For ENCACC_002, we have two sources of uncertainties: DT01 prediction, and temporal discrepancy between DT01 and DT03. Although we restricted the temporal coverage of DT03 to match the temporal coverage of DT01, an encroachment detected in 2024 on a ''særlig viktig natur'' polygon from 2017 is likely to be more uncertain that if detected on a ''særlig viktig natur'' dating from 2024.
```{r encacc_uncertainty_1}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "Step 1 - attribute colored-uncertainty to all sources of uncertainties per municipalities"

# Load necessary functions
source(paste0(functions_path,"/add_uncertainty_individual.R"))

# Uncertainty ENCACCC_001
unc_001_idv <- komenc %>%
                st_drop_geometry() %>%
                group_by(GISCO_ID, year) %>%
                summarise(n_poly = n()) %>%
                select(!n_poly) %>%
                ungroup() %>%
                mutate(uncertainty_source1 = apply(., 1, function(x) {add_uncertainty_individual(data = x, source1 = TRUE, yrsource1 = "year")}))


# Uncertainty ENCACCC_002
unc_002_idv <- encnatur %>%
                st_drop_geometry() %>%
                group_by(GISCO_ID, year_enc, year_natur_recorded, naturtype) %>%
                summarise(n_poly = n()) %>%
                select(!n_poly) %>%
                ungroup() %>%
                mutate(uncertainty_source1 = apply(., 1, function(x) {add_uncertainty_individual(data = x, source4 = TRUE, yrsource4 = c("year_enc", "year_natur_recorded"))[1]}),
                       uncertainty_source4 = apply(., 1, function(x) {add_uncertainty_individual(data = x, source4 = TRUE, yrsource4 = c("year_enc", "year_natur_recorded"))[2]}))


# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "natur_32",
                    "komenc",
                    "encnatur",
                    "unc_001_idv",
                    "unc_002_idv")))

```

The second step is to attribute ONE "global" color-based uncertainty to each municipality for each accounting year and change between 2018 and 2024. For this we need to create a vector of all the color-based uncertainties per municipality per year. After that we look at the frequency of each of the colors and we take the two most predominant. In case of ties, we take the three most predominant. In case the four color-based uncertainty were all equal in their frequency, we decided to classify the "global" uncertainty as "Orange". To see the "global" color-based uncertainty we assign with each combination of two or three predominant colors of uncertainty, you can refer to our NINA rapport <REPORT NUMBER>. For this we first need to create a vector where we concatenate all the individual color-based uncertainties, and then create a function to produce the "global" uncertainty. As we haven't calculated all the metrics yet: cumulative areas and change area, we only summarise and concatenate all the color uncertainties for each municipality for each year. When we will calculate our metrics, we will then compile the final uncertainty going with each metric. We wrote a function to concatenate all the colored-uncertainty into one column per municipality per accounting year. *Note: the "accounting year" is the year of encroachment as we are compiling encroachment extent accounts. Only the area of accounting is changing among the accounts*.


```{r encacc_uncertainty_2}
#| eval: true
#| warning: false
#| message: false
#| output: false
#| code-summary: "Pre-Step 2 - concatenate colored uncertainty from Step 1"

# Call function to concatenate
source(paste0(functions_path,"/concatenate_individual_uncertainty.R"))

# Concatenate individual colored-uncertainties
unc_001_sum <- concatenate_fn(unc_001_idv,
                              col_column_names = "uncertainty_source1", 
                              kom_num = "GISCO_ID",
                              accounting_year = "year")

unc_002_sum <- concatenate_fn(unc_002_idv,
                              col_column_names = c("uncertainty_source1", 
                                                   "uncertainty_source4"),
                              kom_num = "GISCO_ID",
                              habitat_col = "naturtype")

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "project_path",
                    "functions_path",
                    "natur_32",
                    "komenc",
                    "encnatur",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum")))

```

<!--# 

Use this header for documenting the analyses. Put code in separate code chunks, and annotate the code in between using normal text (i.e. between the chunks, and try to avoid too many hashed out comments inside the code chunks). Add subheaders as needed. 

Code folding is activated, meaning the code will be hidden by default in the html (one can click to expand it).

Caching is also activated (from the top YAML), meaning that rendering to html will be quicker the second time you do it. This will create a folder inside you project folder (called INDICATORID_cache). Sometimes caching created problems because some operations are not rerun when they should be rerun. Try deleting the cash folder and try again.

-->

## 10. Results - Thematic Extent Accounts: encroachment in municipalities
Now we can built the accounting tables and maps based on the spatial intersection we performed in the previous section. We chose to build the thematic accounts over the 2018-2024 period. Maps will be created only for the closing (2024) year of the total accounting period 2018-2024.

### 10.1 Accounting tables
#### 10.1.1 ENCACC_001: Extent accounts of encroachment in municipalities
Here we present the extent accounts table for encroachment areas between 2018 and 2024 within each municipality and fylke in Norway. These accounts present both the yearly area of encroachment and the cumulative area of encroachment over the years. We also present a total for Norway. We added the uncertainty as color-based indicators presented in 5.2 Reported uncertainties. We first present the table per municipalities and then per fylke.

*Note: the function calculating the change compiles it as area_closing_year - area_opening_year. This is because we expect an increase in encroachment so we want to show how we gain in built-in areas.*

*Note: We consider the uncertainty of the change in area between the opening and closing accounting years, the same as the cumulative uncertainty for the closing accounting year. This is because when we calculate a change we refer to all the accounting period as the cumulative area for the last accounting year is calculated based on all the accounting years. In our code, add_change_encroachment_area() automatically duplicate the uncertainty of the last accounting year for the change in area.*

```{r res_acc_tables1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "ENCACC_001 - Create accounting table per municipalities"

## Call function to add change in areas 2018-2024
source(paste0(functions_path, "/add_change_encroachment_area.R"))

## Call function to add fylke
source(paste0(functions_path, "/add_fylke.R"))

## Call function to compile individual uncertainty
source(paste0(functions_path,"/add_uncertainty_individual.R"))

## Call function to compile global uncertainty
source(paste0(functions_path, "/add_uncertainty_global.R"))

## ENCACC_001 accounting table
# Accounts for all municipalities
encacc_001_kom <- komenc %>%
                    st_drop_geometry() %>%
                    select(GISCO_ID,
                           year,
                           LAU_NAME,
                           area_inter_m2) %>%
                    mutate(fylke = apply(., 1, function(row) {add_fylke(kommune_nummer = row["GISCO_ID"])}), .before = GISCO_ID) %>%
                    group_by(fylke, GISCO_ID, LAU_NAME, year) %>%
                    summarise(area_dekar_yearly = round2(sum(area_inter_m2)/1000, 2)) %>%
                    mutate(area_dekar_cumulative = round2(cumsum(area_dekar_yearly), 2)) %>%
                    mutate(indicator_id = "ENCACC_001", .before = "fylke") 

# Add uncertainties to encacc_001_kom
encacc_001_kom_un <- encacc_001_kom %>%
                      left_join(., unc_001_sum, by = c("GISCO_ID", "year")) %>%
                      add_uncertainty_global(., 
                                             uncertainty_col_name = "unc_concat_full", 
                                             kom_num = "GISCO_ID", 
                                             accounting_year = "year") %>%
                      select(!unc_concat_full)


# Total across all municipalities
encacc_001_tot <- encacc_001_kom %>%
                    group_by(year) %>%
                    summarise(area_dekar_yearly = sum(area_dekar_yearly)) %>%
                    mutate(area_dekar_cumulative = round2(cumsum(area_dekar_yearly), 2)) %>%
                    mutate(indicator_id = "ENCACC_001",
                           fylke = "00",
                           GISCO_ID = "0000",
                           LAU_NAME = "Norway", .before = year)

# Add uncertainty to encacc_001_tot

encacc_001_tot_un <-encacc_001_tot  %>%
                      mutate(uncertainty_source1 = apply(., 
                                                         1, 
                                                         function(x){add_uncertainty_individual(data = x, source1 = TRUE, yrsource1 = "year")})) %>%
                      
                      add_uncertainty_global(., 
                                             uncertainty_col_name = "uncertainty_source1", 
                                             kom_num = "GISCO_ID", 
                                             accounting_year = "year") %>%
                      select(!uncertainty_source1)

# Add change to all tables
encacc_001_kom_un$year <- as.character(encacc_001_kom_un$year)

encacc_001_kom_un <- encacc_001_kom_un %>%
                      group_split(GISCO_ID) %>% 
                      map(~ add_change_encroachment_area(., 
                                                         year_column = "year",
                                                         area1 = "area_dekar_cumulative")) %>% 
                      compact() %>%
                      bind_rows()

encacc_001_tot_un$year <- as.character(encacc_001_tot_un$year)

encacc_001_tot_un <- encacc_001_tot_un %>%
                      add_change_encroachment_area(., 
                                                   year_column = "year",
                                                   area1 = "area_dekar_cumulative") %>% 
                      compact() %>%
                      bind_rows()

# Correct uncertainty for global_yearly_uncertainty when accounting year is "Change 2018-2024"
encacc_001_kom_un[which(encacc_001_kom_un$year == "Change 2018-2024"), "global_yearly_uncertainty"] <- "NA"
encacc_001_tot_un[which(encacc_001_tot_un$year == "Change 2018-2024"), "global_yearly_uncertainty"] <- "NA"

# Final accounting table               
encacc_001_acc <- bind_rows(encacc_001_kom_un,
                            encacc_001_tot_un)

encacc_001_acc$fylke[which(encacc_001_acc$fylke == "00")] <- NA
encacc_001_acc$GISCO_ID[which(encacc_001_acc$GISCO_ID == "0000")] <- NA

colnames(encacc_001_acc) <- c("Indikatorens navn",
                              "Fylke nummer",
                              "Kommune nummer",
                              "Kommune navn",
                              "Regnskapsår",
                              "Årlig areal i dekar",
                              "Kumulativ areal i dekar",
                              "Usikkerhet for årlig areal",
                              "Usikkerhet for kumulativ areal")


max_kom <- encacc_001_acc %>%
            as.data.frame() %>%
            filter(Regnskapsår == "2024") %>%
            #filter(Regnskapsår == "Change 2018-2024") %>%
            filter(!`Kommune navn` == "Norway") %>%
            arrange(desc(.$`Kumulativ areal i dekar`)) %>%
            filter(`Kumulativ areal i dekar`< 35)

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "add_change_encroachment_area",
                    "add_fylke",
                    "add_uncertainty_individual",
                    "add_uncertainty_global",
                    "encacc_001_acc",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum")))


# Display the table dynamically
datatable(encacc_001_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list( dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                          columnDefs = list(list(targets = c(which(colnames(encacc_001_acc) == "Usikkerhet for årlig areal"),
                                                             which(colnames(encacc_001_acc) == "Usikkerhet for kumulativ areal")),
                                                 visible = FALSE))),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Tabell 1 – Tematiske arealregnskap: areal av all nedbygging i kommunner i perioden 2018–2024."))) %>%
    formatCurrency(
    columns = c("Årlig areal i dekar", "Kumulativ areal i dekar"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",") %>%
  ## Example: color Årlig areal i dekar by var1
  formatStyle(
    "Årlig areal i dekar",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red", "NA"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A", "")
    ),
    valueColumns = "Usikkerhet for årlig areal"           # column used for the intervals
  ) %>%
  ## Example: color Kumulativ areal i dekar by var2
  formatStyle(
    "Kumulativ areal i dekar",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A")
    ),
    valueColumns = "Usikkerhet for kumulativ areal"           # column used for the intervals
  )

```


```{r res_acc_tables2}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "ENCACC_001 - Create accounting table per Fylke"

# Call function to concatenate
source(paste0(functions_path,"/concatenate_individual_uncertainty.R"))

# Create the accounting table based on encacc_001_acc
encacc_001_fy <- encacc_001_acc %>%
                  filter(!Regnskapsår == "Change 2018-2024") %>%
                  filter(!`Kommune navn` == "Norway") %>%
                  group_by(`Indikatorens navn`, `Fylke nummer`, Regnskapsår) %>%
                  summarise(yearly_area = sum(`Årlig areal i dekar`)) %>%
                  mutate(cumulative_area = cumsum(yearly_area))

# Add uncertainties to encacc_001_fy
encacc_001_fy_un <-  unc_001_idv %>%
                      mutate(fylke = add_fylke(kommune_nummer = GISCO_ID)[[1]]) %>%
                      select(!GISCO_ID) %>%
                      concatenate_fn(.,
                                     col_column_names = c("uncertainty_source1"),
                                     kom_num = "fylke",
                                     accounting_year = "year") %>%
                      add_uncertainty_global(., 
                                                      kom_num = "fylke",
                                                      uncertainty_col_name = "unc_concat_full",
                                                      accounting_year = "year") %>%
                      select(!unc_concat_full) %>%
                      mutate(year = as.character(year)) %>%
                      left_join(encacc_001_fy, ., join_by(`Fylke nummer` == "fylke", "Regnskapsår" == "year")) %>%
                      mutate(fylke_name = add_fylke(fylke_nummer = `Fylke nummer`), .after = `Fylke nummer`)


# Add change to encacc_001_fy_un
encacc_001_fy_unc <- encacc_001_fy_un %>%
                      group_split(`Fylke nummer`) %>% 
                      map(~ add_change_encroachment_area(., 
                                                         year_column = "Regnskapsår",
                                                         area1 = "cumulative_area")) %>% 
                      compact() %>%
                      bind_rows()

# Correct uncertainty for global_yearly_uncertainty when accounting year is "Change 2018-2024"
encacc_001_fy_unc[which(encacc_001_fy_unc$Regnskapsår == "Change 2018-2024"), "global_yearly_uncertainty"] <- "NA"
encacc_001_fy_unc[which(encacc_001_fy_unc$Regnskapsår == "Change 2018-2024"), "global_yearly_uncertainty"] <- "NA"

# Modify colnames
colnames(encacc_001_fy_unc) <- c("Indikatorens navn",
                                 "Fylke nummer",
                                 "Fylke navn",
                                 "Regnskapsår",
                                 "Årlig areal i dekar",
                                 "Kumulativ areal i dekar",
                                 "Usikkerhet for årlig areal",
                                 "Usikkerhet for kumulativ areal")
                                        



# Add total across fylke
encacc_001_fy_tot <- encacc_001_acc %>%
                        filter(`Kommune navn` == "Norway") %>% 
                        rename("Fylke navn" = "Kommune navn") %>%
                        select(!`Kommune nummer`)
                        
  
encacc_001_fy_acc <- bind_rows(encacc_001_fy_unc,
                               encacc_001_fy_tot)

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "add_change_encroachment_area",
                    "add_fylke",
                    "add_uncertainty_individual",
                    "add_uncertainty_global",
                    "concatenate_fn",
                    "encacc_001_acc",
                    "encacc_001_fy_acc",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum")))


# Display the table dynamically
datatable(encacc_001_fy_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list( dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                          columnDefs = list(list(targets = c(which(colnames(encacc_001_fy_acc) == "Usikkerhet for årlig areal"),
                                                             which(colnames(encacc_001_fy_acc) == "Usikkerhet for kumulativ areal")),
                                                 visible = FALSE))),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Tabell 2 – Tematiske arealregnskap: areal av all nedbygging i fylker i perioden 2018–2024."))) %>%
    formatCurrency(
    columns = c("Årlig areal i dekar", "Kumulativ areal i dekar"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",") %>%
  ## Example: color Årlig areal i dekar by var1
  formatStyle(
    "Årlig areal i dekar",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red", "NA"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A", "")
    ),
    valueColumns = "Usikkerhet for årlig areal"           # column used for the intervals
  ) %>%
  ## Example: color Kumulativ areal i dekar by var2
  formatStyle(
    "Kumulativ areal i dekar",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A")
    ),
    valueColumns = "Usikkerhet for kumulativ areal"           # column used for the intervals
  )

```

#### 10.1.2 ENCACC_002: Extent accounts of encroachments in all sværlig vikitig natur lost due to encroachments in municipalities
Here we present the extent accounts table for "sværlig vikitig natur" lost to encroachment between 2018 and 2024 within each municipality in Norway. These accounts present both the yearly area of encroachment and the cumulative area of encroachment over the years. We also present a total for Norway. As previously, we first present the results per municipality sand then aggregated at fylke's level.

As the loss of the nature types did not happen homogeneously within municipalities, these accounts only present both the total area lost by nature types over the 2018-2024 period. This is to avoid confusion and present the bigger picture, instead of having different years for different municipalities and nature types. We also present a total across all municipalities in Norway which mapped their recreation areas.


```{r res_acc_tables3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "ENCACC_002 - Create accounting table per municpalities"

## ENCACC_002 accounting table
# Accounts for all municipalities
encacc_002_kom <- encnatur %>%
                    st_drop_geometry() %>%
                    select(GISCO_ID,
                           LAU_NAME,
                           naturtype,
                           year_enc,
                           area_naturtyper_m2) %>%
                    mutate(fylke = add_fylke(kommune_nummer = GISCO_ID)[[1]], .before = GISCO_ID) %>%
                    group_by(fylke, GISCO_ID, LAU_NAME, naturtype) %>%
                    summarise(total_loss_dekar = round2(sum(area_naturtyper_m2)/1000, 2)) %>%
                    mutate(indicator_id = "ENCACC_002", .before = "fylke") %>%
                    mutate(accounting_period = "2018-2024", .before = "total_loss_dekar")

# Add uncertainties to encacc_002_kom
encacc_002_kom_un <- encacc_002_kom %>%
                      left_join(., unc_002_sum, by = c("GISCO_ID", "naturtype"))  %>%
                      add_uncertainty_global(., 
                                             uncertainty_col_name = "unc_concat_full", 
                                             kom_num = "GISCO_ID", 
                                             habitat_accounting = "naturtype") %>%
                      select(!unc_concat_full) %>%
                      rename(global_period_uncertainty = global_habitat_uncertainty)

# National total across all municipalities
encacc_002_tot <- encacc_002_kom %>%
                    group_by(indicator_id, accounting_period) %>%
                    summarise(total_loss_dekar = sum(total_loss_dekar)) %>%
                    mutate(fylke = "00",
                           GISCO_ID = "0000",
                           LAU_NAME = "Norway", .before = accounting_period) 

# Add uncertainties to encacc_002_tot
encacc_002_tot_un <-  unc_002_idv %>%
                        mutate(national = "0000",
                               accounting_period = "2018-2024") %>%
                        concatenate_fn(.,
                              col_column_names = c("uncertainty_source1", 
                                                   "uncertainty_source4"),
                              kom_num = "national",
                              accounting_year = "accounting_period") %>%
                        add_uncertainty_global(., 
                                               kom_num = "national",
                                               uncertainty_col_name = "unc_concat_full",
                                               accounting_year = "accounting_period") %>%
                      ungroup() %>%

                      select(!c(unc_concat_full, 
                                global_cumulative_uncertainty,
                                national,
                                accounting_period)) %>%
  
                      cbind(encacc_002_tot, .) %>%
                      mutate(naturtype = "Alle", .after = LAU_NAME) %>%
                      rename(global_period_uncertainty = global_yearly_uncertainty)


# Final accounting table               
encacc_002_acc <- bind_rows(encacc_002_kom_un,
                            encacc_002_tot_un)

encacc_002_acc$fylke[which(encacc_002_acc$fylke == "00")] <- NA
encacc_002_acc$GISCO_ID[which(encacc_002_acc$GISCO_ID == "0000")] <- NA


colnames(encacc_002_acc) <- c("Indikatorens navn",
                              "Fylke nummer",
                              "Kommune nummer",
                              "Kommune navn",
                              "Særlig viktig naturtype",
                              "Regnskapsperiode",
                              "Areal særlig viktige naturtyper gått tapt (dekar)",
                              "Usikkerhet for areal i regnskapsperiode")


# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "add_change_encroachment_area",
                    "add_fylke",
                    "add_uncertainty_individual",
                    "add_uncertainty_global",
                    "concatenate_fn",
                    "encacc_001_acc",
                    "encacc_001_fy_acc",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum",
                    "encacc_002_acc")))


# Display the table dynamically
datatable(encacc_002_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list( dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                          columnDefs = list(list(targets = c(which(colnames(encacc_002_acc) == "Usikkerhet for areal i regnskapsperiode")),
                                                 visible = FALSE))),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Tabel 3 - Tematiske arealregnskap: areal av særlig viktig natur tapt på grunn av nedbygging i kommuner i perioden 2018–2024."))) %>%
    formatCurrency(
    columns = c("Areal særlig viktige naturtyper gått tapt (dekar)"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",") %>%
  ## Example: color Årlig areal i dekar by var1
  formatStyle(
    "Areal særlig viktige naturtyper gått tapt (dekar)",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red", "NA"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A", "")
    ),
    valueColumns = "Usikkerhet for areal i regnskapsperiode"           # column used for the intervals
  )

```


```{r res_acc_tables4}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "ENCACC_002 - Create accounting table per fylke"

# Create the accounting table based on encacc_001_acc
encacc_002_fy <- encacc_002_acc %>%
                  filter(!`Kommune navn` == "Norway") %>%
                  group_by(`Indikatorens navn`, `Fylke nummer`, `Særlig viktig naturtype`, Regnskapsperiode) %>%
                  summarise(total_area = sum(`Areal særlig viktige naturtyper gått tapt (dekar)`))

# Add uncertainties to encacc_001_fy
encacc_002_fy_un <-  unc_002_idv %>%
                      mutate(fylke = add_fylke(GISCO_ID)[[1]]) %>%
                      concatenate_fn(.,
                                     col_column_names = c("uncertainty_source1",
                                                          "uncertainty_source4"),
                                     kom_num = "fylke",
                                     habitat_col = "naturtype") %>%
                      add_uncertainty_global(., 
                                                      kom_num = "fylke",
                                                      uncertainty_col_name = "unc_concat_full",
                                                      habitat_accounting = "naturtype") %>%
                      select(!unc_concat_full) %>%
                      left_join(encacc_002_fy, ., join_by(`Fylke nummer` == "fylke", "Særlig viktig naturtype" == "naturtype"))  %>%
                      mutate(fylke_name = add_fylke(fylke_nummer = `Fylke nummer`), .after = `Fylke nummer`)

# Modify colnames
colnames(encacc_002_fy_un) <- c("Indikatorens navn",
                                 "Fylke nummer",
                                 "Fylke navn",
                                 "Særlig viktig naturtype",
                                 "Regnskapsperiode",
                                 "Areal særlig viktige naturtyper gått tapt (dekar)",
                                 "Usikkerhet for areal i regnskapsperiode")


# Add total across fylke
encacc_002_fy_tot <- encacc_002_acc %>%
                        filter(`Kommune navn` == "Norway") %>% 
                        rename("Fylke navn" = "Kommune navn") %>%
                        select(!`Kommune nummer`)
  
encacc_002_fy_acc <- bind_rows(encacc_002_fy_un,
                               encacc_002_fy_tot)

# Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "add_change_encroachment_area",
                    "add_fylke",
                    "add_uncertainty_individual",
                    "add_uncertainty_global",
                    "concatenate_fn",
                    "encacc_001_acc",
                    "encacc_001_fy_acc",
                    "encacc_002_acc",
                    "encacc_002_fy_acc",
                    "unc_001_idv",
                    "unc_002_idv",
                    "unc_001_sum",
                    "unc_002_sum")))


# Display the table dynamically
datatable(encacc_002_fy_acc,
          extensions = 'Buttons',
          filter = list(position = 'top', clear = TRUE), 
          options = list( dom = 'Bfrtip',
                          buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                          columnDefs = list(list(targets = c(which(colnames(encacc_002_fy_acc) == "Usikkerhet for areal i regnskapsperiode")),
                                                 visible = FALSE))),
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: center; font-size: 1.5em;",
                                            htmltools::strong("Tabel 4 - Tematiske arealregnskap: areal av særlig viktig natur tapt på grunn av nedbygging i kommuner i perioden 2018–2024."))) %>%
    formatCurrency(
    columns = c("Areal særlig viktige naturtyper gått tapt (dekar)"),
    currency = "",                
    interval = 3,                 
    mark = " " ,
    dec.mark = ",") %>%
      
  formatStyle(
    "Areal særlig viktige naturtyper gått tapt (dekar)",
    backgroundColor = styleEqual(
      c("Green",   "Yellow",    "Orange",    "Red", "NA"),
      c("#94F4B6", "#FFF085", "#FFC68A", "#FF8A8A", "")
    ),
    valueColumns = "Usikkerhet for areal i regnskapsperiode" )

```


### 10.2 Maps
#### 10.2.1 ENCACC_001: Map of the total encroachment area in municipalities over 2018-2024
We first prepare the dataset for mapping then map it both using an interactive and static map. Due to memory issues in the html, we only display the static map.
```{r map_enc001_1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Prepare dataset for mapping"

## Import LAU dataset to then join the name of the municipalities
kom_navn <- lau_no %>%
              select(GISCO_ID)

## Prepare dataset
# Municipalities
encacc_001_map_mun <- encacc_001_acc %>%
                        filter(!c(`Kommune navn` == "Norway" | Regnskapsår == "Change 2018-2024")) %>%
                        filter(Regnskapsår == 2024) %>%
                        left_join(., kom_navn, join_by( `Kommune nummer` == "GISCO_ID")) %>%
                        select("Fylke nummer",
                               "Kommune navn",
                               "Kumulativ areal i dekar",
                               "Usikkerhet for kumulativ areal",
                               "geometry") %>%
                        rename("Kumulativ areal av inngrep i dekar" = "Kumulativ areal i dekar") %>%
                        st_as_sf()

# Fylke

################################

# Replace labels of uncertainty
encacc_001_map_mun$`Usikkerhet for kumulativ areal`[which(encacc_001_map_mun$`Usikkerhet for kumulativ areal` == "Green")]<- "Quite certain"

encacc_001_map_mun$`Usikkerhet for kumulativ areal`[which(encacc_001_map_mun$`Usikkerhet for kumulativ areal` == "Yellow")]<- "Rather certain"

encacc_001_map_mun$`Usikkerhet for kumulativ areal`[which(encacc_001_map_mun$`Usikkerhet for kumulativ areal` == "Orange")]<- "Rather uncertain"

encacc_001_map_mun$`Usikkerhet for kumulativ areal`[which(encacc_001_map_mun$`Usikkerhet for kumulativ areal` == "Red")]<- "Quite uncertain"


```

```{r map_enc001_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Interactive map of total encroachment area in municipality over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/interactive_map.R"))

# Run function
interactive_map(data1 = encacc_001_map, 
                data2 = encacc_001_map, 
                fill_variable1 = "Kumulativ areal av nedbygging i dekar", 
                fill_variable2 = "Usikkerhet for kumulativ areal", 
                legend_title1 = "Totalt areal av nedbygging (dekar)", 
                legend_title2 = "Usikkerhet",
                color_values_map2 = c("Quite certain" = "#94F4B6",
                                      "Rather certain" = "#FFF085", 
                                      "Rather uncertain" = "#FFC68A",
                                      "Quite uncertain" = "#FF8A8A"),
                map_title1 = "Totalt areal av nedbygging (dekar) i fylker over 2018-2024", 
                map_title2 = "Usikkerhet knyttet til det totale arealet av nedbygging over 2018-2024")

```

```{r map_enc001_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of total encroachment area in municipality over 2018-2024"

## Static maps
# Call for the function to do static mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = encacc_001_map, 
                data2 = encacc_001_map, 
                fill_variable1 = "Kumulativ areal av inngrep i dekar", 
                fill_variable2 = "Usikkerhet for kumulativ areal", 
                legend_title1 = "Totalt areal av nedbygging (dekar)", 
                legend_title2 = "Usikkerhet",
                color_values_map2 = c("Quite certain" = "#94F4B6",
                                      "Rather certain" = "#FFF085", 
                                      "Rather uncertain" = "#FFC68A",
                                      "Quite uncertain" = "#FF8A8A"),
                map_title1 = "Totalt areal av nedbygging (dekar) i fylker over 2018-2024", 
                map_title2 = "Usikkerhet til det totale arealet av nedbygging over 2018-2024")

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "encacc_001_acc",
                    "encacc_002_acc")))

```

```{r map_enc001_4}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Clean R environment"

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "encacc_001_acc",
                    "encacc_002_acc")))

```

#### 10.2.2 ENCACC_001: Share of the total encroachment area in municipalities over 2018-2024
We first prepare the dataset for mapping then map it both using an interactive and static map. Due to memory issues in the html, we only display the static map. 

```{r map_enc001_perc_1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Prepare data for mapping"

## Import LAU dataset to then join the name of the municipalities
kom_navn <- lau_no %>%
              select(GISCO_ID) %>%
              mutate(area_kom = round2(unclass(st_area(.))/1000, 2))


## Calculate total area per fylke
fylke_total_area <- encacc_001_acc %>%
                      left_join(., kom_navn, join_by(`Kommune nummer` == "GISCO_ID")) %>%
                      st_drop_geometry() %>%
                      select(`Fylke nummer`, area_kom) %>%
                      group_by(`Fylke nummer`) %>%
                      summarise(area_fylke = sum(area_kom))
 


## Prepare dataset
encacc_001_perc_map <- encacc_001_acc %>%
                          filter(!c(`Kommune navn` == "Norway" | Regnskapsår == "Change 2018-2024")) %>%
                          filter(Regnskapsår == 2024) %>%
                          left_join(., kom_navn, join_by( `Kommune nummer` == "GISCO_ID")) %>%
                          select("Fylke nummer",
                                 "Kommune navn",
                                 "Kumulativ areal i dekar",
                                 "area_kom",
                                 "geometry") %>%
                          mutate(share = round2(`Kumulativ areal i dekar`/area_kom * 100, 2)) %>%
                          st_as_sf()

fylke_map <- encacc_001_perc_map %>%
              group_by(`Fylke nummer`) %>%
              summarise("Kumulativ areal i dekar" = sum(`Kumulativ areal i dekar`)) %>%
              left_join(., fylke_total_area, join_by( "Fylke nummer" == "fylke")) %>%
              mutate(share = round2(`Kumulativ areal i dekar`/area_fylke * 100, 2))

```

```{r map_enc001_perc_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Interactive map of the share of encroachments in municipalities and fylkes over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/interactive_map.R"))

# Run function
interactive_map(data1 = fylke_map, 
                data2 = encacc_001_perc_map, 
                fill_variable1 = "share", 
                fill_variable2 =  "share", 
                legend_title1 = "Fylke - Prosentandel (%) ", 
                legend_title2 = "Kommune - Prosentandel (%) ", 
                map_title1 = "Prosentandel (%) av totalt inngrep i fylker", 
                map_title2 = "Prosentandel (%) av totalt inngrep i kommuner")

```

```{r map_enc001_perc_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of the share of recreation area lost to encroachments in all recreation areas per municipality and fylke over 2018-2024"

## Static maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = fylke_map, 
           data2 = encacc_001_perc_map, 
           fill_variable1 = "share", 
           fill_variable2 =  "share", 
           legend_title1 = "Fylke - Prosentandel (%) ", 
           legend_title2 = "Kommune - Prosentandel (%) ", 
           map_title1 = "Prosentandel (%) av totalt inngrep i fylker", 
           map_title2 = "Prosentandel (%) av totalt inngrep i kommuner")

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "encacc_001_acc",
                    "encacc_002_acc")))

```

```{r map_enc001_perc_4}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Clean R environment"

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kom_navn",
                    "encacc_001_acc",
                    "encacc_002_acc")))

```

#### 10.2.3 ENCACC: Map of the total area of "sæarlig viktig natur" lost due to encroachment in municipalities over 2018-2024
We first prepare the dataset for mapping then map it both using an interactive and static map. Due to memory issues in the html, we only display the static map. As few municipalities have "særlig viktig natur" in their encroachments, we do not provide summary statistics at Fylke Level.
```{r map_enc002_1}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Prepare dataset for mapping"

## Prepare dataset
encacc_002_map <- encacc_002_acc %>%
                    filter(!c(`Kommune navn` == "Norway" | Regnskapsperiode == "Change 2018-2024")) %>%
                    left_join(kom_navn, ., join_by("GISCO_ID" == `Kommune nummer`)) %>%
                    select("Fylke nummer",
                           "Kommune navn",
                           "Areal særlig viktige naturtyper gått tapt (dekar)",
                           "geometry") %>%
                    st_as_sf()

```

```{r map_enc002_2}
#| eval: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Interactive map of total area of ''særlig viktig natur'' lost due to encroachment in municipality and fylke over 2018-2024"

## Interactive maps -- disabled as too heavy to render in html
# Call for the function to do interactive mapping
source(paste0(functions_path, "/interactive_map.R"))

# Run function
interactive_map(data1 = encacc_002_map, 
                fill_variable1 = "Areal særlig viktige naturtyper gått tapt (dekar)", 
                legend_title1 = "Kommune - Areal av inngrep (dekar)", 
                map_title1 = "Totalt areal av ''særlig viktig natur'' gått tapt (dekar) på inngrep i kommuner")
```

```{r map_enc002_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Static map of total encroachment area in important and very important recreation areas per municipality and fylke over 2018-2024"

## Static maps
# Call for the function to do static mapping
source(paste0(functions_path, "/static_map.R"))

# Run function
static_map(data1 = encacc_002_map,
           fill_variable1 = "Areal særlig viktige naturtyper gått tapt (dekar)", 
           legend_title1 = "Kommune - Areal av inngrep (dekar)", 
           map_title1 = "Totalt areal av ''særlig viktig natur''  (dekar) som er gått tapt på inngrep i kommuner")



## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kon_navn",
                    "encacc_001_acc",
                    "encacc_002_acc")))
```


```{r map_enc004_3}
#| eval: true
#| warning: false
#| message: false
#| output: true
#| code-summary: "Clean R environment"

## Clean R environment
rm(list = setdiff(ls(), 
                  c("lau_no",
                    "encroach_dt",
                    "natur_32",
                    "project_path",
                    "functions_path",
                    "komenc",
                    "encrec",
                    "encnatur",
                    "kon_navn",
                    "encacc_001_acc",
                    "encacc_002_acc")))
```

<!--# 

Repeat the final results here. Typically this is a map or table of indicator values.

This is typically where people will harvest data from, so make sure to include all relevant output here, but don't clutter this section with too much output either.

-->

## 11. Export file

<!--# 

Optional: Display the code (don't execute it) or the workflow for exporting the indicator values to file. Ideally the indicator values are exported as a georeferenced shape or raster file with indicators values, reference values and errors. You can also chose to export the raw (un-normalised or unscaled variable) as a separate product. You should not save large sptaial output data on GitHub. You can use eval=FALSE to avoid code from being executed (example below - delete if not relevant) 

-->

```{r export}
#| eval: false
```

<!-- You can leave this last part as it is, unless you have a good reason to alter it. -->

::: {.callout-tip collapse="true"}
## Session Info

This information is mainly aimed at developers that need to recreate the R environment that produced the html.
Important note: this information is about the R session that was running when the html was rendered, and if some of the code chunks were not evaluated when the html was rendered, they may have been executed in a different environment.  

```{r sessionInfo}
#| echo: false
sessionInfo()
```

:::
